---
title: "diagnostic_plots"
author: "Pascale Goertler"
date: "2022-08-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Diagnostic plots (8/4/2022)
# look at each varaible against chl, by region and inundation season
# look at each varaible against dowy with chl and possible interactions

```{r library}
library(ggplot2)
library(lubridate)
library(dplyr)
library(tidyverse)
```

## Get data

```{r data, echo=FALSE}
# from make_gam_byregion.R
alldata0 <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy))

inundation_yes <- subset(alldata0, inundation ==1)
min(inundation_yes$dowy) #65
max(inundation_yes$dowy) #215

alldata0$season <- ifelse(alldata0$dowy >= 65 & alldata0$dowy <= 215, "inundation_possible", "inundation_not_possible")
```

# each variable
```{r chl, echo=FALSE}
ggplot(alldata0, aes(WTmwk, chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(WTrangemwk, chlorophyll, colour = inundation)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(Q_sday, chlorophyll, colour = inundation)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(inund_days, chlorophyll, colour = WTrangemwk)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
# highest chl values for 'below' are during inudation not possible season

ggplot(alldata0, aes(total_inund_last_year, chlorophyll, colour = inundation)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
# highest chl values in 'yolo' were no or low inundation last year with inundation during chl collection (4 out of 6)

ggplot(alldata0, aes(days_since_last_inundation, chlorophyll, colour = Q_sday)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(Sradmwk, chlorophyll, colour = inund_days)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
```


# day of water year vs mean water temperature for the week
```{r WT, echo=FALSE}
ggplot(alldata0, aes(dowy, WTmwk, colour = log(Q_sday), size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
# something si wring with the Qsday in Cache... returning NAs

ggplot(alldata0, aes(dowy, WTmwk, colour = inund_days, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, WTmwk, colour = total_inund_last_year, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, WTmwk, colour = days_since_last_inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
```
# day of water year vs range in water temperature for the week
```{r range, echo=FALSE}
ggplot(alldata0, aes(dowy, WTrangemwk, size = Q_sday, colour = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
# range is only high in yolo and Cache (could be habitat complexity from inundation?)

ggplot(alldata0, aes(dowy, WTrangemwk, colour = total_inund_last_year, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, WTrangemwk, colour = days_since_last_inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

# not as seasonal as we thought... 
```

# look into drainage period for yolo
```{r drainage, echo=FALSE}
yolo <- subset(alldata0, region == "yolo" & water_year == "2011" | water_year == "2016" | water_year == "2017" | water_year == "2019")

hist(yolo$WTrangemwk)

yolo$cols <- ifelse(yolo$WTrangemwk > 14, "red",
                    ifelse(yolo$WTrangemwk <14 & yolo$WTrangemwk >= 5, "blue",
                           ifelse(yolo$WTrangemwk <5 & yolo$WTrangemwk > 1, "darkgreen", "orange")))

plot(yolo$dowy, yolo$Q_sday, cex = (yolo$inund_days/10), color = yolo$cols)

ggplot(yolo, aes(dowy, Q_sday, size = inund_days)) + 
  geom_point(aes(colour = WTrangemwk))+
  scale_color_gradientn(colours = yolo$cols,
                       values = yolo$WTrangemwk)+ 
  facet_grid(vars(water_year))

hist(yolo$WTrangemwk)

# two large values in 2016 (might be outliers?) throughing off the gradient in plot
yolo_sub <- subset(yolo, WTrangemwk < 14)
ggplot(yolo_sub, aes(dowy, Q_sday, size = inund_days)) + 
  geom_point(aes(colour = WTrangemwk))+
  facet_grid(vars(water_year))

# WTrangemwk is higher after flooding, but not showing a clear pattern with drainage (continues into summer)

hist(alldata0$WTrangemwk) # might need to investigate outliers for this variable 
boxplot(alldata0$WTrangemwk)
dotchart(alldata0$WTrangemwk)
```

# day of water year vs solar radiation 
```{r Srad, echo=FALSE}

ggplot(alldata0, aes(dowy, Sradmwk, colour = Q_sday, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = total_inund_last_year, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = days_since_last_inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = chlorophyll, size = inund_days)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = WTmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Sradmwk, colour = WTrangemwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
```

# flow
```{r flow, echo=FALSE}
ggplot(alldata0, aes(dowy, Q_sday, colour = inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = Sradmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = WTmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = WTrangemwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = total_inund_last_year, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = days_since_last_inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, Q_sday, colour = chlorophyll, size = inund_days)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
```

# inundation
```{r inundation, echo=FALSE}
ggplot(alldata0, aes(dowy, inund_days, colour = Sradmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, inund_days, colour = WTmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, inund_days, colour = WTrangemwk, size = Q_sday)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, inund_days, colour = total_inund_last_year, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

```

# inundation other metrics that are not directly related to what is happening on the day of chl collection (lag variables)
```{r metrics, echo=FALSE}
ggplot(alldata0, aes(dowy, total_inund_last_year, colour = inundation, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, total_inund_last_year, colour = Sradmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, total_inund_last_year, colour = WTmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, total_inund_last_year, colour = WTrangemwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, total_inund_last_year, colour = Q_sday, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
#

ggplot(alldata0, aes(dowy, days_since_last_inundation, colour = inund_days, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, days_since_last_inundation, colour = Sradmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, days_since_last_inundation, colour = WTmwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, days_since_last_inundation, colour = WTrangemwk, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))

ggplot(alldata0, aes(dowy, days_since_last_inundation, colour = Q_sday, size = chlorophyll)) + 
  geom_point()+ 
  facet_grid(vars(region), vars(season))
# opposite of inundation and Qsday
```
