---
title: "explore_gam_data"
author: "Catarina Pien"
date: "1/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


    Here we read in cleaned dataset, plot data, look at outliers and correlations between variables. 

*Last updated 1/19/2022 by C. Pien

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(viridis)
library(psych) # for pairs
```


# Read in data
```{r}
covars <- read_csv("data_model/model_data_daily_covars.csv") %>%
  mutate(month = lubridate::month(date)) 

alldata <- read_csv("data_model/model_chla_covars_gam.csv")%>%
  mutate(month = lubridate::month(date),
         year = lubridate::year(date),
         doy = lubridate::yday(date)-60,
         water_year = ifelse(month>9, year + 1, year))

chla_nuts <- read_csv("data_model/model_chla_nuts_combined.csv")
```

## Temporarily use this until data are cleaned up above.
```{r}
inundation_doy1998 <- read_csv("data_model/inundation_with_doy1998.csv")

covars <- readRDS("bayes_models/mod_covariates_complete.rds") %>% 
  select(-yolo, -height_sac, -inund_days, -inun, -past_topped) %>%
  left_join(inundation_doy1998, by = c("doy1998", "date")) %>%
  mutate(month =  month(date))

chla_all <- readRDS("bayes_models/Chla_all.rds")%>%
    # only want chl
    select(station, date, chl) %>%
    filter(chl > 0 & station != '11455420') %>%
    filter(complete.cases(.)) %>% # keep only complete records
    arrange(station, date) %>%
    # add a diff time as above for join purposes
    mutate(doy1998 = as.numeric(difftime(date, as.Date("1998-01-01"), "day")) + 1,
           station_id = as.numeric(as.factor(station))) %>%
    mutate(log_chla = log(chl))

final_covars <- covars %>%
    mutate(Q_1day = lag(flow_usgs_verona, 1), #this one is not working
           Q_mwk = rollapply(flow_usgs_verona, 7, mean, align='right', partial=TRUE),
           T_mwk = rollapply(daymet_tmax, 7, mean, align='right', partial=TRUE),
           Srad_mwk = rollapply(daymet_srad, 7, mean, align='right', partial=TRUE)) %>%
    rename(Q_sday = flow_usgs_verona)

chla_uniq_date = chla_all %>% group_by(doy1998) %>% sample_n(1)

chla_covars <- left_join(chla_uniq_date, final_covars, by=c("doy1998", "date"))
```


# Look at outliers/patterns in data
**We should investigate and clean up any outliers**

## Chla
```{r}
chl_clean <- chla_nuts %>%
  filter(!is.na(chlorophyll)) %>%
  rename(chl = chlorophyll) %>%
  mutate(month = month(date),
         year = year(date),
         water_year = ifelse(month>9, year + 1, year))

chl_clean <- chl_all %>%
  filter(!is.na(chl))

ggplot(chl_clean) + geom_histogram(aes(chl))
ggplot(chl_clean) + geom_point(aes(date, chl, color = month), size = 3) + scale_color_viridis()
ggplot(chl_clean) + geom_boxplot(aes(x = factor(month), y = chl))
ggplot(chl_clean) + geom_boxplot(aes(x = factor(water_year), y = chl)) + theme(axis.text.x = element_text(angle = 90))
par(mfrow = c(1,1))
dotchart(chl_clean$chl, lcolor = "white", xlab = "Chla", ylab = "row")
```

## Covariates
```{r}
model_covars <- covars %>%
  select(c(date, month, water_year, daymet_precip_mm, Q_sday, flow_yolo, height_sac_na, inund_days, daymet_tmax, daymet_srad))

# make data long to plot everything faceted
data_long <- pivot_longer(model_covars, cols = 4:10, names_to = "variable", values_to = "value")

# all variables on one plot
ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5)

# point plot over time
ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5) + facet_wrap(variable~., scales = "free_y")

# boxplot by month
ggplot(data_long) + geom_boxplot(aes(x = factor(month), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw()

# boxplot by water year
ggplot(data_long) + geom_boxplot(aes(x = factor(water_year), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw() + theme(axis.text.x = element_text(angle = 90))
```

## Correlations between variables
```{r}
# precip - stage height
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, height_sac_na)) + geom_point()

# precip - inundation days
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, inund_days, color = month)) + geom_point() + scale_color_viridis()

# precip - flow_yolo
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, flow_yolo)) + geom_point()

# flow_verona - flow_yolo
ggplot(model_covars, aes(Q_sday, flow_yolo, color = month)) + geom_point() + scale_color_viridis()

# radiation - temp
ggplot(filter(model_covars, daymet_srad<1000), aes(daymet_tmax, daymet_srad, color = month)) + geom_point()

# temp - flow
ggplot(model_covars, aes(daymet_tmax, Q_sday)) + geom_point(aes(color = month))+ scale_color_viridis()

# Useful plot (seasonality of temp vs stage height)
ggplot(model_covars, aes(daymet_tmax, height_sac_na)) + geom_point(aes(color = month)) + scale_color_viridis() 
```

# Correlations with chlorophyll
```{r}
ggplot(subset(alldata, inund_days>0), aes(inund_days, chlorophyll, color = region_chla)) + geom_point() + theme_bw() + viridis::scale_color_viridis( discrete = TRUE, option = "turbo") + labs(x = "Inundation Days", color = "Region") + geom_smooth( aes(inund_days, chlorophyll))


stations <- alldata %>% select(station_wq_chl, latitude, longitude, region_chla) %>%
  distinct()

stations_plot <- ggplot(stations, aes(longitude, latitude,  color = station_wq_chl)) + geom_point()

plotly::ggplotly(stations_plot)
```


# Model exploration 

## Create a unique date dataset, and a dataset containing only days with chla data.
* sample, slice, mean, pick a station hierarchy?
```{r}
# uncomment this once we are using cleaned dataset

# data_unique_date = alldata %>% group_by(doy1998) %>% sample_n(1) %>%
#   mutate(log_chla = log(chl))
# 
# chla_covars = data_unique_date %>%
#   filter(!is.na(log_chla))
```

## See span and distribution of data 
Nice and normal!
```{r}
ggplot(chla_covars) + geom_point(aes(date, log_chla))
ggplot(chla_covars) + geom_histogram(aes(log_chla))
```

## Check for correlation between all vars
Strong correlations between all flow metrics. 
```{r}
# check for correlation between all predictors, and the response variable
x = na.omit(chla_covars) %>% select(log_chla, Q_sday, Q_1day, Q_mwk, T_mwk, Srad_mwk, inund_days, doy1998)

(corr_plot <- pairs.panels(x, hist.col = "white", cex.cor = 1.5, pch = 21, stars = TRUE))
```

### Save correlation plot
```{r}
 tiff(filename = "figures/corr_plot_gamvars.tiff", width = 10, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
corr_plot <- pairs.panels(x, hist.col = "white", cex.cor = 1, pch = 21, stars = TRUE)
 dev.off()
```

### add water year day
```{r}
devtools::install_github("ryanpeek/wateRshedTools")
library(wateRshedTools)

model_covars <- add_WYD(model_covars, "date")

chla_covars <- add_WYD(chla_covars, "date")
```

### add inundation year category
```{r}
model_covars$inundation <- ifelse(model_covars$inund_days > 0, 1, 0)

WY_summary <- model_covars %>% group_by(WY) %>% summarise(sum = sum(inundation))

WY_summary$IY <- ifelse(WY_summary$sum > 0, "Yes", "No") #drop 2021 and 2022 (not in all datasets)

model_covars <- merge(model_covars,  WY_summary, by = "WY")

model_covars <-subset(model_covars, WY < 2021)
```

### trying some more plots for Jan 24th
```{r}
tiff(filename = "figures/DOWY_Q_WY_inund.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(model_covars, aes(DOWY, Q_sday, colour = factor(WY), size = inund_days)) + geom_point() + facet_grid(cols = vars(IY))

 dev.off()
 
tiff(filename = "figures/Q_inund_tmax_precip.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(model_covars, aes(Q_sday, inund_days, colour = daymet_tmax, size = daymet_precip_mm)) + geom_point()

 dev.off()
 
tiff(filename = "figures/Q_chl_srad_inund.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(chla_covars, aes(Q_sday, chl, colour = daymet_srad, size = inund_days)) + geom_point() + ylim(0,30)

 dev.off()
```
