---
title: "explore_gam_data"
author: "Catarina Pien"
date: "1/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


    Here we read in cleaned dataset, plot data, look at outliers and correlations between variables. 

*Last updated 8/24/2022 by C. Pien


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(viridis)
library(psych) # for pairs
```


# Read in data
```{r}
#covars <- read_csv("data_model/model_data_daily_covars.csv") %>%
#  mutate(month = lubridate::month(date)) 

alldata <- read_csv("data_model/model_chla_covars_gam.csv")%>%
  mutate(month = lubridate::month(date),
         year = lubridate::year(date),
         doy = lubridate::yday(date)-60,
         water_year = ifelse(month>9, year + 1, year),
         log_chla = log(chlorophyll+1),
         log_qsdy = log(Q_sday)) %>%
  rename(station = station_wq_chl)

chla_nuts <- read_csv("data_model/model_covars_chla_nuts_extravars.csv") %>%
  filter(region!="cache") %>%
  select(-c(diss_ammonia_sign:diss_nitrate_nitrite, diss_orthophos_sign))%>%
  mutate(month = lubridate::month(date),
         year = lubridate::year(date),
         doy = lubridate::yday(date)-60,
         water_year = ifelse(month>9, year + 1, year),
         log_chla = log(chlorophyll+1),
         log_qsdy = log(Q_sday),
         inund_fac = factor(ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short"))),
         inund_fac2 = factor(ifelse(inund_days == 0, "none", ifelse(inund_days > 21, "long", "short")))) %>%
  rename(station = station_wq_chl)
```

## Full dataset
```{r}
data_long <- pivot_longer(chla_nuts, cols = c(din, diss_orthophos, diss_silica, inund_days,total_inund_last_year:precip_mm, log_qsdy), names_to = "Parameter", values_to = "Value")

ggplot(data_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

### Scale covars
```{r}
data_scaled <- chla_nuts %>%
  mutate(across(.cols = c(din, diss_orthophos, diss_silica, inund_days,total_inund_last_year:precip_mm), ~ scale(.x)))

data_s_long <- pivot_longer(data_scaled, cols = c(din, diss_orthophos, diss_silica, inund_days,total_inund_last_year:precip_mm), names_to = "Parameter", values_to = "Value")
```

### Explore data

#### See span and distribution of data 
Chlorophyll
```{r}
ggplot(alldata) + geom_point(aes(date, log_chla))
ggplot(alldata) + geom_histogram(aes(log_chla)) + facet_wrap(~region, scales="free")
```

Covars
```{r}
# point plot over time
ggplot(data_long) + geom_point(aes(x = date, y = Value, color = Parameter), alpha = 0.5) + facet_wrap(Parameter~., scales = "free_y")

# boxplot by month
(boxplot_month <- ggplot(data_long) + geom_boxplot(aes(x = factor(month), y = Value, fill = Parameter)) + facet_wrap(Parameter~., scales = "free_y") + theme_bw())

# boxplot by water year
(boxplot_year <- ggplot(data_long) + geom_boxplot(aes(x = factor(water_year), y = Value, fill = Parameter)) + facet_wrap(Parameter~., scales = "free_y") + theme_bw() + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position = "top"))
```

Relationships between transformed vars (flow, log_flow, chl, log_chl, also month)
```{r}
ggplot(chla_nuts) + geom_boxplot(aes(factor(month), log_qsdy )) + facet_wrap(~region) + theme_bw()
ggplot(chla_nuts) + geom_boxplot(aes(factor(month), Q_sday)) + facet_wrap(~region) + theme_bw()
ggplot(chla_nuts) + geom_boxplot(aes(factor(month), log_chla)) + facet_wrap(~region) + theme_bw()

ggplot(chla_nuts) + geom_jitter(aes(log_qsdy, Q_sday, color = region, shape = region), alpha = 0.5, size = 3) + theme_bw()
ggplot(chla_nuts) + geom_jitter(aes(log_chla, chlorophyll, color = region, shape = region), alpha = 0.5, size = 3) + theme_bw()
```

Save
```{r}
tiff(filename = "figures/boxplot_year_covars.tiff", width = 11, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
boxplot_year
 dev.off()
 
 
png(filename = "figures/boxplot_month_covars.png", width = 11, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
boxplot_month
 dev.off()
 
```



#### Chlorophyll Outliers
```{r}
chl_clean <- alldata %>%
  filter(!is.na(chlorophyll))

ggplot(chl_clean) + geom_histogram(aes(chlorophyll))
ggplot(chl_clean) + geom_point(aes(date, chlorophyll, color = month), size = 3) + scale_color_viridis()
ggplot(chl_clean) + geom_boxplot(aes(x = factor(month), y = chlorophyll))
ggplot(chl_clean) + geom_boxplot(aes(x = factor(water_year), y = chlorophyll)) + theme(axis.text.x = element_text(angle = 90))
par(mfrow = c(1,1))
dotchart(chl_clean$chlorophyll, lcolor = "white", xlab = "Chla", ylab = "row")
```
Possibly two outliers 

#### Relationship with chlorophyll
```{r}
ggplot(data_s_long) + geom_point(aes(x = Value, y = chlorophyll, color = factor(month))) +
  geom_smooth(aes(x = Value, y = chlorophyll), method = "gam") +
  facet_wrap(~Parameter, scales = "free") 
```

####  Correlation Plots

##### Nutrients
```{r}
nuts <- chla_nuts %>%
  select(c(date, month, water_year, region, Q_sday, WTmwk, inund_days, din, diss_orthophos, diss_silica, chlorophyll)) 

nuts_corr <- nuts %>%
  na.omit() %>%
  dplyr::select(Q_sday, WTmwk, inund_days, din, diss_orthophos, diss_silica, chlorophyll)
  
(corr_plot_nut <- pairs.panels(nuts_corr, hist.col = "white", cex.cor = 1.5, pch = 21, stars = TRUE))
```

##### Covariates
```{r}
covars <- chla_nuts %>%
  select(c(date, month, water_year, region, Q_sday, log_qsdy, WTmwk, WTrangemwk, Sradmwk, inund_days, days_since_last_inundation, total_inund_last_year, chlorophyll, diurnal_range, inund_fac2))

covars_corr <- covars %>%
  na.omit() %>%
  dplyr::select(Q_sday, log_qsdy, WTmwk, WTrangemwk, Sradmwk, diurnal_range, inund_days, days_since_last_inundation, total_inund_last_year, inund_fac2, chlorophyll) %>%
  mutate(log_chl = log(chlorophyll+1))

(corr_plot_covar <- pairs.panels(covars_corr, hist.col = "white", cex.cor = 1.5, pch = 21, stars = TRUE))
```

Save correlation plot
```{r}
png(filename = "figures/corr_plot_gamvars.png", width = 10, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
corr_plot_covar
 dev.off()
 
tiff(filename = "figures/corr_plot_nuts.tiff", width = 10, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
corr_plot_nut
 dev.off()
```


#### Regional Patterns

Look at nutrients
```{r}
data_s_long %>%
  filter(inundation == 0) %>%
  filter(Parameter %in% c("din", "diss_orthophos", "diss_silica")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

```{r}
data_long %>%
  filter(inundation == 1) %>%
  filter(Parameter %in% c("din", "diss_orthophos", "diss_silica")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

Temp, Srad
```{r}
data_s_long %>%
  filter(inundation == 0) %>%
  filter(Parameter %in% c("WTmwk", "Sradmwk")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, log_chl, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

```{r}
data_s_long %>%
  filter(inundation == 1) %>%
  filter(Parameter %in% c("WTmwk", "Sradmwk")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, log_chl, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

Flow, diurnal range, WTrange
* Interesting pattern in flow
```{r}
data_s_long %>%
  filter(inundation == 0) %>%
  filter(Parameter %in% c("Q_sday", "WTrangemwk", "diurnal_range")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

```{r}
data_s_long %>%
  filter(inundation == 1) %>%
  filter(Parameter %in% c("Q_sday", "WTrangemwk", "diurnal_range")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

Inundation metrics
```{r}
data_s_long %>%
  filter(inundation == 0) %>%
  filter(Parameter %in% c("total_inund_last_year", "inund_days", "days_since_last_inundation")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```

```{r}
data_s_long %>%
  filter(inundation == 1) %>%
  filter(Parameter %in% c("total_inund_last_year", "inund_days", "days_since_last_inundation")) %>%
  mutate(log_chl = log(chlorophyll+1)) %>%
  ggplot(aes(Value, chlorophyll, color = region)) + geom_point() + geom_smooth() +facet_wrap(~Parameter, scales = "free")
```




## Downstream only 
### Filter to downstream
```{r}
downstream <- alldata %>%
  filter(region== "below")
```

### Look at covariates
```{r}
downstream_long <- pivot_longer(downstream, cols = c(inund_days,total_inund_last_year, days_since_last_inundation, diurnal_range:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(downstream_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

#### These are very different scales. Scale covars.
```{r}
downstream_scaled <- downstream %>%
  mutate(across(.cols = c(inund_days,diurnal_range:Sradmwk), ~ scale(.x)))
```

#### Check distribution.
```{r}
downstream_s_long <- pivot_longer(downstream_scaled, cols = c(inund_days,total_inund_last_year, days_since_last_inundation, diurnal_range:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(downstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

### Explore data.

Relationship with chlorophyll
```{r}
ggplot(downstream_s_long) + geom_point(aes(x = Value, y = chlorophyll, color = factor(month))) +
  geom_smooth(aes(x = Value, y = chlorophyll), method = "gam") +
  facet_wrap(~Parameter, scales = "free") 
```

#### Break up inundation = 0/1
```{r}
downstream_inund <- downstream_s_long %>%
  filter(inundation == 1)
downstream_noinund <- downstream_s_long %>%
  filter(inundation == 0)

ggplot(downstream_inund) + geom_point(aes(x = Value, y = chlorophyll, color = factor(month))) +
  geom_smooth(aes(x = Value, y = chlorophyll), method = "gam") +
  facet_wrap(~Parameter, scales = "free") 

ggplot(downstream_noinund) + geom_point(aes(x = Value, y = chlorophyll, color = factor(month))) +
  geom_smooth(aes(x = Value, y = chlorophyll), method = "gam") +
  facet_wrap(~Parameter, scales = "free") 
```

#### By month
```{r}
ggplot(downstream_scaled) + geom_point(aes(x = inund_days, y = chlorophyll)) +
  geom_smooth(aes(x = inund_days, y = chlorophyll), method = "gam") + facet_wrap(~month, scales = "free") 
```



















## Temporarily use this until data are cleaned up above.
```{r}
inundation_doy1998 <- read_csv("data_model/inundation_with_doy1998.csv")

covars <- readRDS("bayes_models/mod_covariates_complete.rds") %>% 
  select(-yolo, -height_sac, -inund_days, -inun, -past_topped) %>%
  left_join(inundation_doy1998, by = c("doy1998", "date")) %>%
  mutate(month =  month(date))

chla_all <- readRDS("bayes_models/Chla_all.rds")%>%
    # only want chl
    select(station, date, chl) %>%
    filter(chl > 0 & station != '11455420') %>%
    filter(complete.cases(.)) %>% # keep only complete records
    arrange(station, date) %>%
    # add a diff time as above for join purposes
    mutate(doy1998 = as.numeric(difftime(date, as.Date("1998-01-01"), "day")) + 1,
           station_id = as.numeric(as.factor(station))) %>%
    mutate(log_chla = log(chl))

final_covars <- covars %>%
    mutate(Q_1day = lag(flow_usgs_verona, 1), #this one is not working
           Q_mwk = rollapply(flow_usgs_verona, 7, mean, align='right', partial=TRUE),
           T_mwk = rollapply(daymet_tmax, 7, mean, align='right', partial=TRUE),
           Srad_mwk = rollapply(daymet_srad, 7, mean, align='right', partial=TRUE)) %>%
    rename(Q_sday = flow_usgs_verona)

chla_uniq_date = chla_all %>% group_by(doy1998) %>% sample_n(1)

chla_covars <- left_join(chla_uniq_date, final_covars, by=c("doy1998", "date"))
```


# Look at outliers/patterns in data
**We should investigate and clean up any outliers**

## Chla
```{r}
chl_clean <- chla_nuts %>%
  filter(!is.na(chlorophyll)) %>%
  rename(chl = chlorophyll) %>%
  mutate(month = month(date),
         year = year(date),
         water_year = ifelse(month>9, year + 1, year))

chl_clean <- chl_all %>%
  filter(!is.na(chl))

ggplot(chl_clean) + geom_histogram(aes(chl))
ggplot(chl_clean) + geom_point(aes(date, chl, color = month), size = 3) + scale_color_viridis()
ggplot(chl_clean) + geom_boxplot(aes(x = factor(month), y = chl))
ggplot(chl_clean) + geom_boxplot(aes(x = factor(water_year), y = chl)) + theme(axis.text.x = element_text(angle = 90))
par(mfrow = c(1,1))
dotchart(chl_clean$chl, lcolor = "white", xlab = "Chla", ylab = "row")
```

## Covariates
```{r}
model_covars <- covars %>%
  select(c(date, month, water_year, daymet_precip_mm, Q_sday, flow_yolo, height_sac_na, inund_days, daymet_tmax, daymet_srad))

# make data long to plot everything faceted
data_long <- pivot_longer(model_covars, cols = 4:10, names_to = "variable", values_to = "value")

# all variables on one plot
ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5)

# point plot over time
ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5) + facet_wrap(variable~., scales = "free_y")

# boxplot by month
ggplot(data_long) + geom_boxplot(aes(x = factor(month), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw()

# boxplot by water year
ggplot(data_long) + geom_boxplot(aes(x = factor(water_year), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw() + theme(axis.text.x = element_text(angle = 90))
```

## Correlations between variables
```{r}
# precip - stage height
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, height_sac_na)) + geom_point()

# precip - inundation days
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, inund_days, color = month)) + geom_point() + scale_color_viridis()

# precip - flow_yolo
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, flow_yolo)) + geom_point()

# flow_verona - flow_yolo
ggplot(model_covars, aes(Q_sday, flow_yolo, color = month)) + geom_point() + scale_color_viridis()

# radiation - temp
ggplot(filter(model_covars, daymet_srad<1000), aes(daymet_tmax, daymet_srad, color = month)) + geom_point()

# temp - flow
ggplot(model_covars, aes(daymet_tmax, Q_sday)) + geom_point(aes(color = month))+ scale_color_viridis()

# Useful plot (seasonality of temp vs stage height)
ggplot(model_covars, aes(daymet_tmax, height_sac_na)) + geom_point(aes(color = month)) + scale_color_viridis() 
```

# Model exploration 

### Save correlation plot
```{r}
 tiff(filename = "figures/corr_plot_gamvars.tiff", width = 10, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
corr_plot <- pairs.panels(x, hist.col = "white", cex.cor = 1, pch = 21, stars = TRUE)
 dev.off()
```

### add water year day
```{r}
devtools::install_github("ryanpeek/wateRshedTools")
library(wateRshedTools)

model_covars <- add_WYD(model_covars, "date")

chla_covars <- add_WYD(chla_covars, "date")
```

### add inundation year category
```{r}
model_covars$inundation <- ifelse(model_covars$inund_days > 0, 1, 0)

WY_summary <- model_covars %>% group_by(WY) %>% summarise(sum = sum(inundation))

WY_summary$IY <- ifelse(WY_summary$sum > 0, "Yes", "No") #drop 2021 and 2022 (not in all datasets)

model_covars <- merge(model_covars,  WY_summary, by = "WY")

model_covars <-subset(model_covars, WY < 2021)
```

### trying some more plots for Jan 24th
```{r}
tiff(filename = "figures/DOWY_Q_WY_inund.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(model_covars, aes(DOWY, Q_sday, colour = factor(WY), size = inund_days)) + geom_point() + facet_grid(cols = vars(IY))

 dev.off()
 
tiff(filename = "figures/Q_inund_tmax_precip.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(model_covars, aes(Q_sday, inund_days, colour = daymet_tmax, size = daymet_precip_mm)) + geom_point()

 dev.off()
 
tiff(filename = "figures/Q_chl_srad_inund.tiff", width = 10, height = 8, units = "in", res = 300)

ggplot(chla_covars, aes(Q_sday, chl, colour = daymet_srad, size = inund_days)) + geom_point() + ylim(0,30)

 dev.off()
```
