---
title: "explore_gam_data"
author: "Catarina Pien"
date: "1/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


    Here we read in cleaned dataset, plot data, look at outliers and correlations between variables. 

*Last updated 1/14/2022 by C. Pien

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(viridis)
library(psych) # for pairs
```


# Read in data
```{r}
covars <- read_csv("data_model/gam_data_covars.csv") %>%
  mutate(month = lubridate::month(date)) 

alldata <- read_csv("data_model/gam_data_covars_chla.csv")%>%
  mutate(month = lubridate::month(date),
         year = lubridate::year(date)) 
```


# Look at outliers/patterns in data
**We should investigate and clean up any outliers**

## Chla
```{r}
chl_clean <- alldata %>%
  filter(!is.na(chl))

ggplot(chl_clean) + geom_histogram(aes(chl))
ggplot(chl_clean) + geom_point(aes(date, chl, color = month), size = 3) + scale_color_viridis()
ggplot(chl_clean) + geom_boxplot(aes(x = factor(month), y = chl))
ggplot(chl_clean) + geom_boxplot(aes(x = factor(water_year), y = chl)) + theme(axis.text.x = element_text(angle = 90))
dotchart(chla_all$chl, lcolor = "white", xlab = "Chla", ylab = "row")
```

```{r}
model_covars <- covars %>%
  select(c(date, month, water_year, daymet_precip_mm, Q_sday, flow_yolo, height_sac_na, inund_days, daymet_tmax, daymet_srad))

data_long <- pivot_longer(model_covars, cols = 4:10, names_to = "variable", values_to = "value")

ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5)

ggplot(data_long) + geom_point(aes(x = date, y = value, color = variable), alpha = 0.5) + facet_wrap(variable~., scales = "free_y")

ggplot(data_long) + geom_boxplot(aes(x = factor(month), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw()

ggplot(data_long) + geom_boxplot(aes(x = factor(water_year), y = value, fill = variable)) + facet_wrap(variable~., scales = "free_y") + theme_bw() + theme(axis.text.x = element_text(angle = 90))
```

## Correlations between variables
```{r}
# precip - stage height
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, height_sac_na)) + geom_point()

# precip - inundation days
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, inund_days, color = month)) + geom_point() + scale_color_viridis()

# precip - flow_yolo
ggplot(filter(model_covars, daymet_precip_mm>0), aes(daymet_precip_mm, flow_yolo)) + geom_point()

# flow_verona - flow_yolo
ggplot(model_covars, aes(Q_sday, flow_yolo, color = month)) + geom_point() + scale_color_viridis()

# radiation - temp
ggplot(filter(model_covars, daymet_srad<1000), aes(daymet_tmax, daymet_srad, color = month)) + geom_point()

# temp - flow
ggplot(model_covars, aes(daymet_tmax, Q_sday)) + geom_point(aes(color = month))+ scale_color_viridis()

# Useful plot (seasonality of temp vs stage height)
ggplot(model_covars, aes(daymet_tmax, height_sac_na)) + geom_point(aes(color = month)) 
```

# Model exploration 

## Create a unique date dataset, and a dataset containing only days with chla data.
* sample, slice, mean, pick a station hierarchy?
```{r}
data_unique_date = alldata %>% group_by(doy1998) %>% sample_n(1) %>%
  mutate(log_chla = log(chl))

chla_covars = data_unique_date %>%
  filter(!is.na(log_chla))
```

## See span and distribution of data 
Nice and normal!
```{r}
ggplot(chla_covars) + geom_point(aes(date, log_chla))
ggplot(chla_covars) + geom_histogram(aes(log_chla))
```

## Check for correlation between all vars
Strong correlations between all flow metrics. 
```{r}
# check for correlation between all predictors, and the response variable
x = na.omit(chla_covars) %>% select(log_chla, Q_sday, Q_1day, Q_mwk, T_mwk, Srad_mwk, inund_days, doy1998)

(corr_plot <- pairs.panels(x, hist.col = "white", cex.cor = 1.5, pch = 21, stars = TRUE))
```

### Save correlation plot
```{r}
 tiff(filename = "figures/corr_plot_gamvars.tiff", width = 10, height = 8, units = "in", res = 300)
 # plot with R2 values on the diagnolly opposite side
corr_plot
 dev.off()
```


```{r}

```

