---
title: "flow_constituent_viz"
author: "Elizabeth Stumpner"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)
library(ggbeeswarm)
```
### Borrowed from a script I now can't find
```{r}
alldata <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         fwater_year = factor(water_year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         t_qsdy = kader:::cuberoot(Q_sday), 
         log_qsdy = log(Q_sday),
         inund_fac = ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short")),
         inund_fac2 = ifelse(inund_days == 0, "none", ifelse(inund_days > 21, "long", "short")),
         inundation = as.factor(inundation),
         station_wq_chl = as.factor(station_wq_chl),
         tily_fac = case_when(total_inund_last_year == 0 ~ "none",
                              total_inund_last_year>0 & total_inund_last_year < 16 ~ "2wk",
                              total_inund_last_year>16 & total_inund_last_year < 41 ~ "month",
                              TRUE ~ "months"),
         tily_fac = as.factor(tily_fac),
         inund_fac = as.factor(inund_fac),
         inund_fac2 = factor(inund_fac2, levels = c("none", "short", "long"))) %>%
  rename(station = station_wq_chl, tily = total_inund_last_year, 
         dsli = days_since_last_inundation, idysun = days_of_inundation_until_now)
```
## Set 1

### First, does it matter whether I use mwk or mdy (mean of week and mean of day, right?)

### As suggested in group meeting, viewing WY temp by boxplot is not an apples-to-apple comparison - details like the timing and water temperature of inundation is glossed over with this view.

### Lastly, breaking down by region or inundation period is not especially useful.
```{r}
#incase use later
decade <- subset(alldata, water_year == c(2011:2021))

#compare two means
alldata %>% 
ggplot(aes(x=dowy, y=WTmwk, color = "mwk")) +
  geom_smooth() +
  geom_smooth(aes(dowy, WTmday, color = "mday"))
  

alldata %>% 
ggplot(aes(x=fwater_year, y=WTmday))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90)) 

alldata %>% 
ggplot(aes(x=fwater_year, y=WTmday))+
  geom_boxplot()+
  facet_grid(~region)+
  theme(axis.text.x = element_text(angle = 90)) 

alldata %>% 
ggplot(aes(x=fwater_year, y=WTmday))+
  geom_boxplot()+
  facet_grid(~inund_fac2)+
  theme(axis.text.x = element_text(angle = 90)) 
```

## Set 2 

### Explore temperature and chlorophyll and split into region. Plot A shows all water years in the last decade (probably should choose a different color scheme) and Plot B includes length of inundation - but not all shapes seem to be included. Plot C is all data over wtr_dy 50 - 225 and Plot D is limited to short and long periods of inundation - again i'm doing something wrong bc not all water years are assigned a shape. Plot E is another variation.
```{r}

ggplot(decade, aes(x=dowy, y=WTmwk, color = as.factor(water_year), size = chlorophyll))+
  #geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  geom_point()+
  facet_grid(region~., scales="free_y")+
  ggtitle('A')+
  #coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()

#looks like not all water years are assigned a shape?

ggplot(decade, aes(x=dowy, y=WTmwk, color = inund_fac, size = chlorophyll, shape = as.factor(water_year)))+
  #geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  geom_point()+
  facet_grid(region~., scales="fixed")+
  ggtitle('B')+
  #coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()

ggplot(alldata, aes(x=dowy, y=WTmwk, color = inund_fac, size = chlorophyll))+
  #geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  geom_point()+
  xlim(c(50, 225))+
  facet_grid(region~., scales="fixed")+
  ggtitle('C')+
  #coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()

#remove 'no inundation'
alldata_inun <- subset(alldata, inund_fac == c("short", "long"))

#looks like not all water years are assigned a shape?
ggplot(alldata_inun, aes(x=dowy, y=WTmwk, color = inund_fac, size = chlorophyll, shape = as.factor(water_year)))+
  #geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  geom_point()+
  facet_grid(region~., scales="fixed")+
  ggtitle('D')+
  #coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()

inund <- subset(alldata, inund_fac == c("short", "long"))
#change color scheme
ggplot(inund, aes(x=dowy, y=WTmday, color = fwater_year, shape = inund_fac))+
  #geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  geom_point()+
  facet_grid(region~., scales="fixed")+
  ggtitle('E')+
  #coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()

```

## Set 3 

### Beeswarms! Stopped here but can continue if anyone likes them. Probably need a better color/shape scheme. I seem to have trouble with assigning shapes.
```{r}
ggplot(alldata, aes(x=region, y=WTmday, color = chlorophyll))+
  facet_wrap(~inund_fac2, scales="fixed")+
  theme_bw() +
  geom_quasirandom(             
    size=2,                      
    alpha=0.75    
  )+ 

  theme_classic(base_size=20) +
  theme(text = element_text (size = 15), axis.text.x = element_text(angle = 90)) +
  scale_shape_manual(values = c(1, 3))

ggplot(alldata, aes(x=inund_fac, y=WTmday, color = chlorophyll))+
  facet_wrap(~region, scales="fixed")+
  theme_bw() +
  geom_quasirandom(             
    size=2,                      
    alpha=0.75    
  )+ 

  theme_classic(base_size=20) +
  theme(text = element_text (size = 15), axis.text.x = element_text(angle = 90)) +
  scale_shape_manual(values = c(1, 3))
```


