---
title: "make_gam_hypothesis"
author: "NCEAS"
date: '2022-08-11'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)

# tinytex::install_tinytex() - run in console before knitting
```

# Notes 
* Srad = WT and dowy (highest correlation with both) or choose most ecologically relevant

Test: log_chla~te(WTmwk, log_qsdy, by = "inundation")

* Plot residuals of model against other variables we think are important
* Or separate analysis for legacy effects
* Or don't analyze and mention in discussion
* break into factor-combos (inundation & tily_fac combos, one by statement)

To do:
* Run models with un-transformed data
* Predict upstream and yolo models
* Figure out autocorrelation and if it matters
* More visualizations

# Read in data
```{r}
alldata0 <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         t_qsdy = kader:::cuberoot(Q_sday), 
         log_qsdy = log(Q_sday),
         inund_fac = ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short")),
         inund_fac2 = ifelse(inund_days == 0, "none", ifelse(inund_days > 21, "long", "short")),
         inundation = as.factor(inundation),
         station_wq_chl = as.factor(station_wq_chl),
         tily_fac = case_when(total_inund_last_year == 0 ~ "none",
                              total_inund_last_year>0 & total_inund_last_year < 16 ~ "2wk",
                              total_inund_last_year>16 & total_inund_last_year < 41 ~ "month",
                              TRUE ~ "months"),
         tily_fac = as.factor(tily_fac),
         inund_fac = as.factor(inund_fac),
         inund_fac2 = as.factor(inund_fac2)) %>%
  rename(station = station_wq_chl, tily = total_inund_last_year, 
         dsli = days_since_last_inundation, idysun = days_of_inundation_until_now)

str(alldata0)
alldata0 %>% filter(region == "above") %>% summary()
```

```{r}
cor(alldata0$Sradmwk, alldata0$WTmwk, method = "spearman")
cor(alldata0$Sradmwk, alldata0$dowy, method = "spearman")
cor(alldata0$WTmwk, alldata0$dowy, method = "spearman")
```


Filter to inundation period
```{r}
inundPd <- alldata0 %>% filter(inundation == 1) 
inMin <- min(inundPd$dowy)
inMax <- max(inundPd$dowy)

alldata <- alldata0 %>%
  filter(dowy >= inMin & dowy <= inMax)
```

```{r}
dataplot <- alldata0 %>%
  mutate(date2 = paste0("2000-", month(date), "-", day(date)),
         date2 = ymd(date2))

ggplot(dataplot) + 
  geom_point(aes(x = date2, y = inundation, color = inund_fac2)) + 
  scale_x_date(date_breaks = "month", date_labels = "%b") + 
  scale_color_viridis(discrete = TRUE, option = "plasma") + 
  labs(x = "Date", color = "Inundation Duration") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 13),
        axis.text.y = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "top",
        axis.title.x = element_blank(),
        panel.grid.minor = element_blank())
```


## Filter datasets to region
```{r}
downstream <- alldata %>% filter(region == "below")
upstream <- alldata %>% filter(region == "above")
yolo <- alldata %>% filter(region == "yolo")
```
## Scale datasets
```{r}
d.idys <- downstream %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
u.idys <- upstream %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
yo.idys <- yolo %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
```

# Model

## Run Models
### Upstream
```{r}
gamu1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu1)

gamu2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu2)

gamu3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = u.idys, family = "gaussian")
summary(gamu3)

gamu4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu4)


gamu5 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")

######################
gamu6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")
summary(gamu6)

gamu6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = upstream, family = "gaussian")
summary(gamu6b)

######################

AIC(gamu1, gamu2, gamu3, gamu4, gamu6)


```

### Yolo
```{r}
gamyo1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo1) # best

gamyo2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo2)

gamyo3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo3)

gamyo4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo4)

gamyo5 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo5)

#######################
gamyo6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo6)

#######################
gamyo6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = yolo, family = "gaussian")
summary(gamyo6)
######################

AIC(gamyo1, gamyo2, gamyo3, gamyo4, gamyo5, gamyo6)
```

### Downstream
```{r}
gamd1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd1)

gamd2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd2)

gamd3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = d.idys, family = "gaussian")
summary(gamd3)

gamd4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd4)

#####################
gamd6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = d.idys, family = "gaussian")
summary(gamd6)

gamd6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = downstream, family = "gaussian")
summary(gamd6b)
#####################

AIC(gamd1, gamd2, gamd3, gamd4, gamd6)
```

## Validation

gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")

### gamd6
```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamd6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamd6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(d.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(d.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(d.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
```

### gamu6
* some minor autocorrelation at t-1
```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamu6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamu6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamu6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(u.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(u.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(u.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  pacf(resid)
```

### gamyo6
```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamyo6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamyo6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamyo6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(yo.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(yo.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(yo.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  pacf(resid)
```

# Predict/Visualize data

* chlorophyll vs flow 5% WT, 95% WT
* chlorophyll vs WT 5% flow, 95% flow

## Downstream
### PREDICT
```{r}
library(tidymv)

# Get values 
(d_table <- downstream %>%
  group_by(inund_fac2)%>%
 summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())


# Prediction datasets - Downstream (unscaled)
pred_gamd6_none <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[2,2], d_table[2,3], by = 0.08),
                                                    WTmwk = seq(d_table[2,4],d_table[2,5], by = 0.1),
                                                    inund_fac2 = "none"))
pred_gamd6_short <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[3,2], d_table[3,3], by = 0.05),
                                                    WTmwk = seq(d_table[3,4], d_table[3,5], by = 0.1),
                                                    inund_fac2 = "short"))
pred_gamd6_long <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[1,2], d_table[1,3], by = 0.05),
                                                    WTmwk = seq(d_table[1,4], d_table[1,5], by = 0.1),
                                                    inund_fac2 = "long"))
    
pred_downstream <- rbind(pred_gamd6_short, pred_gamd6_long, pred_gamd6_none)
pred_downstream$inund_fac2 <- factor(pred_downstream$inund_fac2, levels = c("none", "short", "long"))
```

###VISUALIZE
```{r}
png(filename = "figures/downstream_predictions.png", width = 10, height = 4, units = "in", res = 300)
(downstream_plots <- pred_downstream %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point(size = 0.5) +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  facet_wrap(~inund_fac2) +
  scale_color_viridis() +
  labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Downstream") +
  theme_bw() +
    theme(strip.text = element_text(size = 13)))
dev.off()
```

```{r}
# by flow and WT
pred_gamd6_none %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "no inundation downstream") +
  theme_bw()

down_none <- pred_gamd6_none %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "no inundation downstream") +
  theme_bw()

pred_gamd6_short %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "short inundation downstream") +
  theme_bw()

down_short <- pred_gamd6_short %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "short inundation downstream") +
  theme_bw()

pred_gamd6_long %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "long inundation downstream") +
  theme_bw()





down_long <- pred_gamd6_long %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "long inundation downstream") +
  theme_bw()

# Long: Flow is dominant factor
# Short: At high temps






# limited data by WT divisions
pred_gamd6_q %>%
  ggplot(aes(log_qsdy, fit)) +
  geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~inund_fac2,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()


# All the prediction data
pred_gamd6 %>%
ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  #geom_smooth_ci(WTmwk, size = 1.5) +
  geom_point() +
 # geom_errorbar(aes(ymax = fit + se.fit, ymin = fit-se.fit)) + 
  #scale_color_manual(values = c("red", "blue")) +
  scale_color_viridis() +
  facet_wrap(~inund_fac2,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()
```

## Yolo
### PREDICT
```{r}
# Get values 
(y_table <- yolo %>%
  group_by(inund_fac2)%>%
 summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())

# Prediction datasets - Yolo (unscaled)
pred_gamyo6_none <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[2,2], y_table[2,3], by = 0.05),
                                                    WTmwk = seq(y_table[2,4],y_table[2,5], by = 0.2),
                                                    inund_fac2 = "none"))
pred_gamyo6_short <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[3,2], y_table[3,3], by = 0.05),
                                                    WTmwk = seq(y_table[3,4], y_table[3,5], by = 0.2),
                                                    inund_fac2 = "short"))
pred_gamyo6_long <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[1,2], y_table[1,3], by = 0.05),
                                                    WTmwk = seq(y_table[1,4], y_table[1,5], by = 0.2),
                                                    inund_fac2 = "long"))

pred_yolo <- rbind(pred_gamyo6_short, pred_gamyo6_long, pred_gamyo6_none)
pred_yolo$inund_fac2 <- factor(pred_yolo$inund_fac2, levels = c("none", "short", "long"))
```

### VISUALIZE
```{r}
png(filename = "figures/yolo_predictions.png", width = 9, height = 4, units = "in", res = 300)
(yolo_plots <- pred_yolo %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point(size = 0.5) +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  facet_wrap(~inund_fac2) +
  scale_color_viridis() +
  labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Yolo") +
  theme_bw() +
    theme(strip.text = element_text(size = 13)))
dev.off()
```

## Upstream
### PREDICT
```{r}
# Get values 
(u_table <- upstream %>%
  group_by(inund_fac2)%>%
 summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())

# Prediction datasets - Upstream (unscaled)
pred_gamu6_none <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[2,2], u_table[2,3], by = 0.02),
                                                    WTmwk = seq(u_table[2,4],u_table[2,5], by = 0.2),
                                                    inund_fac2 = "none")) 
pred_gamu6_short <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[3,2], u_table[3,3], by = 0.02),
                                                    WTmwk = seq(u_table[3,4], u_table[3,5], by = 0.2),
                                                    inund_fac2 = "short"))
pred_gamu6_long <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[1,2], u_table[1,3], by = 0.005),
                                                    WTmwk = seq(u_table[1,4], u_table[1,5], by = 0.2),
                                                    inund_fac2 = "long"))

pred_upstream <- rbind(pred_gamu6_short, pred_gamu6_long, pred_gamu6_none)
pred_upstream$inund_fac2 <- factor(pred_upstream$inund_fac2, levels = c("none", "short", "long"))
```

### VISUALIZE
```{r}
png(filename = "figures/upstream_predictions.png", width = 9, height = 4, units = "in", res = 300)
(upstream_plots <- pred_upstream %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point(size = 0.5) +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  facet_wrap(~inund_fac2) +
  scale_color_viridis() +
 labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Upstream") +
  theme_bw() +
    theme(strip.text = element_text(size = 13)))
dev.off()
```
