---
title: "make_gam_hypothesis"
author: "NCEAS"
date: '2022-08-11'
output: html_document
editor_options: 
  chunk_output_type: console
---


* How to change scale - predict over the whole range of values for all regions
* How to mask out the values we don't have 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)

# tinytex::install_tinytex() - run in console before knitting
```

# Notes 
* Srad = WT and dowy (highest correlation with both) or choose most ecologically relevant

Test: log_chla~te(WTmwk, log_qsdy, by = "inundation")

* Plot residuals of model against other variables we think are important
* Or separate analysis for legacy effects
* Or don't analyze and mention in discussion
* break into factor-combos (inundation & tily_fac combos, one by statement)

To do:
* Run models with un-transformed data
* Predict upstream and yolo models
* Figure out autocorrelation and if it matters
* More visualizations

# Read in data
```{r}
alldata0 <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         fwater_year = factor(water_year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         t_qsdy = kader:::cuberoot(Q_sday), 
         log_qsdy = log(Q_sday),
         inund_fac = ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short")),
         inund_fac2 = ifelse(inund_days == 0, "none", ifelse(inund_days > 21, "long", "short")),
         inundation = as.factor(inundation),
         station_wq_chl = as.factor(station_wq_chl),
         tily_fac = case_when(total_inund_last_year == 0 ~ "none",
                              total_inund_last_year>0 & total_inund_last_year < 16 ~ "2wk",
                              total_inund_last_year>16 & total_inund_last_year < 41 ~ "month",
                              TRUE ~ "months"),
         tily_fac = as.factor(tily_fac),
         inund_fac = as.factor(inund_fac),
         inund_fac2 = factor(inund_fac2, levels = c("none", "short", "long"))) %>%
  rename(station = station_wq_chl, tily = total_inund_last_year, 
         dsli = days_since_last_inundation, idysun = days_of_inundation_until_now)

str(alldata0)
alldata0 %>% filter(region == "above") %>% summary()
```

```{r}
cor(alldata0$Sradmwk, alldata0$WTmwk, method = "spearman")
cor(alldata0$Sradmwk, alldata0$dowy, method = "spearman")
cor(alldata0$WTmwk, alldata0$dowy, method = "spearman")
```


Filter to inundation period
```{r}
inundPd <- alldata0 %>% filter(inundation == 1) 
inMin <- min(inundPd$dowy)
inMax <- max(inundPd$dowy)

alldata <- alldata0 %>%
  filter(dowy >= inMin & dowy <= inMax)
```

```{r}
dataplot <- alldata0 %>%
  mutate(date2 = paste0("2000-", month(date), "-", day(date)),
         date2 = ymd(date2))

ggplot(dataplot) + 
  geom_point(aes(x = date2, y = inundation, color = inund_fac2)) + 
  scale_x_date(date_breaks = "month", date_labels = "%b") + 
  scale_color_viridis(discrete = TRUE, option = "plasma") + 
  labs(x = "Date", color = "Inundation Duration") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 13),
        axis.text.y = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "top",
        axis.title.x = element_blank(),
        panel.grid.minor = element_blank())
```


## Filter datasets to region
```{r}
downstream <- alldata %>% filter(region == "below")
upstream <- alldata %>% filter(region == "above")
yolo <- alldata %>% filter(region == "yolo")
```


## Scale datasets
```{r}
d.idys <- downstream %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
u.idys <- upstream %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
yo.idys <- yolo %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))


```

# Model

## Run Models

### Upstream

Still being run on scaled data, not original data
To decide whether to use WTmwk, Sradmwk or dowy for the time/seasonal component
To decide whether to use inundation days or inundation flag or inundation factor
1. inundation days (log_idys) is duration of inundation up to the day of observation
2. inundationn is 0 if not inundated; 1 if inundated
3. inund_fac  is "none" if no inundation; "short" if inundation < 2wk; "long" if inundation > 2wk
4. inund_fac2 is "none" if no inundation; "short" if inundation < 3wk; "long" if inundation > 3wk

```{r}
gamu1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu1)

gamu2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu2)

gamu3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = u.idys, family = "gaussian")
summary(gamu3)

gamu4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = u.idys, family = "gaussian")
summary(gamu4)

gamu5 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")

######################
gamu6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")
summary(gamu6)

# the final model run on non-scaled data (best model is actually gamu4 for upstream)
# final chosen covars are log_qsdy, WTmwk and inund_fac2
gamu6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = upstream, family = "gaussian")
summary(gamu6b)

# final model with random effect for station
gamu6c <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2)  + s(station, bs = "re"), method = "REML", data = upstream, family = "gaussian")
summary(gamu6c)
gam.vcomp(gamu6c)

gamu6d <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) +inund_fac2 + s(station, bs = "re"), method = "REML", data = upstream, family = "gaussian")
summary(gamu6d)
gam.vcomp(gamu6d)

# add year as random effect
gamu6e <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + s(station, bs = "re", k = 3) + s(water_year, bs = "re", k = 3), method = "REML", data = upstream, family = "gaussian")
summary(gamu6e)
gam.vcomp(gamu6e) # what does the stddev tell us?

AIC(gamu6c, gamu6d, gamu6e)
######################

AIC(gamu1, gamu2, gamu3, gamu4, gamu6, gamu6b, gamu6c)

vis.gam(gamu6c, view = c("log_qsdy", "WTmwk"), color = "topo", too.far = 0.1, type = "response")
```
* https://gavinsimpson.github.io/gratia/
* extract data from vis.gam?

### Yolo
```{r}
gamyo1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo1) # best

gamyo2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo2)

gamyo3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo3)

gamyo4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo4)

gamyo5 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo5)

#######################
gamyo6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = yo.idys, family = "gaussian")
summary(gamyo6)

# the final model  run on non-scaled data (best model is this one)
# final chosen covars are log_qsdy, WTmwk and inund_fac2
gamyo6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = yolo, family = "gaussian")
summary(gamyo6b)

# final model with random effect for station
gamyo6c <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + s(station, bs = "re"), method = "REML", data = yolo, family = "gaussian")
summary(gamyo6c)
gam.vcomp(gamyo6c)

# final model with random effect for station
gamyo6d <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + inund_fac2 + s(station, bs = "re"), method = "REML", data = yolo, family = "gaussian")
summary(gamyo6d)

######################

AIC(gamyo1, gamyo2, gamyo3, gamyo4, gamyo5, gamyo6, gamyo6b, gamyo6c)
```

### Downstream
```{r}
gamd1 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd1)

gamd2 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd2)

gamd3 <- gam(log_chla ~ te(log_qsdy, WTmwk, log_idys), method = "REML", data = d.idys, family = "gaussian")
summary(gamd3)

gamd4 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
summary(gamd4)

#####################
gamd6 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = d.idys, family = "gaussian")
summary(gamd6)

# the final model  run on non-scaled data (best model is this one)
# final chosen covars are log_qsdy, WTmwk and inund_fac2
gamd6b <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = downstream, family = "gaussian")
summary(gamd6b)

# with station random effect
gamd6c <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + inund_fac2 + s(station, bs = "re"), method = "REML", data = downstream, family = "gaussian")
summary(gamd6c)
gam.vcomp(gamd6c)
#####################

AIC(gamd1, gamd2, gamd3, gamd4, gamd6, gamd6b, gamd6c)
```

## Validation

gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2), method = "REML", data = u.idys, family = "gaussian")

### gamd6 - downstream model validation

```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamd6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamd6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(d.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(d.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(d.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
```

### gamu6 - upstream model validation

* some minor autocorrelation at t-1

```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamu6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamu6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamu6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(u.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(u.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(u.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  pacf(resid)
```

### gamyo6 - Yolo model validation

```{r}
  # validate the final model by plotting residuals and save to a tiff
  png(filename = "figures/gamyo6.tiff", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamyo6)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamyo6), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(yo.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(yo.idys$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(yo.idys$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  pacf(resid)
```


# Predict/Visualize data

* chlorophyll vs flow 5% WT, 95% WT
* chlorophyll vs WT 5% flow, 95% flow

## Downstream

### PREDICT

```{r}

library(tidymv)

# Get values 
(d_table <- downstream %>%
  group_by(inund_fac2)%>%
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())

# divide all covars into this many intervals in simulated dataset
grid.size = 49
# calculate interval size for all 3 inundation factor simulation datasets
int.none.lqsdy = (d_table[2,3] - d_table[2,2])/(grid.size)
int.none.wtmwk = (d_table[2,5] - d_table[2,4])/(grid.size)
int.shrt.lqsdy = (d_table[3,3] - d_table[3,2])/(grid.size)
int.shrt.wtmwk = (d_table[3,5] - d_table[3,4])/(grid.size)
int.long.lqsdy = (d_table[1,3] - d_table[1,2])/(grid.size)
int.long.wtmwk = (d_table[1,5] - d_table[1,4])/(grid.size)

seq.none.lqsdy = seq(d_table[2,2], d_table[2,3], by = int.none.lqsdy)
seq.none.wtmwk = seq(d_table[2,4], d_table[2,5], by = int.none.wtmwk)
seq.shrt.lqsdy = seq(d_table[3,2], d_table[3,3], by = int.shrt.lqsdy)
seq.shrt.wtmwk = seq(d_table[3,4], d_table[3,5], by = int.shrt.wtmwk)
seq.long.lqsdy = seq(d_table[1,2], d_table[1,3], by = int.long.lqsdy)
seq.long.wtmwk = seq(d_table[1,4], d_table[1,5], by = int.long.wtmwk)

# Prediction datasets - Downstream (unscaled)
pred_gamd6_none  <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[2,2], d_table[2,3], by = int.none.lqsdy),
                                                         WTmwk = seq(d_table[2,4], d_table[2,5], by = int.none.wtmwk),
                                                         inund_fac2 = "none"))
pred_gamd6_short <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[3,2], d_table[3,3], by = int.shrt.lqsdy),
                                                         WTmwk = seq(d_table[3,4], d_table[3,5], by = int.shrt.wtmwk),
                                                         inund_fac2 = "short"))
pred_gamd6_long  <- predict_gam(gamd6b, values = list(log_qsdy = seq(d_table[1,2], d_table[1,3], by = int.long.lqsdy),
                                                         WTmwk = seq(d_table[1,4], d_table[1,5], by = int.long.wtmwk),
                                                         inund_fac2 = "long"))

# combine all three inundation factor predictions into one data frame    
pred_downstream <- rbind(pred_gamd6_short, pred_gamd6_long, pred_gamd6_none)
# set the order of factors for inund_fac2
pred_downstream$inund_fac2 <- factor(pred_downstream$inund_fac2, levels = c("none", "short", "long"))

```

### VISUALIZE

```{r}
# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)


for (i in 1:(grid.size+1)) {
  for (j in 1:(grid.size+1)) {
    pred.none = predict_gam(gamd6b, values = list(log_qsdy = seq.none.lqsdy[i], WTmwk = seq.none.wtmwk[j], inund_fac2 = "none"))
    pred.shrt = predict_gam(gamd6b, values = list(log_qsdy = seq.shrt.lqsdy[i], WTmwk = seq.shrt.wtmwk[j], inund_fac2 = "short"))
    pred.long = predict_gam(gamd6b, values = list(log_qsdy = seq.long.lqsdy[i], WTmwk = seq.long.wtmwk[j], inund_fac2 = "long"))
    chl.mat.none[i, j] = pred.none$fit
    chl.mat.shrt[i, j] = pred.shrt$fit
    chl.mat.long[i, j] = pred.long$fit
  }
}
# Color palette
col_range = colorRampPalette(viridis(12, option = "turbo"))
col_pal = col_range(10000)

# this is a way to make a single palette across all 3 panels
# image(x = seq.none.lqsdy, y = seq.none.wtmwk, z = chl.mat.none,
#       col = col_pal[round(min(chl.mat.none)*length(col_pal)):round(max(chl.mat.none)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.none.wtmwk), xlim = range(seq.none.lqsdy))
# image(x = seq.shrt.lqsdy, y = seq.shrt.wtmwk, z = chl.mat.shrt,
#       col = col_pal[round(min(chl.mat.shrt)*length(col_pal)):round(max(chl.mat.shrt)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.shrt.wtmwk), xlim = range(seq.shrt.lqsdy))
# image(x = seq.long.lqsdy, y = seq.long.wtmwk, z = chl.mat.long,
#       col = col_pal[round(min(chl.mat.long)*length(col_pal)):round(max(chl.mat.long)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.long.wtmwk), xlim = range(seq.long.lqsdy))

# calculating axes labels
lbl.none.lqsdy = round(seq(d_table[2,2], d_table[2,3], by = (d_table[2,3] - d_table[2,2])/5), digits = 2)
lbl.none.wtmwk = round(seq(d_table[2,4], d_table[2,5], by = (d_table[2,5] - d_table[2,4])/5), digits = 2)
lbl.shrt.lqsdy = round(seq(d_table[3,2], d_table[3,3], by = (d_table[3,3] - d_table[3,2])/5), digits = 2)
lbl.shrt.wtmwk = round(seq(d_table[3,4], d_table[3,5], by = (d_table[3,5] - d_table[3,4])/5), digits = 2)
lbl.long.lqsdy = round(seq(d_table[1,2], d_table[1,3], by = (d_table[1,3] - d_table[1,2])/5), digits = 2)
lbl.long.wtmwk = round(seq(d_table[1,4], d_table[1,5], by = (d_table[1,5] - d_table[1,4])/5), digits = 2)

# subset the downstream data frame each inundation factor
d.none = downstream %>% filter(inund_fac2 == "none")
d.shrt = downstream %>% filter(inund_fac2 == "short")
d.long = downstream %>% filter(inund_fac2 == "long")

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(d.none$log_qsdy, d.none$WTmwk), chull(d.none$log_qsdy, d.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(d.shrt$log_qsdy, d.shrt$WTmwk), chull(d.shrt$log_qsdy, d.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(d.long$log_qsdy, d.long$WTmwk), chull(d.long$log_qsdy, d.long$WTmwk)[1])

# subset to only the hull rows - constraining to 5% and 95% intervals
d.none.hull = d.none[cnvx.hull.none,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[2,2], d_table[2,2], ifelse(log_qsdy > d_table[2,3], d_table[2,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < d_table[2,4], d_table[2,4], ifelse(WTmwk > d_table[2,5], d_table[2,5], WTmwk)))
d.shrt.hull = d.none[cnvx.hull.shrt,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[3,2], d_table[3,2], ifelse(log_qsdy > d_table[3,3], d_table[3,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < d_table[3,4], d_table[3,4], ifelse(WTmwk > d_table[3,5], d_table[3,5], WTmwk)))
d.long.hull = d.none[cnvx.hull.long,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[1,2], d_table[1,2], ifelse(log_qsdy > d_table[1,3], d_table[1,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < d_table[1,4], d_table[1,4], ifelse(WTmwk > d_table[1,5], d_table[1,5], WTmwk)))

# this way maximizes contrast within each panel hence the scale within category is stretched
# par(mfrow = c(1,3), mar = c(4,2.75,1,2.5) + 0.1, oma = c(0,2.25,2,8), las = 1)

###-------------- Downstream: No inundation --------------###
png(filename = "figures/d_none_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.none.lqsdy, y = seq.none.wtmwk, z = chl.mat.none, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.none.wtmwk), xlim = range(seq.none.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.none.lqsdy, labels = lbl.none.lqsdy)
axis(side = 2, at = lbl.none.wtmwk, labels = lbl.none.wtmwk)

# draw the simple polygon where data exists
#polygon(x = d.none.hull$log_qsdy, y = d.none.hull$WTmwk, col = gray(1, 0.2), border = NA)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Downstream: No inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()

###--------- Downstream: short inundation (< 3 weeks) ---------###
png(filename = "figures/d_shrt_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.shrt.lqsdy, y = seq.shrt.wtmwk, z = chl.mat.shrt, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.shrt.wtmwk), xlim = range(seq.shrt.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.shrt.lqsdy, labels = lbl.shrt.lqsdy)
axis(side = 2, at = lbl.shrt.wtmwk, labels = lbl.shrt.wtmwk)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Downstream: short inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()

# draw the simple polygon where data exists
polygon(x = d.shrt.hull$log_qsdy, y = d.shrt.hull$WTmwk, col = gray(1, 0.2), border = NA)

###--------- Downstream: long inundation (> 3 weeks) ---------###
png(filename = "figures/d_long_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.long.lqsdy, y = seq.long.wtmwk, z = chl.mat.long, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.long.wtmwk), xlim = range(seq.long.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.long.lqsdy, labels = lbl.long.lqsdy)
axis(side = 2, at = lbl.long.wtmwk, labels = lbl.long.wtmwk)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Downstream: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()




# (downstream_plots <- pred_downstream %>%
#   ggplot(aes(WTmwk, fit, color = log_qsdy)) +
#   geom_point(size = 0.5) +
#   geom_smooth() + 
#   #geom_smooth_ci(WTmwk, size = 1.5) +
#   facet_wrap(~inund_fac2) +
#   scale_color_viridis() +
#   labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Downstream") +
#   theme_bw() +
#     theme(strip.text = element_text(size = 13)))

dev.off()

```



```{r}
# by flow and WT
pred_gamd6_none %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "no inundation downstream") +
  theme_bw()

down_none <- pred_gamd6_none %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "no inundation downstream") +
  theme_bw()

pred_gamd6_short %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "short inundation downstream") +
  theme_bw()

down_short <- pred_gamd6_short %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "short inundation downstream") +
  theme_bw()

pred_gamd6_long %>%
  ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "long inundation downstream") +
  theme_bw()


down_long <- pred_gamd6_long %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point() +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis() +
  labs(y = "chlorophyll", title = "long inundation downstream") +
  theme_bw()

# Long: Flow is dominant factor
# Short: At high temps


# limited data by WT divisions
pred_gamd6_q %>%
  ggplot(aes(log_qsdy, fit)) +
  geom_smooth_ci(WTmwk, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~inund_fac2,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()


# All the prediction data
pred_gamd6 %>%
ggplot(aes(log_qsdy, fit, color = WTmwk)) +
  #geom_smooth_ci(WTmwk, size = 1.5) +
  geom_point() +
 # geom_errorbar(aes(ymax = fit + se.fit, ymin = fit-se.fit)) + 
  #scale_color_manual(values = c("red", "blue")) +
  scale_color_viridis() +
  facet_wrap(~inund_fac2,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()
```

## Yolo
### PREDICT
```{r}
# Get values 
(y_table <- yolo %>%
  group_by(inund_fac2)%>%
 summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())

# Prediction datasets - Yolo (unscaled)
pred_gamyo6_none <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[2,2], y_table[2,3], by = 0.05),
                                                    WTmwk = seq(y_table[2,4],y_table[2,5], by = 0.2),
                                                    inund_fac2 = "none"))
pred_gamyo6_short <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[3,2], y_table[3,3], by = 0.05),
                                                    WTmwk = seq(y_table[3,4], y_table[3,5], by = 0.2),
                                                    inund_fac2 = "short"))
pred_gamyo6_long <- predict_gam(gamyo6b, values = list(log_qsdy = seq(y_table[1,2], y_table[1,3], by = 0.05),
                                                    WTmwk = seq(y_table[1,4], y_table[1,5], by = 0.2),
                                                    inund_fac2 = "long"))

pred_yolo <- rbind(pred_gamyo6_short, pred_gamyo6_long, pred_gamyo6_none)
pred_yolo$inund_fac2 <- factor(pred_yolo$inund_fac2, levels = c("none", "short", "long"))
```

```{r}
# divide all covars into this many intervals in simulated dataset
grid.size = 49
# calculate interval size for all 3 inundation factor simulation datasets
int.none.lqsdy = (y_table[2,3] - y_table[2,2])/(grid.size)
int.none.wtmwk = (y_table[2,5] - y_table[2,4])/(grid.size)
int.shrt.lqsdy = (y_table[3,3] - y_table[3,2])/(grid.size)
int.shrt.wtmwk = (y_table[3,5] - y_table[3,4])/(grid.size)
int.long.lqsdy = (y_table[1,3] - y_table[1,2])/(grid.size)
int.long.wtmwk = (y_table[1,5] - y_table[1,4])/(grid.size)

seq.none.lqsdy = seq(y_table[2,2], y_table[2,3], by = int.none.lqsdy)
seq.none.wtmwk = seq(y_table[2,4], y_table[2,5], by = int.none.wtmwk)
seq.shrt.lqsdy = seq(y_table[3,2], y_table[3,3], by = int.shrt.lqsdy)
seq.shrt.wtmwk = seq(y_table[3,4], y_table[3,5], by = int.shrt.wtmwk)
seq.long.lqsdy = seq(y_table[1,2], y_table[1,3], by = int.long.lqsdy)
seq.long.wtmwk = seq(y_table[1,4], y_table[1,5], by = int.long.wtmwk)
```

### VISUALIZE
```{r}
png(filename = "figures/yolo_predictions.png", width = 9, height = 4, units = "in", res = 300)
(yolo_plots <- pred_yolo %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point(size = 0.5) +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  facet_wrap(~inund_fac2) +
  scale_color_viridis() +
  labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Yolo") +
  theme_bw() +
    theme(strip.text = element_text(size = 13)))
dev.off()
```


```{r}
# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)


for (i in 1:(grid.size+1)) {
  for (j in 1:(grid.size+1)) {
    pred.none = predict_gam(gamyo6b, values = list(log_qsdy = seq.none.lqsdy[i], WTmwk = seq.none.wtmwk[j], inund_fac2 = "none"))
    pred.shrt = predict_gam(gamyo6b, values = list(log_qsdy = seq.shrt.lqsdy[i], WTmwk = seq.shrt.wtmwk[j], inund_fac2 = "short"))
    pred.long = predict_gam(gamyo6b, values = list(log_qsdy = seq.long.lqsdy[i], WTmwk = seq.long.wtmwk[j], inund_fac2 = "long"))
    chl.mat.none[i, j] = pred.none$fit
    chl.mat.shrt[i, j] = pred.shrt$fit
    chl.mat.long[i, j] = pred.long$fit
  }
}
# Color palette
col_range = colorRampPalette(viridis(12, option = "turbo"))
col_pal = col_range(10000)

# this is a way to make a single palette across all 3 panels
# image(x = seq.none.lqsdy, y = seq.none.wtmwk, z = chl.mat.none,
#       col = col_pal[round(min(chl.mat.none)*length(col_pal)):round(max(chl.mat.none)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.none.wtmwk), xlim = range(seq.none.lqsdy))
# image(x = seq.shrt.lqsdy, y = seq.shrt.wtmwk, z = chl.mat.shrt,
#       col = col_pal[round(min(chl.mat.shrt)*length(col_pal)):round(max(chl.mat.shrt)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.shrt.wtmwk), xlim = range(seq.shrt.lqsdy))
# image(x = seq.long.lqsdy, y = seq.long.wtmwk, z = chl.mat.long,
#       col = col_pal[round(min(chl.mat.long)*length(col_pal)):round(max(chl.mat.long)*length(col_pal))],
#       ylab = "", xlab = "", axes = F, ylim = range(seq.long.wtmwk), xlim = range(seq.long.lqsdy))

# calculating axes labels
lbl.none.lqsdy = round(seq(y_table[2,2], y_table[2,3], by = (y_table[2,3] - y_table[2,2])/5), digits = 2)
lbl.none.wtmwk = round(seq(y_table[2,4], y_table[2,5], by = (y_table[2,5] - y_table[2,4])/5), digits = 2)
lbl.shrt.lqsdy = round(seq(y_table[3,2], y_table[3,3], by = (y_table[3,3] - y_table[3,2])/5), digits = 2)
lbl.shrt.wtmwk = round(seq(y_table[3,4], y_table[3,5], by = (y_table[3,5] - y_table[3,4])/5), digits = 2)
lbl.long.lqsdy = round(seq(y_table[1,2], y_table[1,3], by = (y_table[1,3] - y_table[1,2])/5), digits = 2)
lbl.long.wtmwk = round(seq(y_table[1,4], y_table[1,5], by = (y_table[1,5] - y_table[1,4])/5), digits = 2)

# subset the yolo data frame each inundation factor
y.none = yolo %>% filter(inund_fac2 == "none")
y.shrt = yolo %>% filter(inund_fac2 == "short")
y.long = yolo %>% filter(inund_fac2 == "long")

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(y.none$log_qsdy, y.none$WTmwk), chull(y.none$log_qsdy, y.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(y.shrt$log_qsdy, y.shrt$WTmwk), chull(y.shrt$log_qsdy, y.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(y.long$log_qsdy, y.long$WTmwk), chull(y.long$log_qsdy, y.long$WTmwk)[1])

# subset to only the hull rows - constraining to 5% and 95% intervals
y.none.hull = y.none[cnvx.hull.none,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[2,2], y_table[2,2], ifelse(log_qsdy > y_table[2,3], y_table[2,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < y_table[2,4], y_table[2,4], ifelse(WTmwk > y_table[2,5], y_table[2,5], WTmwk)))
y.shrt.hull = y.none[cnvx.hull.shrt,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[3,2], y_table[3,2], ifelse(log_qsdy > y_table[3,3], y_table[3,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < y_table[3,4], y_table[3,4], ifelse(WTmwk > y_table[3,5], y_table[3,5], WTmwk)))
y.long.hull = y.none[cnvx.hull.long,] %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1,2], y_table[1,2], ifelse(log_qsdy > y_table[1,3], y_table[1,3], log_qsdy))) %>%
  mutate(WTmwk = ifelse(WTmwk < y_table[1,4], y_table[1,4], ifelse(WTmwk > y_table[1,5], y_table[1,5], WTmwk)))

# this way maximizes contrast within each panel hence the scale within category is stretched
# par(mfrow = c(1,3), mar = c(4,2.75,1,2.5) + 0.1, oma = c(0,2.25,2,8), las = 1)

###-------------- Yolo: No inundation --------------###
png(filename = "figures/d_none_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.none.lqsdy, y = seq.none.wtmwk, z = chl.mat.none, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.none.wtmwk), xlim = range(seq.none.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.none.lqsdy, labels = lbl.none.lqsdy)
axis(side = 2, at = lbl.none.wtmwk, labels = lbl.none.wtmwk)

# draw the simple polygon where data exists
#polygon(x = d.none.hull$log_qsdy, y = d.none.hull$WTmwk, col = gray(1, 0.2), border = NA)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: No inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()

###--------- Yolo: short inundation (< 3 weeks) ---------###
png(filename = "figures/d_shrt_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.shrt.lqsdy, y = seq.shrt.wtmwk, z = chl.mat.shrt, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.shrt.wtmwk), xlim = range(seq.shrt.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.shrt.lqsdy, labels = lbl.shrt.lqsdy)
axis(side = 2, at = lbl.shrt.wtmwk, labels = lbl.shrt.wtmwk)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: short inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()

# draw the simple polygon where data exists
polygon(x = d.shrt.hull$log_qsdy, y = d.shrt.hull$WTmwk, col = gray(1, 0.2), border = NA)

###--------- Yolo: long inundation (> 3 weeks) ---------###
png(filename = "figures/d_long_2dfigure.png", width = 8, height = 6, units = "in", res = 300)

image(x = seq.long.lqsdy, y = seq.long.wtmwk, z = chl.mat.long, col = col_pal,
      ylab = "", xlab = "", axes = F, ylim = range(seq.long.wtmwk), xlim = range(seq.long.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.long.lqsdy, labels = lbl.long.lqsdy)
axis(side = 2, at = lbl.long.wtmwk, labels = lbl.long.wtmwk)

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)
dev.off()

```


## Upstream
### PREDICT
```{r}
# Get values 
(u_table <- upstream %>%
  group_by(inund_fac2)%>%
 summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  ungroup() %>%
  as.data.frame())

# Prediction datasets - Upstream (unscaled)
pred_gamu6_none <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[2,2], u_table[2,3], by = 0.02),
                                                    WTmwk = seq(u_table[2,4],u_table[2,5], by = 0.2),
                                                    inund_fac2 = "none")) 
pred_gamu6_short <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[3,2], u_table[3,3], by = 0.02),
                                                    WTmwk = seq(u_table[3,4], u_table[3,5], by = 0.2),
                                                    inund_fac2 = "short"))
pred_gamu6_long <- predict_gam(gamu6b, values = list(log_qsdy = seq(u_table[1,2], u_table[1,3], by = 0.005),
                                                    WTmwk = seq(u_table[1,4], u_table[1,5], by = 0.2),
                                                    inund_fac2 = "long"))

pred_upstream <- rbind(pred_gamu6_short, pred_gamu6_long, pred_gamu6_none)
pred_upstream$inund_fac2 <- factor(pred_upstream$inund_fac2, levels = c("none", "short", "long"))
```

### VISUALIZE
```{r}
png(filename = "figures/upstream_predictions.png", width = 9, height = 4, units = "in", res = 300)
(upstream_plots <- pred_upstream %>%
  ggplot(aes(WTmwk, fit, color = log_qsdy)) +
  geom_point(size = 0.5) +
  geom_smooth() + 
  #geom_smooth_ci(WTmwk, size = 1.5) +
  facet_wrap(~inund_fac2) +
  scale_color_viridis() +
 labs(y = "log chlorophyll", x = "Mean Weekly Water Temperature (°C)", title = "Upstream") +
  theme_bw() +
    theme(strip.text = element_text(size = 13)))
dev.off()
```
