---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---
---
title: "make_explore_gam_structure"
author: "NCEAS Connectivity Group"
date: "4/28/2022"
output: html_document
editor_options: 
  chunk_output_type: console



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Connectivity Modeling - Chlorophyll a

## Preparation

### Call packages
```{r}
detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)

# tinytex::install_tinytex() - run in console before knitting
```

### Read in data

```{r}

alldata <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         log_qsdy = log(Q_sday)) %>%
  rename(station  = station_wq_chl) %>% 
  filter(region  != "cache")

```

Filter to downstream
```{r}
downstream <- alldata %>% filter(region == "below")
downstream$inundation = as.factor(downstream$inundation)
downstream$station = as.factor(downstream$station)

```

Look at covariates
```{r}
downstream_long <- pivot_longer(downstream, cols = c(diurnal_range:Sradmwk, log_idys, log_qsdy), names_to = "Parameter", values_to = "Value")

ggplot(downstream_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")

```

These are very different scales. Scale covars.
```{r}
downstream_scaled <- downstream %>%
  mutate(across(.cols = c(log_idys,log_qsdy,diurnal_range:Sradmwk), ~ scale(.x)))
```

Check distribution.
```{r}
downstream_s_long <- pivot_longer(downstream_scaled, cols = c(log_idys,log_qsdy,diurnal_range:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(downstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```


#### Check for autocorrelation in model, predictor, response
```{r}

x <- downstream_scaled

op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)

acf(x$log_chla)
acf(x$Q_sday)
acf(x$WTmwk)
acf(x$Sradmwk)

par(op)

```

### Explore GAMM structure

## first covar

```{r}
# Flow, inundation days, water temperature, day of year explored one at a time
# compared to null model
gam1.0 <- gam(log_chla ~ 1, method = "REML", data = x, family = "gaussian")

# gam1.1 <- gam(log_chla ~ s(log_qsdy) + s(station, bs="re"), method = "REML", data = x, family = "gaussian")
# gam1.2 <- gam(log_chla ~ s(log_idys) + s(station, bs="re"), method = "REML", data = x, family = "gaussian")
# gam1.3 <- gam(log_chla ~ s(WTmwk)    + s(station, bs="re"), method = "REML", data = x, family = "gaussian")
# gam1.4 <- gam(log_chla ~ s(Sradmwk)  + s(station, bs="re"), method = "REML", data = x, family = "gaussian")
# gam1.5 <- gam(log_chla ~ s(Q_sday)   + s(station, bs="re"), method = "REML", data = x, family = "gaussian")
# gam1.6 <- gam(log_chla ~ s(Inund_days) + s(station, bs="re"), method = "REML", data = x, family = "gaussian")

gam1.1 <- gam(log_chla ~ s(log_qsdy), method = "REML", data = x, family = "gaussian")
gam1.2 <- gam(log_chla ~ s(log_idys), method = "REML", data = x, family = "gaussian")
gam1.3 <- gam(log_chla ~ s(WTmwk),    method = "REML", data = x, family = "gaussian")
gam1.4 <- gam(log_chla ~ s(Sradmwk),  method = "REML", data = x, family = "gaussian")
gam1.5 <- gam(log_chla ~ s(Q_sday),   method = "REML", data = x, family = "gaussian")
gam1.6 <- gam(log_chla ~ s(inund_days), method = "REML", data = x, family = "gaussian")

# to make sure that k is not too high or too low
gam.check(gam1.0)
gam.check(gam1.1)
gam.check(gam1.2)
gam.check(gam1.3)
gam.check(gam1.4)
gam.check(gam1.5)
gam.check(gam1.6)

# to compare model fit; lowest AIC value will be first covar to be included in model
AIC(gam1.0, gam1.1, gam1.2, gam1.3, gam1.4, gam1.5, gam1.6)

# selects gam1.4 or Sradmwk
```

## explore other spline definitions

```{r}

# simple spline; no k given, bs is by default "tp"
gam1.4 <- gam(log_chla ~ s(Sradmwk),  method = "REML", data = x, family = "gaussian")

# added week + 52 knots as a tensor product with Srad (Pederson et al. paper)
# unclear on the difference between k and knots
gam4.4 <- gam(log_chla ~ te(week, Sradmwk,  bs=c("cc", "tp"), k = c(10, 10)), 
              method = "REML", data = x, family = "gaussian", knots = list(week=c(0, 52)))

# another way of including cyclic nature of data i.e. seasonality; performs as well as gam4.4
gam4.3 <- gam(log_chla ~ s(Sradmwk) + s(week, bs = "cc", k = 51), 
              method = "REML", data = x, family = "gaussian")

gam.check(gam1.4)
gam.check(gam4.4)

AIC(gam1.4, gam4.4, gam4.3)

# gam4.3 formulation carried forward as the simplest but best fitting

```


## second term of covar with model incorporating seasonality; first term Sradmwk

```{r}

gam5.1 <- gam(log_chla ~ s(Sradmwk) + s(week, bs = "cc", k = 51) + s(log_qsdy), 
              method = "REML", data = x, family = "gaussian")
gam5.2 <- gam(log_chla ~ s(Sradmwk) + s(week, bs = "cc", k = 51) + s(log_idys), 
              method = "REML", data = x, family = "gaussian")

AIC(gam4.3, gam5.1, gam5.2)

# 5.1 wins; log_qsdy is included next
# but also discovered that by including week, Sradmwk becomes 'not significant'
# thus Srad maybe only important as a marker to seasonality
```

## test out inundation flag as a by= with Q from same day

# also test out tensor product of log(inundation days)

```{r}

gam5.3 <- gam(log_chla ~ s(Sradmwk) + s(week, bs = "cc", k = 51) + 
                s(log_qsdy, by = inundation, m = 1, bs = "tp") + s(inundation, bs="re"), 
              method = "REML", data = x, family = "gaussian")

gam5.4 <- gam(log_chla ~ s(Sradmwk) + s(week, bs = "cc", k = 51) + te(log_qsdy, log_idys), 
              method = "REML", data = x, family = "gaussian")

AIC(gam5.1, gam5.3, gam5.4)

gam.check(gam5.1)
gam.check(gam5.4)

summary(gam5.1)
summary(gam5.4)

# 5.1 wins but 5.4 performs equally well

```



## include inundation days as tensor product with Q from same day

also remove week and include cyclic nature within Sradmwk (did not work)
also include week and remove Sradmwk which works better

```{r}

# include cyclic nature within Sradmwk
gam6.1 <- gam(log_chla ~ s(Sradmwk, bs = "cc", k = 51) + s(log_qsdy), 
              method = "REML", data = x, family = "gaussian")

# include cyclic nature within Sradmwk + tensor(log of Q and inundation days)
gam6.2 <- gam(log_chla ~ s(Sradmwk, bs = "cc", k = 51) + te(log_qsdy, log_idys), 
              method = "REML", data = x, family = "gaussian")

# include cyclic nature through week and drop Srad  + te(log_qsdy, log_idys)
gam6.3 <- gam(log_chla ~ s(week, bs = "cc", k = 51) + te(log_qsdy, log_idys), 
              method = "REML", data = x, family = "gaussian")

# include cyclic nature through week and drop Srad (no inundation days)
gam6.4 <- gam(log_chla ~ s(week, bs = "cc", k = 51) + s(log_qsdy), 
              method = "REML", data = x, family = "gaussian")

# include cyclic nature through week and drop Srad (inundation flag with inundation days)
gam6.5 <- gam(log_chla ~ s(week, bs = "cc", k = 51) + s(log_idys, by = inundation, m = 1, bs = "tp"), 
              method = "REML", data = x, family = "gaussian")

# compare all
AIC(gam5.1, gam5.3, gam6.1, gam6.2, gam6.3, gam6.4, gam6.5)

AIC(gam6.3, gam6.4, gam6.5)

# gam6.4 (only log_qsdy required for a good model)

```

## color ramp for diagnostic plots
```{r}
# add color bar to include Sradmwk to plots below
range(x$Sradmwk)
length(unique(x$Sradmwk))
rbPal <- colorRampPalette(c('red','blue'))
x$Col <- rbPal(10)[as.numeric(cut(x$Sradmwk,breaks = 10))]
length(unique(x$Col))
```

## plot residuals with fitted values, etc.
```{r}
# plot residuals with fitted values, etc.
par(mfrow = c(2,2))
gam.check(gam6.3)
gam.check(gam6.4)

# check if temporal autocorrelation is taken care of
resid = residuals(gam6.3, type = "working")
par(mfrow = c(2,2))
acf(resid)
pacf(resid)
plot(x$log_qsdy, resid, xlab = "Flow same day",   ylab = "Residuals", col = x$Col)
plot(x$log_idys, resid, xlab = "Inundation days", ylab = "Residuals", col = x$Col)

resid = residuals(gam6.4)
acf(resid)
pacf(resid)
plot(x$log_qsdy, resid, xlab = "Flow same day",   ylab = "Residuals", col = x$Col)

#install.packages("visreg")
library("visreg")

visreg2d(gam6.3, xvar='log_qsdy', yvar='log_idys', scale='response')

summary(gam6.3)
summary(gam6.4)

```

```{r}
library("visreg")

# attempt overlay of data
visreg2d(gam6.3, xvar='log_qsdy', yvar='log_idys', scale='response')
par(new = TRUE)
plot(x$log_qsdy, x$log_idys, xlim = c(-2.5,5)) # looks weird, but answers the question about no data in blue area of the plot

# 3-D plot that includes log(chl)
library(rgl)
visreg2d(gam6.3, xvar='log_qsdy', yvar='log_idys', scale='response', plot.type = "rgl")
```

## Include all regions (except cache) and scale parameters

```{r}

alldata.scaled <- alldata %>%
  mutate(across(.cols = c(log_idys,log_qsdy,diurnal_range:Sradmwk), ~ scale(.x)))
  
```

## New trials with random effects for station and time (maybe)

```{r}

gam1.0 <- gam(log_chla ~ 1, method = "REML", data = x, family = "gaussian")
gam1.2 <- gam(log_chla ~ s(Q_sday), method = "REML", data = x, family = "gaussian")
gam1.3 <- gam(log_chla ~ inundation + s(Q_sday, by = inundation, bs = "tp"), 
               method = "REML", data = x, family = "gaussian")
gam1.4 <- gamm(log_chla ~ inundation + s(Q_sday, by = inundation, bs = "tp"), 
               correlation = corAR1(), 
               method = "REML", data = x, family = "gaussian")
gam1.5 <- gamm(log_chla ~ inundation + s(Q_sday, by = inundation, bs = "tp"), 
               correlation = corAR1(form = ~1|doy1998), 
               method = "REML", data = x, family = "gaussian")
gam1.6 <- gamm(log_chla ~ inundation + s(Q_sday, by = inundation, bs = "tp"), 
               correlation = corAR1(form = ~1|doy1998), random = list(station = ~1),
               method = "REML", data = x, family = "gaussian")

AIC(gam1.0, gam1.1, gam1.2, gam1.3, gam1.4, gam1.5, gam1.6)

# Best model here is gam1.4
# accounting for autocorr. without giving any time var
# I assume this means that it is implying autocorr in residuals
# plus inundation flag - check is this is necessary

gam1.4 <- gamm(log_chla ~ inundation + s(Q_sday, by = inundation, bs = "tp"), 
               correlation = corAR1(), 
               method = "REML", data = x, family = "gaussian")
gam1.7 <- gamm(log_chla ~ s(Q_sday, bs = "tp"), correlation = corAR1(), 
               method = "REML", data = x, family = "gaussian")

AIC(gam1.4, gam1.7)

```

### discussion with Jereme

1. try constraining the time series
2. categorize inundation (high, low, no_inundation)
3. year as random effect
4. plots of normalized data
5. plots of response vs. covar
6. residuals mean 0 and variance similar across the range


1. "Days_since_last_inundation"
2. "days_of_inundation_until_now"
3. "total_inund_last_year"
4. "inundation_flag"

Model 0: log_chla = s(Q_sday, by = inundated)
Model 1: log_chla = s(Q_sday, by = inundated) + s(doy)
Model 2: log_chla = te(Q_sday, doy, by = inundated)
Model 3: log_chla = te(Q_sday, cumulative_idys, by = inundated (0,1))
Model 4: log_chls = te(Q_sday, doy, days_since_inundation, by = inundated)

```{r}

library("mgcv")
gam0 <- gam(log_chla ~ 1, method = "REML", data = x, family = "gaussian")

gam1.0 <- gam(log_chla ~ s(log_qsdy, by = inundation), method = "REML", data = x, family = "gaussian")
gam1.1 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = x, family = "gaussian")
gam1.2 <- gam(log_chla ~ te(log_qsdy, dowy, by = inundation), method = "REML", data = x, family = "gaussian")
gam1.3 <- gam(log_chla ~ te(log_qsdy, log_idys, dowy, by = inundation), method = "REML", data = x, family = "gaussian")

AIC(gam0, gam1.0, gam1.1, gam1.2, gam1.3)

gam.check(gam1.0)
gam.check(gam1.1)
gam.check(gam1.2)
gam.check(gam1.3)

summary(gam1.1)
summary(gam1.2)
summary(gam1.3)

# plot residuals with fitted values, etc.
par(mfrow = c(2,2), pch=18)
gam.check(gam1.2)
gam.check(gam1.3)

par(mfrow = c(1,2), pch = 18)
plot(gam1.3)


```

## constrain to range when Yolo Bypass gets inundated

```{r}

x.idys = subset(x, x$inundation == 1)
range(x.idys$dowy)
# min "day of water year" 65; max 214

x.idys = subset(x, (x$dowy >= 65 & x$dowy <= 214))
table(x.idys$inundation)

gam1.0 <- gam(log_chla ~ s(log_qsdy, by = inundation), method = "REML", data = x.idys, family = "gaussian")
gam1.1 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = x.idys, family = "gaussian")
gam1.2 <- gam(log_chla ~ te(log_qsdy, dowy, by = inundation), method = "REML", data = x.idys, family = "gaussian")
gam1.3 <- gam(log_chla ~ te(log_idys, dowy, by = inundation), method = "REML", data = x.idys, family = "gaussian")

AIC(gam0, gam1.0, gam1.1, gam1.2, gam1.3)

gam.check(gam1.1)
gam.check(gam1.2)
gam.check(gam1.3)
gam.check(gam1.4)

summary(gam1.1)
summary(gam1.2)
summary(gam1.3)
summary(gam1.4)

resid = residuals(gam1.3)
acf(resid)
pacf(resid)

```


#### Model validation for the gam model based on hypothesis testing

```{r}

  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_qsdy.tiff", width = 10, height = 8, units = "in", res = 200)
  resid = residuals(gam1.3)
  op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gam1.3), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(x.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  par(op)
  dev.off()

  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_idys.tiff", width = 10, height = 8, units = "in", res = 200)
  resid = residuals(gam1.4)
  op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gam1.4), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(x.idys$log_idys, resid, xlab = "Consecutive inundation days",   ylab = "Residuals")
  par(op)
  dev.off()

```
