---
title: "make_gam_model_sk"
author: "Shruti Khanna"
date: "4/28/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
Variety of Modeling Techniques for Connectivity Synthesis
* LM
* GLS
* GLMM
* GAM
* GAMM
What will we choose?



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Connectivity Modeling - Chlorophyll a

## Preparation

### Call packages
```{r}
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)

# tinytex::install_tinytex() - run in console before knitting
```

### Read in data

```{r}

alldata <- read_csv("data_model/model_chla_covars_gam.csv")%>%
  mutate(month = lubridate::month(date),
         year = lubridate::year(date),
         doy = lubridate::yday(date)-60,
         water_year = ifelse(month>9, year + 1, year),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days+1),
         log_qsdy = log(Q_sday)) %>%
  rename(station = station_wq_chl)

```

Filter to downstream
```{r}
downstream <- alldata %>% filter(region == "below")
```

Look at covariates
```{r}
downstream_long <- pivot_longer(downstream, cols = c(diurnal_range:Sradmwk, log_idys, log_qsdy), names_to = "Parameter", values_to = "Value")

ggplot(downstream_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

These are very different scales. Scale covars.
```{r}
downstream_scaled <- downstream %>%
  mutate(across(.cols = c(log_idys,log_qsdy,diurnal_range:Sradmwk), ~ scale(.x)))
```

Check distribution.
```{r}
downstream_s_long <- pivot_longer(downstream_scaled, cols = c(log_idys,log_qsdy,diurnal_range:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(downstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```


#### Check for autocorrelation in model, predictor, response
```{r}

x <- downstream_scaled

op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)

acf(x$log_chla)
acf(x$Q_sday)
acf(x$WTmwk)
acf(x$Sradmwk)

par(op)

```



### lme trials
All variables

```{r}
  
  # make sure that only one date per station is present in the final dataset
  y = x %>% group_by(station, doy1998) %>% sample_n(1)

  # decide the random structure & autocorr. with a full fixed model and method REML
  ms.0 = gls(log_chla ~ 1 + inundation + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = x)

  ms.1 = lme(log_chla ~ 1 + inundation + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = x,
             random = ~ 1 | station)
  
  ms.2 = lme(log_chla ~ 1 + inundation + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = x,
             random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))
  
  AIC(ms.0, ms.1, ms.2)
  BIC(ms.0, ms.1, ms.2)
  # include both random and corARMA structure
  
```



```{r}

  ms.11 = gls(log_chla ~ 1, method = "ML", data = x)

  # which predictor should be included next
  ms.21 = lme(log_chla ~ 1 + Q_sday,   method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))
  
  ms.22 = lme(log_chla ~ 1 + inund_days, method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))
  
  ms.23 = lme(log_chla ~ 1 + WTmwk,    method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))
  
  ms.24 = lme(log_chla ~ 1 + Sradmwk, method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))

  ms.25 = lme(log_chla ~ 1 + inundation, method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))

  AIC(ms.21, ms.22, ms.23, ms.24, ms.25)
  BIC(ms.21, ms.22, ms.23, ms.24, ms.25)
  # first var to be included is Sradmwk; best model ms.24
  
```
  

```{r}

  # which second predictor should be included (or not)
  ms.31 = lme(log_chla ~ 1 + Sradmwk + Q_sday,   method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.32 = lme(log_chla ~ 1 + Sradmwk + WTmwk,    method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.33 = lme(log_chla ~ 1 + Sradmwk + inund_days, method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.34 = lme(log_chla ~ 1 + Sradmwk + inundation, method = "ML", data = x, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))

  AIC(ms.24, ms.31, ms.32, ms.33, ms.34)
  BIC(ms.24, ms.31, ms.32, ms.33, ms.34)
  # best model as yet - ms.33 i.e. Sradmwk + inund_days but ms.31 is close
  
```


```{r}

  # interaction should be included (or not)
  ms.33 = lme(log_chla ~ 1 + Sradmwk + inund_days, method = "ML", data = y, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.43 = lme(log_chla ~ 1 + Sradmwk + inund_days + Sradmwk*inund_days, method = "ML", data = y, 
              random = ~ 1 | station,
              correlation = corARMA(form =~ doy1998|station, p=1, q=1))

  AIC(ms.33, ms.43)
  BIC(ms.33, ms.43)
  # best model is without interaction
  
```



```{r}

  # which third predictor should be included (or not)
  ms.51 = lme(log_chla ~ 1 + Sradmwk + inund_days + Q_sday, method = "ML", data = x, 
              random = ~ 1 | station,
              correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.52 = lme(log_chla ~ 1 + Sradmwk + inund_days + WTmwk,  method = "ML", data = x, 
              random = ~ 1 | station,
              correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.53 = lme(log_chla ~ 1 + Sradmwk + inund_days + inundation, method = "ML", data = x, 
              random = ~ 1 | station,
              correlation = corARMA(form =~ doy1998|station, p=1, q=1))

  AIC(ms.33, ms.51, ms.52, ms.53)
  BIC(ms.33, ms.51, ms.52, ms.53)
  # best model as yet - ms.33 i.e. Sradmwk + inund_days; no third covar or interaction
  
```



```{r}
  # best model was ms.33 with just Sradmwk, inundation, random effect and autocorr. structure
  ms.33.reml = lme(log_chla ~ 1 + Sradmwk + inund_days, method = "REML", data = y, 
                   random = ~ 1 | station,
                   correlation = corARMA(form =~ doy1998|station, p=1, q=1))

  resid = residuals(ms.33.reml, type = "normalized")
  acf(resid)
  pacf(resid)
  summary(ms.33.reml)
  
```


#### Model validation for the lme model selection
```{r}

  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/lme_model_validation.tiff", width = 10, height = 8, units = "in", res = 200)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(ms.33.reml), resid, xlab = "Fitted values", ylab = "Norm. Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Norm. Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(x$Sradmwk,    resid, xlab = "Mean Solar Radiation per week", ylab = "Norm. Residuals")
  plot(x$inund_days, resid, xlab = "Consecutive inundation days",   ylab = "Norm. Residuals")
  par(op)
  dev.off()

  tiff(filename = "figures/lme_model_visualization.tiff", width = 10, height = 8, units = "in", res = 200)
  
  plot(ms.33.reml)
  par(op)
  dev.off()

```


### lme trials separting inundated from non-inundated periods

## All except inundation related variables because for dry period

```{r}
  
  # make sure that only one date per station is present in the final dataset
  # divide the inundated dataset from the non-inundated dataset
  y.wet = subset(x, inundation == 1)
  y.dry = subset(x, inundation == 0)

  # decide the random structure & autocorr. with a full fixed model and method REML
  ms.0 = gls(log_chla ~ 1 + Q_sday + WTmwk + Sradmwk, method = "REML", data = y.dry)

  ms.1 = lme(log_chla ~ 1 + Q_sday + WTmwk + Sradmwk, method = "REML", data = y.dry,
             random = ~ 1 | station)
  
  ms.2 = lme(log_chla ~ 1 + Q_sday + WTmwk + Sradmwk, method = "REML", data = y.dry,
             random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  BIC(ms.0, ms.1, ms.2)
  
  # best model is ms.2

  ms.11 = gls(log_chla ~ 1, method = "ML", data = y.dry)

  # which predictor should be included next
  ms.21 = lme(log_chla ~ 1 + Q_sday,   method = "ML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.22 = lme(log_chla ~ 1 + Sradmwk, method = "ML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.23 = lme(log_chla ~ 1 + WTmwk,    method = "ML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  BIC(ms.11, ms.21, ms.22, ms.23)
  
  # best model is ms.22 so Sradmwk most important
  
  # which second predictor should be included (or not)
  ms.31 = lme(log_chla ~ 1 + Sradmwk + WTmwk,    method = "ML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.32 = lme(log_chla ~ 1 + Sradmwk + Q_sday, method = "ML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  BIC(ms.22, ms.31, ms.32)
  
  # best model was still ms.22 with just Srad mean/week, random effect and autocorr. structure
  ms.22.reml = lme(log_chla ~ 1 + Sradmwk, method = "REML", data = y.dry, random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998|station, p=1, q=1))

  resid = residuals(ms.22.reml, type = "normalized")
  acf(resid)
  pacf(resid)
  summary(ms.22.reml)
  
```


### lme trials for inundated period only - y.wet
All variables but only for inundated period hence no "inundation" flag

```{r}
  
  # decide the random structure & autocorr. with a full fixed model and method REML
  ms.0 = gls(log_chla ~ 1 + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = y.wet)

  ms.1 = lme(log_chla ~ 1 + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = y.wet,
             random = ~ 1 | station)
  
  ms.2 = lme(log_chla ~ 1 + Q_sday + WTmwk + inund_days + Sradmwk, method = "REML", data = y.wet,
             random = ~ 1 | station,
             correlation = corARMA(form =~ doy1998, p=1, q=1))
  
  AIC(ms.0, ms.1, ms.2)
  BIC(ms.0, ms.1, ms.2)
  # include both random and corARMA structure
  
```



```{r}

  ms.11 = gls(log_chla ~ 1, method = "ML", data = y.wet)

  # # which predictor should be included next
  # ms.21 = lme(log_chla ~ 1 + Q_sday,   method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998, p=1, q=1))
  # 
  # ms.22 = lme(log_chla ~ 1 + inund_days, method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998, p=1, q=1))
  # 
  # ms.23 = lme(log_chla ~ 1 + WTmwk,    method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998, p=1, q=1))
  # 
  # ms.24 = lme(log_chla ~ 1 + Sradmwk, method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998, p=1, q=1))

  # which predictor should be included next
  ms.21 = gls(log_chla ~ 1 + Q_sday,     method = "ML", data = y.wet)
  
  ms.22 = gls(log_chla ~ 1 + inund_days, method = "ML", data = y.wet)
  
  ms.23 = gls(log_chla ~ 1 + WTmwk,      method = "ML", data = y.wet)
  
  ms.24 = gls(log_chla ~ 1 + Sradmwk,    method = "ML", data = y.wet)

  AIC(ms.11, ms.21, ms.22, ms.23, ms.24)
  BIC(ms.11, ms.21, ms.22, ms.23, ms.24)
  # first var to be included is inundation days; best model ms.22
  
```
  

```{r}

  # which second predictor should be included (or not)
  # ms.31 = lme(log_chla ~ 1 + inund_days + Q_sday,   method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  # 
  # ms.32 = lme(log_chla ~ 1 + inund_days + WTmwk,    method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  # 
  # ms.33 = lme(log_chla ~ 1 + inund_days + Sradmwk, method = "ML", data = y.wet, random = ~ 1 | station,
  #            correlation = corARMA(form =~ doy1998|station, p=1, q=1))
  
  ms.31 = gls(log_chla ~ 1 + inund_days + Q_sday,   method = "ML", data = y.wet)
  
  ms.32 = gls(log_chla ~ 1 + inund_days + WTmwk,    method = "ML", data = y.wet)
  
  ms.33 = gls(log_chla ~ 1 + inund_days + Sradmwk, method = "ML", data = y.wet)

  AIC(ms.22, ms.31, ms.32, ms.33)
  BIC(ms.22, ms.31, ms.32, ms.33)
  # best model as yet - ms.22 i.e. inund_days only
  
```


```{r}
  # best model was ms.33 with just Sradmwk, inundation, random effect and autocorr. structure
  ms.22.reml = gls(log_chla ~ 1 + inund_days, method = "REML", data = y.wet)

  resid = residuals(ms.33.reml, type = "normalized")
  acf(resid)
  pacf(resid)
  summary(ms.22.reml)
  
```


#### Model validation for the lme model selection
```{r}

  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/lme_model_validation.tiff", width = 10, height = 8, units = "in", res = 200)
  op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(ms.22.reml), resid, xlab = "Fitted values", ylab = "Norm. Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Norm. Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(x$inund_days, resid, xlab = "Consecutive inundation days",   ylab = "Norm. Residuals")
  par(op)
  dev.off()

  tiff(filename = "figures/lme_model_visualization.tiff", width = 10, height = 8, units = "in", res = 200)
  
  plot(ms.33.reml)
  par(op)
  dev.off()

```



