---
title: "make_gam_noninund"
author: "Cat"
date: '2022-08-09'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)
library("AICcmodavg")

# tinytex::install_tinytex() - run in console before knitting
```

# Read in data
```{r}
alldata0 <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         month_fac = as.factor(month),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         doy = lubridate::yday(date),
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         wy_fac = as.factor(water_year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         t_qsdy = kader:::cuberoot(Q_sday), 
         log_qsdy = log(Q_sday),
         inund_fac = ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short")),
         inundation = as.factor(inundation),
         station_wq_chl = as.factor(station_wq_chl),
         tily_fac = case_when(total_inund_last_year == 0 ~ "none",
                              total_inund_last_year>0 & total_inund_last_year < 16 ~ "2wk",
                              total_inund_last_year>16 & total_inund_last_year < 41 ~ "month",
                              TRUE ~ "months"),
         tily_fac = as.factor(tily_fac),
         inund_fac = as.factor(inund_fac)) %>%
  rename(station = station_wq_chl, tily = total_inund_last_year, 
         dsli = days_since_last_inundation, idysun = days_of_inundation_until_now)

alldatalook <- select(alldata0, date, tily, tily_fac)
str(alldata0)
```


Filter to inundation period
```{r}
inundPd <- alldata0 %>% filter(inundation == 1) 
inMin <- min(inundPd$dowy)
inMax <- max(inundPd$dowy)

alldata <- alldata0 %>%
  filter(dowy < inMin | dowy > inMax)
  #mutate(dnoni0 = ifelse(dowy < 65, dowy+365, dowy),
  #       dnoni = ifelse(year%%4>0, dnoni0-214, dnoni0-215))
```

```{r}
ggplot(alldata, aes(x = month, y = doy, color = month)) + geom_point() + theme_bw()
```
May - early December

Model upstream

## Upstream
```{r}
upstream <- alldata %>% filter(region == "above")
```

### Filter to upstream

```{r}
upstream_scaled <- upstream %>%
  mutate(across(.cols = c(diurnal_range:Sradmwk, t_qsdy:log_qsdy), ~ scale(.x)))
  
u.noni <- upstream_scaled

upstream_s_long <- pivot_longer(u.noni, cols = c(diurnal_range:Sradmwk, log_idys:log_qsdy), names_to = "Parameter", values_to = "Value")

ggplot(upstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

### Model 

* log_qsdy
* WTrangemwk
* WTmwk
* diurnal_range
* doy
* month

```{r}
gamu0 <- gam(log_chla ~ 1, method = "REML", data = u.noni, family = "gaussian")
gamu1 <- gam(log_chla ~ s(log_qsdy), method = "REML", data = u.noni, family = "gaussian")
gamu2 <- gam(log_chla ~ s(log_qsdy) + s(doy, bs = "cc"), method = "REML", data = u.noni, family = "gaussian")
gamu3 <- gam(log_chla ~ s(WTmwk), method = "REML", data = u.noni, family = "gaussian")
gamu4 <- gam(log_chla ~ s(WTmwk) + s(doy, bs = "cc"), method = "REML", data = u.noni, family = "gaussian")
gamu5 <- gam(log_chla ~ s(Sradmwk), method = "REML", data = u.noni, family = "gaussian")
gamu6 <- gam(log_chla ~ s(Sradmwk) + s(doy, bs = "cc"), method = "REML", data = u.noni, family = "gaussian")
gamu7 <- gam(log_chla ~ s(WTrangemwk), method = "REML", data = u.noni, family = "gaussian")
gamu8 <- gam(log_chla ~ s(WTrangemwk) + s(doy, bs = "cc"), method = "REML", data = u.noni, family = "gaussian")
gamu9 <- gam(log_chla ~ te(WTmwk, doy, bs = c( "tp", "cc")), method = "REML", data = u.noni, family = "gaussian")
gamu10 <- gam(log_chla ~ te(WTmwk, log_qsdy, doy, bs = c("tp", "tp", "cc")), method = "REML", data = u.noni, family = "gaussian")
gamu11 <- gam(log_chla ~ te(WTmwk, Sradmwk, doy, bs = c("tp", "tp", "cc")), method = "REML", data = u.noni, family = "gaussian")
gamu12 <- gam(log_chla ~ te(WTmwk, WTrangemwk, doy, bs = c("tp", "tp", "cc")), method = "REML", data = u.noni, family = "gaussian")
gamu13 <- gam(log_chla ~ te(WTmwk, diurnal_range, doy, bs = c("tp", "tp", "cc")), method = "REML", data = u.noni, family = "gaussian")
gamu14 <- gam(log_chla ~ te(WTmwk, month, log_qsdy), method = "REML", data = u.noni, family = "gaussian")
gamu15 <- gam(log_chla ~ te(Sradmwk, month, log_qsdy), method = "REML", data = u.noni, family = "gaussian")
#gamu15 <- gam(log_chla ~ s(month), method = "REML", data = u.noni, family = "gaussian")
# AIC

(table_aic <- AIC(gamu0, gamu1, gamu2, gamu3, gamu4, gamu5, gamu6, gamu7, gamu8, gamu9, gamu10, gamu11, gamu12, gamu13, gamu14, gamu15)%>% arrange(AIC))

summary(gamu10) #0.503 | 55.4% 
summary(gamu14) #0.494 | 54.1%
summary(gamu13) #0.455|  52.2%
summary(gamu9)  #0.413|  45.5%


```

There is some autocorrelation here
```{r}
par(mfrow = c(2,2))
gam.check(gamu10)
gam.check(gamu13)
gam.check(gamu9)

resid = residuals(gamu10)
acf(resid)
pacf(resid)
```

## Prediction
https://cran.r-project.org/web/packages/tidymv/vignettes/predict-gam.html

```{r}
library(tidymv)

pred_gamu10_q <- predict_gam(gamu10, values = list(log_qsdy = c(quantile(u.noni$log_qsdy, 0.2),
                                                          quantile(u.noni$log_qsdy, 0.8), 
                                                          median(u.noni$log_qsdy)),
                                             WTmwk = c(quantile(u.noni$WTmwk, 0.2),
                                                       quantile(u.noni$WTmwk, 0.8)))) %>%
  mutate(WTmwk_orig = (WTmwk*sd(upstream$WTmwk)+mean(upstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamu10_wt <- predict_gam(gamu10, values = list(WTmwk = c(quantile(u.noni$WTmwk, 0.2),
                                                          quantile(u.noni$WTmwk, 0.8), 
                                                          median(u.noni$WTmwk)),
                                             log_qsdy = c(quantile(u.noni$log_qsdy, 0.2),
                                                          quantile(u.noni$log_qsdy, 0.8)))) %>%
   mutate(WTmwk_orig = (WTmwk*sd(upstream$WTmwk)+mean(upstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamu10_doy <- predict_gam(gamu10, values = list(doy = c(quantile(u.noni$doy, 0.2),
                                                          quantile(u.noni$doy, 0.8), 
                                                          median(u.noni$doy)),
                                             log_qsdy = c(quantile(u.noni$log_qsdy, 0.2),
                                                          quantile(u.noni$log_qsdy, 0.8)))) %>%
    mutate(WTmwk_orig = (WTmwk*sd(upstream$WTmwk)+mean(upstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamu10_doy_qsdy <- predict_gam(gamu10, values = list(doy = c(quantile(u.noni$doy, 0.1),
                                                          quantile(u.noni$doy, 0.3),
                                                          quantile(u.noni$doy, 0.5), 
                                                          quantile(u.noni$doy, 0.7), 
                                                          quantile(u.noni$doy, 0.9)),
                                             log_qsdy = c(quantile(u.noni$log_qsdy, 0.2),
                                                          quantile(u.noni$log_qsdy, 0.8), 
                                                          median(u.noni$log_qsdy)))) %>%
  mutate(WTmwk_orig = (WTmwk*sd(upstream$WTmwk)+mean(upstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamu10_all <- predict_gam(gamu10) %>%
   mutate(WTmwk_orig = (WTmwk*sd(upstream$WTmwk)+mean(upstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))
```


## Visualization
* CI = 95% CIs

Temperatures are only applicable to certain months. This may be true for flow too. 
```{r}
ggplot(upstream, aes(x = doy, y = WTmwk, color = factor(month))) + geom_point(size = 3) + theme_bw() + scale_color_viridis(discrete = TRUE) + labs(title = "Temperature possibilities")

ggplot(upstream, aes(x = doy, y = log_qsdy, color = factor(month))) + geom_point(size = 3) + theme_bw() + scale_color_viridis(discrete = TRUE) + labs(title = "Flow possibilities")
```


### by q
* We are not able to really predict middles of these, because 16 degrees not relevant July/August and 21 degrees not relevant 207+
```{r}
pred_gamu10_q %>%
  ggplot(aes(doy, fit)) +
  geom_smooth_ci(log_qsdy, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~WTmwk_orig,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()
```

### by water temp
Low chlorophyll around July, higher more towards May and October.
Similar patterns for low and high flow. Low water temp has different pattern from higher water temps. 
```{r}
pred_gamu10_wt %>%
  ggplot(aes(doy, fit)) +
  geom_smooth_ci(WTmwk_orig, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  labs(y = "chlorophyll")+
  theme_bw()
```

### day of year
Pattern similar across low and high flow.
Increase in chlorophyll with increase in water temp (generally), though weird decrease around WT = 18
```{r}
pred_gamu10_doy %>%
  ggplot(aes(WTmwk_orig, fit)) +
  geom_smooth_ci(doy, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = "chlorophyll")+
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  theme_bw()
```

### by flow, and faceted by ~approximately month?
For higher flows, increase with increasing temperature. For lower flow, not much clear pattern
```{r}
pred_gamu10_doy_qsdy %>%
  ggplot(aes(WTmwk_orig, fit)) +
  geom_smooth_ci(log_qsdy, size = 1.5) +
  facet_wrap(~doy) + 
  scale_color_viridis(discrete = TRUE) +
  labs(y = "chlorophyll")+
  theme_bw()
```


```{r}
pred_gamu10_all %>%
  ggplot(aes(WTmwk_orig, log_qsdy, z = fit)) +
  geom_tile(aes(fill = fit)) +
  geom_contour(colour = "white") +
  scale_fill_viridis(name = "chla") +
  theme_minimal() +
  theme(legend.position = "top")

pred_gamu10_all %>%
  ggplot(aes(doy, log_qsdy, z = fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(colour = "white") +
  scale_fill_viridis(name = "chla") +
  theme_minimal() +
  theme(legend.position = "top")
```

# Conclusions
- Higher chlorophyll with higher flow and temperature
- Some dips in chlorophyll around doy 250 (July)

# Questions
* How much should we be playing with k? 
* How to deal with autocorrelation
* Should we add random effects?
* Should doy be scaled?
