---
title: "make_gam_byregion"
author: "NCEAS Connectivity Group"
date: '2022-07-12'
output: html_document
editor_options: 
  chunk_output_type: console
---
This code runs gams by region, looking at the relationship between various covariates and chla. 


Questions/Notes:

1. For inundation filtering, do we want to do this for each region separately or use the min and max overall? 
2. Are we filtering inundation for upstream?
3. We didn't model upstream yet, right? 
4. Add other figures to look at beginning and end of flooding?
5. Can we turn some of this code into functions to reduce repetition?
6. Rename repeated variables - I tried to do this for some of it, but needs checking.
7. Nutrients
8. Inundation metrics

To do 
- Add water temperature
- Explore WTrange, diurnal range (variance)
- Compare Srad vs dowy
- Random effects of station, maybe water year
- Explore total_inund_last_year, days_since_last_inundation, 3-category inundation flag (short, long, none)
- Model non-inundation periods

To do 7/26/2022
- Build model for non-inundation periods
- Visualize all models
- Random effects
- de-seasonalized water temp


To do 8/4/2022
- Shruti to look at yolo (tily_fac, visualizations)
- Shruti to look at downstream
- Cat to look at upstream - look at diff between inundation and non inundation subset days - see if same variables come up
- Cat see if she can streamline code



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)

# tinytex::install_tinytex() - run in console before knitting
```

# Read in data
```{r}

alldata0 <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(month = lubridate::month(date),
         year  = lubridate::year(date),
         rdoy  = lubridate::yday(date) + 92,
         week  = lubridate::week(date),
         water_year = ifelse(month > 9, year + 1, year),
         dowy = ifelse(rdoy > 366, rdoy - 366, rdoy),
         log_chla = log(chlorophyll),
         log_idys = log(inund_days + 1),
         t_qsdy = kader:::cuberoot(Q_sday), 
         log_qsdy = log(Q_sday),
         inund_fac = ifelse(inund_days == 0, "none", ifelse(inund_days > 14, "long", "short")),
         inundation = as.factor(inundation),
         station_wq_chl = as.factor(station_wq_chl),
         tily_fac = case_when(total_inund_last_year == 0 ~ "none",
                              total_inund_last_year>0 & total_inund_last_year < 16 ~ "2wk",
                              total_inund_last_year>16 & total_inund_last_year < 41 ~ "month",
                              TRUE ~ "months"),
         tily_fac = as.factor(tily_fac),
         inund_fac = as.factor(inund_fac)) %>%
  rename(station = station_wq_chl, tily = total_inund_last_year, 
         dsli = days_since_last_inundation, idysun = days_of_inundation_until_now)

alldatalook <- select(alldata0, date, tily, tily_fac)
str(alldata0)


```


Filter to inundation period
```{r}
inundPd <- alldata0 %>% filter(inundation == 1) 
inMin <- min(inundPd$dowy)
inMax <- max(inundPd$dowy)

alldata <- alldata0 %>%
  filter(dowy >= inMin & dowy <= inMax)
```


## Downstream

### Filter to downstream

```{r}
downstream <- alldata %>% filter(region == "below")
```

```{r}
downstream_scaled <- downstream %>%
  mutate(across(.cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), ~ scale(.x)))
  
d.idys <- downstream_scaled

downstream_s_long <- pivot_longer(d.idys, cols = c(log_idys, t_qsdy, log_qsdy, tily:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(downstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```


### Model

1. try constraining the time series
2. categorize inundation (high, low, no_inundation)
3. year as random effect (didn't do)
4. plots of normalized data
5. plots of response vs. covar
6. residuals mean 0 and variance similar across the range

1. "Days_since_last_inundation"
2. "days_of_inundation_until_now"
3. "total_inund_last_year"
4. "inundation_flag"

Model 0: log_chla = s(Q_sday, by = inundated)
Model 1: log_chla = s(Q_sday, by = inundated) + s(dowy)
Model 2: log_chla = te(Q_sday, dowy, by = inundated)
Model 3: log_chla = te(Q_sday, cumulative_idys, by = inundated (0,1))
Model 4: log_chla = te(Q_sday, dowy, by = inundation level)
Model 5: log_chla = te(Q_sday, dowy, days_since_inundation, by = inundated)


```{r}
gamd0 <- gam(log_chla ~ 1, method = "REML", data = d.idys, family = "gaussian")
gamd1.0 <- gam(log_chla ~ s(log_qsdy), method = "REML", data = d.idys, family = "gaussian")
gamd1.1 <- gam(log_chla ~ s(log_qsdy) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.1b <- gam(log_chla ~ te(log_qsdy, dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.2 <- gam(log_chla ~ s(log_idys), method = "REML", data = d.idys, family = "gaussian")
gamd1.3 <- gam(log_chla ~ s(log_idys) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.4 <- gam(log_chla ~ s(WTmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.5 <- gam(log_chla ~ s(WTmwk) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.6 <- gam(log_chla ~ s(Sradmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.7 <- gam(log_chla ~ s(Sradmwk) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.8 <- gam(log_chla ~ s(WTrangemwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.9 <- gam(log_chla ~ s(WTrangemwk) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.10 <- gam(log_chla ~ s(diurnal_range), method = "REML", data = d.idys, family = "gaussian")
gamd1.11 <- gam(log_chla ~ s(diurnal_range) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.12 <- gam(log_chla ~ s(tily), method = "REML", data = d.idys, family = "gaussian")
gamd1.13 <- gam(log_chla ~ s(tily) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.14 <- gam(log_chla ~ s(dsli), method = "REML", data = d.idys, family = "gaussian")
gamd1.15 <- gam(log_chla ~ s(dsli) + s(dowy), method = "REML", data = d.idys, family = "gaussian")

gamd1.16 <- gam(log_chla ~ s(log_qsdy, by = inundation), method = "REML", data = d.idys, family = "gaussian")
gamd1.17 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.18 <- gam(log_chla ~ te(log_qsdy, dowy, by = inundation), method = "REML", data = d.idys, family = "gaussian")
gamd1.19 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
gamd1.20 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = d.idys, family = "gaussian")
gamd1.21 <- gam(log_chla ~ te(log_qsdy, dowy, by = inund_fac),  method = "REML", data = d.idys, family = "gaussian")

gamd1.22 <- gam(log_chla ~ te(log_qsdy, tily, dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.23 <- gam(log_chla ~ te(log_qsdy, dsli, dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.24 <- gam(log_chla ~ te(log_qsdy, diurnal_range, dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.25 <- gam(log_chla ~ te(log_qsdy, WTrangemwk, dowy), method = "REML", data = d.idys, family = "gaussian")
gamd1.26 <- gam(log_chla ~ te(log_qsdy, WTmwk, dowy), method = "REML", data = d.idys, family = "gaussian")

gamd1.27 <- gam(log_chla ~ te(log_qsdy, tily, Sradmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.28 <- gam(log_chla ~ te(log_qsdy, dsli, Sradmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.29 <- gam(log_chla ~ te(log_qsdy, diurnal_range, Sradmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.30 <- gam(log_chla ~ te(log_qsdy, WTrangemwk, Sradmwk), method = "REML", data = d.idys, family = "gaussian")
gamd1.31 <- gam(log_chla ~ te(log_qsdy, WTmwk, Sradmwk), method = "REML", data = d.idys, family = "gaussian")

resd <- (AIC(gamd0, gamd1.0, gamd1.1,gamd1.1b, gamd1.2, gamd1.3, gamd1.4, gamd1.5, gamd1.6, gamd1.7, gamd1.8, gamd1.9, gamd1.10, gamd1.11, gamd1.12, gamd1.13, gamd1.14, gamd1.15, gamd1.16, gamd1.17, gamd1.18, gamd1.19, gamd1.20, gamd1.21, gamd1.22, gamd1.23, gamd1.24,
gamd1.25, gamd1.26, gamd1.27, gamd1.28, gamd1.29, gamd1.30, gamd1.31))
resd %>% arrange(AIC)

gam.check(gamd1.28)
gam.check(gamd1.26)
gam.check(gamd1.3)
gam.check(gamd1.13)
gam.check(gamd1.24)
gam.check(gamd1.23)

summary(gamd1.28) # 32.3%, 0.29
summary(gamd1.26) #33.7%, 0.293
summary(gamd1.3) #28% R2 0.26
summary(gamd1.13) # 26.4% R2 0.243
summary(gamd1.24) #31% 0.272
summary(gamd1.23) # 30% 0.264

resid = residuals(gamd1.28)
acf(resid)
pacf(resid)

```

#### Model validation for the gam model based on hypothesis testing

```{r}

  # validate model gamd1.2 - log_chla ~ te(log_qsdy, dowy, by = inundation)
  # -----------------------------------------------------------------------
  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_qsdy.tiff", width = 6, height = 9, units = "in", res = 100)
  resid = residuals(gamd1.2)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd1.2), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(d.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(d.idys$dowy, resid, xlab = "day of water year",   ylab = "Residuals")
  plot(d.idys$inundation, resid, xlab = "inundation flag",   ylab = "Residuals")
  par(op)
  dev.off()

  # validate model gamd1.3 - log_chla ~ te(log_idys, dowy, by = inundation)
  # -----------------------------------------------------------------------
  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_idys.tiff", width = 8, height = 6, units = "in", res = 100)
  resid = residuals(gamd1.3)
  op = par(mfrow = c(2, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd1.3), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(d.idys$log_idys, resid, xlab = "Consecutive inundation days",   ylab = "Residuals")
  par(op)
  dev.off()

  # validate model gamd1.4 - log_chla ~ te(log_qsdy, dowy, by = inund_fac)
  # -----------------------------------------------------------------------
  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_qsdy_ifac.tiff", width = 8, height = 10, units = "in", res = 100)
  resid = residuals(gamd1.4)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd1.4), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(d.idys$log_qsdy, resid, xlab = "Flow same day",   ylab = "Residuals")
  #plot(d.idys$dowy, resid, xlab = "day of water year",   ylab = "Residuals")
  plot(d.idys$inund_fac, resid, xlab = "inundation factor",   ylab = "Residuals")
  plot(d.idys$WTmwk, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()

```

#### Visualize gamd1.28

```{r}
library(tidymv)
gamd1.28 <- gam(log_chla ~ te(log_qsdy, dsli, Sradmwk), method = "REML", data = d.idys, family = "gaussian")


pred_gamd1.28_q <- predict_gam(gamd1.28, values = list(log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8), 
                                                          median(d.idys$log_qsdy)),
                                             Sradmwk = c(quantile(d.idys$Sradmwk, 0.2),
                                                       quantile(d.idys$Sradmwk, 0.8)))) %>%
    mutate(dsli_orig = (dsli*sd(downstream$dsli)+mean(downstream$dsli))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.28_srad <- predict_gam(gamd1.28, values = list(Sradmwk = c(quantile(d.idys$Sradmwk, 0.2),
                                                          quantile(d.idys$Sradmwk, 0.8), 
                                                          median(d.idys$Sradmwk)),
                                             log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8)))) %>%
   mutate(dsli_orig = (dsli*sd(downstream$dsli)+mean(downstream$dsli))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.28_dsli <- predict_gam(gamd1.28, values = list(dsli = c(-0.873, -0.021, 1.992
),
                                             log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8)))) %>%
   mutate(dsli_orig = (dsli*sd(downstream$dsli)+mean(downstream$dsli))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.28_srad_dsli <- predict_gam(gamd1.28, values = list(dsli = c(-0.873, -0.021, 1.992
),
                                             Sradmwk = c(quantile(d.idys$Sradmwk, 0.2),
                                                          quantile(d.idys$Sradmwk, 0.8)))) %>%
   mutate(dsli_orig = (dsli*sd(downstream$dsli)+mean(downstream$dsli))) %>%
  mutate(across(.cols = everything(), round, digits = 3))


pred_gamd1.28_all <- predict_gam(gamd1.28) %>%
   mutate(dsli_orig = (dsli*sd(downstream$dsli)+mean(downstream$dsli))) %>%
  mutate(across(.cols = everything(), round, digits = 3))


```

Plot predictions
```{r}
# by Srad and dsli
pred_gamd1.28_srad_dsli %>%
  ggplot(aes(log_qsdy, fit)) +
  geom_smooth_ci(dsli_orig, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~Sradmwk,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()

# difference by Srad and flow
pred_gamd1.28_srad %>%
  ggplot(aes(dsli_orig, fit)) +
  geom_smooth_ci(Sradmwk, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  labs(y = "chlorophyll")+
  theme_bw()

# difference by days since last inundation and flow
pred_gamd1.28_dsli %>%
  ggplot(aes(Sradmwk, fit)) +
  geom_smooth_ci(dsli_orig, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = "chlorophyll")+
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  theme_bw()

# difference by days since last inundation and flow
pred_gamd1.28_dsli %>%
  ggplot(aes(Sradmwk, fit)) +
  geom_smooth_ci(log_qsdy, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = "chlorophyll")+
  facet_wrap(~dsli_orig,
             labeller = label_both) +
  theme_bw()
```


#### Visualize gamd1.26
gamd1.26 <- gam(log_chla ~ te(log_qsdy, WTmwk, dowy, bs = c("tp", "tp", "cc")), method = "REML", data = d.idys, family = "gaussian")

Predictions 
```{r}
library(tidymv)

pred_gamd1.26_q <- predict_gam(gamd1.26, values = list(log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8), 
                                                          median(d.idys$log_qsdy)),
                                             WTmwk = c(quantile(d.idys$WTmwk, 0.2),
                                                       quantile(d.idys$WTmwk, 0.8)))) %>%
  mutate(WTmwk_orig = (WTmwk*sd(downstream$WTmwk)+mean(downstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.26_wt <- predict_gam(gamd1.26, values = list(WTmwk = c(quantile(d.idys$WTmwk, 0.2),
                                                          quantile(d.idys$WTmwk, 0.8), 
                                                          median(d.idys$WTmwk)),
                                             log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8)))) %>%
   mutate(WTmwk_orig = (WTmwk*sd(downstream$WTmwk)+mean(downstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.26_dowy <- predict_gam(gamd1.26, values = list(dowy = c(quantile(d.idys$dowy, 0.2),
                                                          quantile(d.idys$dowy, 0.8), 
                                                          median(d.idys$dowy)),
                                             log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8)))) %>%
    mutate(WTmwk_orig = (WTmwk*sd(downstream$WTmwk)+mean(downstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.26_dowy_qsdy <- predict_gam(gamd1.26, values = list(dowy = c(quantile(d.idys$dowy, 0.1),
                                                          quantile(d.idys$dowy, 0.3),
                                                          quantile(d.idys$dowy, 0.5), 
                                                          quantile(d.idys$dowy, 0.7), 
                                                          quantile(d.idys$dowy, 0.9)),
                                             log_qsdy = c(quantile(d.idys$log_qsdy, 0.2),
                                                          quantile(d.idys$log_qsdy, 0.8), 
                                                          median(d.idys$log_qsdy)))) %>%
  mutate(WTmwk_orig = (WTmwk*sd(downstream$WTmwk)+mean(downstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))

pred_gamd1.26_all <- predict_gam(gamd1.26) %>%
   mutate(WTmwk_orig = (WTmwk*sd(downstream$WTmwk)+mean(downstream$WTmwk))) %>%
  mutate(across(.cols = everything(), round, digits = 3))
```

Higher chlorophyll at higher temps. Increase in chlorophyll from 100-180 (Spring)
```{r}
pred_gamd1.26_q %>%
  ggplot(aes(dowy, fit)) +
  geom_smooth_ci(log_qsdy, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~WTmwk_orig,
             labeller= label_both) +
  labs(y = "chlorophyll") +
  theme_bw()
```

Higher chlorophyll at higher flow during higher temps
```{r}
pred_gamd1.26_wt %>%
  ggplot(aes(dowy, fit)) +
  geom_smooth_ci(WTmwk_orig, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  labs(y = "chlorophyll")+
  theme_bw()
```

Increase in chlorophyll with higher temps. No difference by dowy. 
```{r}
pred_gamd1.26_dowy %>%
  ggplot(aes(WTmwk_orig, fit)) +
  geom_smooth_ci(dowy, size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = "chlorophyll")+
  facet_wrap(~log_qsdy,
             labeller = label_both) +
  theme_bw()
```

```{r}
pred_gamd1.26_all %>%
  ggplot(aes(WTmwk_orig, log_qsdy, z = fit)) +
  geom_tile(aes(fill = fit)) +
  geom_contour(colour = "white") +
  scale_fill_viridis(name = "chla") +
  theme_minimal() +
  theme(legend.position = "top")
```


#### Visualize gamd1.2
```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.95
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(d.idys$log_qsdy))
#med.dowy = median(d.idys$dowy)
med.dowy = quantile(d.idys$dowy, probs = 0.9)
qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
n = length(qsdy.ser)
new_dat_d1 = data.frame(cbind(log_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_d1$inundation = sample(c(0,1), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gamd1.2, newdata = new_dat_d1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_d1 = cbind(new_dat_d1, fit, upper, lower)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# separate by inundation flag
gamd.i0 = subset(new_dat_d1, new_dat_d1$inundation == 0)
gamd.i1 = subset(new_dat_d1, new_dat_d1$inundation == 1)

# plot the actual data points
plot(d.idys$log_chla ~ d.idys$log_qsdy, pch = 20, col = gray(0.1,0.2), las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamd.i0$log_qsdy, max(gamd.i0$log_qsdy), rev(gamd.i0$log_qsdy), gamd.i0$log_qsdy[1]),
        y = c(gamd.i0$lower, gamdd.i0$upper[length(gamd.i0$upper)], rev(gamd.i0$upper), gamd.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i0$log_qsdy, y = gamd.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamd.i1$log_qsdy, max(gamd.i1$log_qsdy), rev(gamd.i1$log_qsdy), gamd.i1$log_qsdy[1]),
        y = c(gamd.i1$lower, gamd.i1$upper[length(gamd.i1$upper)], rev(gamd.i1$upper), gamd.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i1$log_qsdy, y = gamd.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

```

#### Visualize gamd1.3

```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.95
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(d.idys$log_idys))
med.dowy = median(d.idys$dowy)
qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
n = length(qsdy.ser)
new_dat_d1 = data.frame(cbind(log_idys = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_d1$inundation = sample(c(0,1), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gamd1.3, newdata = new_dat_d1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_d1 = cbind(new_dat_d1, fit, upper, lower)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# separate by inundation flag
gamd.i0 = subset(new_dat_d1, new_dat_d1$inundation == 0)
gamd.i1 = subset(new_dat_d1, new_dat_d1$inundation == 1)

# plot the actual data points
plot(d.idys$log_chla ~ d.idys$log_idys, pch = 20, col = gray(0.1,0.2), las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Inundation Days)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamd.i0$log_idys, max(gamd.i0$log_idys), rev(gamd.i0$log_idys), gamd.i0$log_idys[1]),
        y = c(gamd.i0$lower, gamd.i0$upper[length(gamd.i0$upper)], rev(gamd.i0$upper), gamd.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i0$log_idys, y = gamd.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamd.i1$log_idys, max(gamd.i1$log_idys), rev(gamd.i1$log_idys), gamd.i1$log_idys[1]),
        y = c(gamd.i1$lower, gamd.i1$upper[length(gamd.i1$upper)], rev(gamd.i1$upper), gamd.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i1$log_idys, y = gamd.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

```



#### Visualize gamd1.4

```{r}

# gamd1.4 <- gam(log_chla ~ te(log_qsdy, dowy, by = inund_fac), method = "REML", data = d.idys, family = "gaussian")


# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(d.idys$log_qsdy))
med.dowy = median(d.idys$dowy)
qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
n = length(qsdy.ser)
new_dat_d1 = data.frame(cbind(log_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_d1$inund_fac = sample(c("none", "long", "short"), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gamd1.4, newdata = new_dat_d1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_d1 = cbind(new_dat_d1, fit, upper, lower)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# separate by inundation flag
gamd.i0 = subset(new_dat_d1, new_dat_d1$inund_fac == "none")
gamd.i1 = subset(new_dat_d1, new_dat_d1$inund_fac == "long")

# plot the actual data points
plot(d.idys$log_chla ~ d.idys$log_qsdy, pch = 20, col = gray(0.1,0.2), las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow same day)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamd.i0$log_qsdy, max(gamd.i0$log_qsdy), rev(gamd.i0$log_qsdy), gamd.i0$log_qsdy[1]),
        y = c(gamd.i0$lower, gamd.i0$upper[length(gamd.i0$upper)], rev(gamd.i0$upper), gamd.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i0$log_qsdy, y = gamd.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamd.i1$log_qsdy, max(gamd.i1$log_qsdy), rev(gamd.i1$log_qsdy), gamd.i1$log_qsdy[1]),
        y = c(gamd.i1$lower, gamd.i1$upper[length(gamd.i1$upper)], rev(gamd.i1$upper), gamd.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamd.i1$log_qsdy, y = gamd.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

```



## Yolo

### Filter to yolo

```{r}
yolo <- alldata %>% filter(region == "yolo")

yolo_scaled <- yolo %>%
  mutate(across(.cols = c(log_idys,log_qsdy,tily:Sradmwk), ~ scale(.x)))
  
yo.idys <- yolo_scaled
```

### Model

```{r}
gamyo0 <- gam(log_chla ~ 1, method = "REML", data = yo.idys, family = "gaussian")
gamyo1.0 <- gam(log_chla ~ s(log_qsdy, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.1 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.2 <- gam(log_chla ~ te(log_qsdy, dowy, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.3 <- gam(log_chla ~ te(log_idys, dowy, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.4 <- gam(log_chla ~ te(log_qsdy, dowy, by = inund_fac),  method = "REML", data = yo.idys, family = "gaussian")
gamyo1.5 <- gam(log_chla ~ s(WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.6 <- gam(log_chla ~ s(WTmwk, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.7 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(WTmwk, bs = "cc"), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.8 <- gam(log_chla ~ s(log_qsdy, by = inundation) + s(Sradmwk, bs = "cc"), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.9 <- gam(log_chla ~ te(log_qsdy, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.10 <- gam(log_chla ~ te(log_qsdy, Sradmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.11 <- gam(log_chla ~ te(log_idys, Sradmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.12 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.13 <- gam(log_chla ~ te(log_qsdy, dowy), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.14 <- gam(log_chla ~ te(log_idys, dowy), method = "REML", data = yo.idys, family = "gaussian")

gamyo1.15 <- gam(log_chla ~ s(tily, k = 6), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.16 <- gam(log_chla ~ s(dsli), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.17 <- gam(log_chla ~ s(diurnal_range), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.18 <- gam(log_chla ~ s(WTrangemwk), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.19 <- gam(log_chla ~ s(log_qsdy), method = "REML", data = yo.idys, family = "gaussian")

gamyo1.20 <- gam(log_chla ~ te(log_qsdy, tily, dowy), method = "REML", data = yo.idys, family = "gaussian") # BEST MODEL

gamyo1.21 <- gam(log_chla ~ te(log_qsdy, dsli, dowy), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.22 <- gam(log_chla ~ te(log_qsdy, diurnal_range, dowy), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.23 <- gam(log_chla ~ te(log_qsdy, WTrangemwk, dowy), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.24 <- gam(log_chla ~ te(log_qsdy, WTmwk, dowy), method = "REML", data = yo.idys, family = "gaussian")
gamyo1.25 <- gam(log_chla ~ te(log_qsdy, dowy, by = tily_fac), method = "REML", data = yo.idys, family = "gaussian")

# AIC
res <- (AIC(gamyo0, gamyo1.0, gamyo1.1, gamyo1.2, gamyo1.3, gamyo1.4, gamyo1.5, gamyo1.6, gamyo1.7, gamyo1.8, gamyo1.9, gamyo1.10, gamyo1.11, gamyo1.12, gamyo1.13, gamyo1.14, gamyo1.15, gamyo1.16, gamyo1.17, gamyo1.18, gamyo1.19, gamyo1.20, gamyo1.21, gamyo1.22, gamyo1.23, gamyo1.24, gamyo1.25))
res %>% arrange(AIC)

write.csv(res,"CompRegYolo.csv")

summary(gamyo1.2) # 64.3% R2 0.603
summary(gamyo1.13) # 61.2% R2 0.578
summary(gamyo1.20) # 72% R2 0.66 BEST ONE
summary(gamyo1.21) # 69.5% R2 0.635 

par(mfrow = c(2,2))
gam.check(gamyo1.1)
gam.check(gamyo1.2)
gam.check(gamyo1.3)
gam.check(gamyo1.4)
gam.check(gamyo1.20)

summary(gamyo1.19)
summary(gamyo1.22)
summary(gamyo1.23)
summary(gamyo1.24)

resid = residuals(gamyo1.20)
acf(resid)
pacf(resid)
```

#### Model validation for the gam model based on hypothesis testing

```{r}

  # validate model gamyo1.20 - log_chla ~ te(log_qsdy, dowy, tily)
  # -----------------------------------------------------------------------
  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_val_qsdy_dowy_tily_yolo.tiff", width = 6, height = 9, units = "in", res = 200)
  resid = residuals(gamyo1.20)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamyo1.20), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(yo.idys$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(yo.idys$dowy, resid, xlab = "day of water year",   ylab = "Residuals")
  plot(yo.idys$tily, resid, xlab = "total inundation last year",   ylab = "Residuals")
  par(op)
  dev.off()

```

#### Visualize gamyo1.20 across day of water year

* Low Flow (10 percentile), High Flow (90 percentile)
* Inundation last year to 5 and 95 percentile
* Visualize dowy at 5-95 percentile


```{r}
# gamyo1.20 <- gam(log_chla ~ te(log_qsdy, tily, dowy), method = "REML", data = yo.idys, family = "gaussian") # BEST MODEL


# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# series of 5-95 percentile values of the predictor variable: day of water year
dowy.1 =   floor(quantile(yo.idys$dowy, probs = 0.05))
dowy.9 = ceiling(quantile(yo.idys$dowy, probs = 0.95))
dowy.ser = seq(from = dowy.1, to = dowy.9, by = 1)
n = length(dowy.ser)

# take 10th and 90th percentile values of all other predictors
lqsdy.1 = quantile(yo.idys$log_qsdy, probs = 0.1)
lqsdy.9 = quantile(yo.idys$log_qsdy, probs = 0.9)

tily.1 = quantile(yo.idys$tily, probs = 0.1)
tily.9 = quantile(yo.idys$tily, probs = 0.9)

# dataset for log flow at 10 percentile; tily at 10 percentile
yo_dat_1 = data.frame(cbind(log_qsdy = rep(lqsdy.1, n), tily = rep(tily.1, n), dowy = dowy.ser))

# dataset for log flow at 90 percentile; tily at 10 percentile
yo_dat_2 = data.frame(cbind(log_qsdy = rep(lqsdy.9, n), tily = rep(tily.1, n), dowy = dowy.ser))

# dataset for log flow at 10 percentile; tily at 90 percentile
yo_dat_3 = data.frame(cbind(log_qsdy = rep(lqsdy.1, n), tily = rep(tily.9, n), dowy = dowy.ser))

# dataset for log flow at 90 percentile; tily at 90 percentile
yo_dat_4 = data.frame(cbind(log_qsdy = rep(lqsdy.9, n), tily = rep(tily.9, n), dowy = dowy.ser))

yo_dat = rbind(yo_dat_1, yo_dat_2, yo_dat_3, yo_dat_4)

#~ Use the *predict()* function to estimate model fit and standard error
preds.yo = predict(gamyo1.20, newdata = yo_dat, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
fit = preds.yo$fit
upper = fit - CI_factor*preds.yo$se.fit
lower = fit + CI_factor*preds.yo$se.fit
yo_dat = cbind(yo_dat, fit, upper, lower)

# separate by log_qsdy at high and low value & tily at high and low value
yo.lq.ltily = subset(yo_dat, yo_dat$log_qsdy == lqsdy.1 & yo_dat$tily == tily.1)
yo.hq.ltily = subset(yo_dat, yo_dat$log_qsdy == lqsdy.9 & yo_dat$tily == tily.1)
yo.hq.htily = subset(yo_dat, yo_dat$log_qsdy == lqsdy.1 & yo_dat$tily == tily.9)
yo.lq.htily = subset(yo_dat, yo_dat$log_qsdy == lqsdy.9 & yo_dat$tily == tily.9)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for low and high flow for day of water year
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

tiff(filename = "figures/gam_yolo_vis_lqsdy_dowy_tily.tiff", width = 10, height = 5, units = "in", res = 200)

# Plot the raw data and make the axes and labels aesthetically pleasing (margins = b,l,t,r)
par(mfrow = c(1,2), mar = c(2.5, 4.5, 2, 1) + 0.1, oma = c(1.5, 0, 0.5, 0))

# plot the actual data points separately for inundated and non-inundated
plot(yo.idys$log_chla ~ yo.idys$dowy, pch = 20, col = "green", las = 1, ylab = "", xlab = "", main = "Low inundation last year")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Day of water year", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(yo.lq.ltily$dowy, max(yo.lq.ltily$dowy), rev(yo.lq.ltily$dowy), yo.lq.ltily$dowy[1]),
        y = c(yo.lq.ltily$lower, yo.lq.ltily$upper[length(yo.lq.ltily$upper)], rev(yo.lq.ltily$upper), yo.lq.ltily$lower[1]), border = NA,
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = yo.lq.ltily$dowy, y = yo.lq.ltily$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(yo.hq.ltily$dowy, max(yo.hq.ltily$dowy), rev(yo.hq.ltily$dowy), yo.hq.ltily$dowy[1]),
        y = c(yo.hq.ltily$lower, yo.hq.ltily$upper[length(yo.hq.ltily$upper)], rev(yo.hq.ltily$upper), yo.hq.ltily$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = yo.hq.ltily$dowy, y = yo.hq.ltily$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))


# plot the actual data points separately for inundated and non-inundated
plot(yo.idys$log_chla ~ yo.idys$dowy, pch = 20, col = "green", las = 1, ylab = "", xlab = "", main = "High inundation last year")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Day of water year", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(yo.hq.htily$dowy, max(yo.hq.htily$dowy), rev(yo.hq.htily$dowy), yo.hq.htily$dowy[1]),
        y = c(yo.hq.htily$lower, yo.hq.htily$upper[length(yo.hq.htily$upper)], rev(yo.hq.htily$upper), yo.hq.htily$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = yo.hq.htily$dowy, y = yo.hq.htily$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(yo.lq.htily$dowy, max(yo.lq.htily$dowy), rev(yo.lq.htily$dowy), yo.lq.htily$dowy[1]),
        y = c(yo.lq.htily$lower, yo.lq.htily$upper[length(yo.lq.htily$upper)], rev(yo.lq.htily$upper), yo.lq.htily$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = yo.lq.htily$dowy, y = yo.lq.htily$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

dev.off()

```

## Trying other ways to visulize two covars at once

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for low and high water tily for day of water year
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
library("plot3D")
library("ggplot2")

# surf3D(new.dat$log_qsdy, new.dat$WTmwk, new.dat$log_chla, colvar = new.dat$log_chla,
#        colkey = TRUE, box = TRUE, bty = "b", phi = 20, theta = 120)

#install.packages("visreg")
library("visreg")
par(new=TRUE)
visreg2d(gamyo1.20, xvar='log_qsdy', yvar='total_inund_last_year', scale='response', plot.type="image")

# Figure of flow vs dowy with points
visreg2d(gamyo1.20, xvar='dowy', yvar='log_qsdy', scale='response', plot.type="image")
par(new=TRUE)
plot(yo.idys$dowy, yo.idys$log_qsdy, xlim = c(75, 240), cex = yo.idys$WTmwk+2, xaxt = "n", yaxt = "n", ylab =)

# Figure of tily vs dowy with points
visreg2d(gamyo1.20, xvar='dowy', yvar='total_inund_last_year', scale='response', plot.type="image")
par(new=TRUE)
plot(yo.idys$dowy, yo.idys$total_inund_last_year, xlim = c(75, 240), cex = yo.idys$WTmwk+2, xaxt = "n", yaxt = "n", ylab = "")

visreg2d(gamyo1.20, xvar='log_qsdy', yvar='dowy', scale='response', plot.type="persp")

tempmeans <- alldata %>% group_by(dowy, region) %>%
  summarize(meanTemp = mean(WTmwk))

```


### Visualize gamyo1.20 across flow

## not working for some reason - troubleshoot later!!

* Across all flows
* Inundation last year for 10 and 90 percentile
* day of water year at 0.05, 0.5 and 0.95 percentile


```{r}
# gamyo1.20 <- gam(log_chla ~ te(log_qsdy, tily, dowy), method = "REML", data = yo.idys, family = "gaussian") # BEST MODEL

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

# day of water year = 130 when there is max. difference between the two types
dowy.05 = quantile(yo.idys$log_qsdy, probs = 0.05)
dowy.50 = quantile(yo.idys$log_qsdy, probs = 0.50)
dowy.95 = quantile(yo.idys$log_qsdy, probs = 0.95)

# take 10th and 90th percentile values of all other predictors
lqsdy.1 = quantile(yo.idys$log_qsdy, probs = 0.05)
lqsdy.9 = quantile(yo.idys$log_qsdy, probs = 0.95)
lqsdy.ser = seq(from = lqsdy.1, to = lqsdy.9, by = 0.01)
n = length(lqsdy.ser)

tily.1 = quantile(yo.idys$tily, probs = 0.1)
tily.9 = quantile(yo.idys$tily, probs = 0.9)

# dataset for log flow at 10 percentile; tily at 10 percentile
yo_dat_1 = data.frame(cbind(log_qsdy = lqsdy.ser, tily = tily.1, dowy = dowy.95))

# dataset for log flow at 90 percentile; tily at 10 percentile
yo_dat_2 = data.frame(cbind(log_qsdy = lqsdy.ser, tily = tily.9, dowy = dowy.95))

yo_dat = rbind(yo_dat_1, yo_dat_2)

#~ Use the *predict()* function to estimate model fit and standard error
preds.yo = predict(gamyo1.20, newdata = yo_dat, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
fit = preds.yo$fit
upper = fit - CI_factor*preds.yo$se.fit
lower = fit + CI_factor*preds.yo$se.fit
yo_dat = cbind(yo_dat, fit, upper, lower)

# separate by log_qsdy at high and low value & tily at high and low value
yo.ltily = subset(yo_dat, yo_dat$tily == tily.1)
yo.htily = subset(yo_dat, yo_dat$tily == tily.9)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for low and high flow for day of water year
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

tiff(filename = "figures/gam_yolo_vis_lqsdy_dowy95_tily.tiff", width = 6, height = 8, units = "in", res = 200)

# Plot the raw data and make the axes and labels aesthetically pleasing (margins = b,l,t,r)
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2, 1) + 0.1, oma = c(1.5, 0, 0.5, 0))

# plot the actual data points separately for inundated and non-inundated
plot(yo.idys$log_chla ~ yo.idys$log_qsdy, pch = 20, col = "green", las = 1, ylab = "", xlab = "", main = "95 percentile Day-of-water-year")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow)"), side = 1, line = 2.5)

# Use the polygon function to add the CI
polygon(x = c(yo.ltily$log_qsdy, max(yo.ltily$log_qsdy), rev(yo.ltily$log_qsdy), yo.ltily$log_qsdy[1]),
        y = c(yo.ltily$lower, yo.ltily$upper[length(yo.ltily$upper)], rev(yo.ltily$upper), yo.ltily$lower[1]), border = NA,
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = yo.ltily$log_qsdy, y = yo.ltily$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(yo.htily$log_qsdy, max(yo.htily$log_qsdy), rev(yo.htily$log_qsdy), yo.htily$log_qsdy[1]),
        y = c(yo.htily$lower, yo.htily$upper[length(yo.htily$upper)], rev(yo.htily$upper), yo.htily$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = yo.htily$log_qsdy, y = yo.htily$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

dev.off()

```



#### Visualize gamyo1.2

```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)
#~ Generate a data frame with the sequence from the minimum to the mayo.imum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(yo.idys$log_qsdy))
lqsdy.05 = quantile(yo.idys$log_qsdy, probs = 0.05)
lqsdy.95 = quantile(yo.idys$log_qsdy, probs = 0.95)
#med.dowy = median(yo.idys$dowy)
med.dowy = quantile(yo.idys$dowy, probs = 0.5)
#qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
qsdy.ser = seq(from = lqsdy.05, to = lqsdy.95, by = 0.02)
n = length(qsdy.ser)
new_dat_yo1 = data.frame(cbind(log_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_yo1$inundation = sample(c(0,1), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gam1.2, newdata = new_dat_yo1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_yo1 = cbind(new_dat_yo1, fit, upper, lower)

# separate by inundation flag
gamyo.i0 = subset(new_dat_yo1, new_dat_yo1$inundation == 0)
gamyo.i1 = subset(new_dat_yo1, new_dat_yo1$inundation == 1)

# separate actual data by inundation flag
act.i0 = subset(yo.idys, yo.idys$inundation == 0)
act.i1 = subset(yo.idys, yo.idys$inundation == 1)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# plot the actual data points separately for inundated and non-inundated
colors = c("green", "blue")
plot(yo.idys$log_chla ~ yo.idys$log_qsdy, pch = 20, col = colors[yo.idys$inundation], las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamyo.i0$log_qsdy, max(gamyo.i0$log_qsdy), rev(gamyo.i0$log_qsdy), gamyo.i0$log_qsdy[1]),
        y = c(gamyo.i0$lower, gamyo.i0$upper[length(gamyo.i0$upper)], rev(gamyo.i0$upper), gamyo.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i0$log_qsdy, y = gamyo.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamyo.i1$log_qsdy, max(gamyo.i1$log_qsdy), rev(gamyo.i1$log_qsdy), gamyo.i1$log_qsdy[1]),
        y = c(gamyo.i1$lower, gamyo.i1$upper[length(gamyo.i1$upper)], rev(gamyo.i1$upper), gamyo.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i1$log_qsdy, y = gamyo.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))


```



#### Visualize gamyo1.6

```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(yo.idys$log_qsdy))
lqsdy.05 = quantile(yo.idys$log_qsdy, probs = 0.05)
lqsdy.95 = quantile(yo.idys$log_qsdy, probs = 0.95)
#med.dowy = median(yo.idys$dowy)
med.dowy = quantile(yo.idys$dowy, probs = 0.5)
#qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
qsdy.ser = seq(from = lqsdy.05, to = lqsdy.95, by = 0.02)
n = length(qsdy.ser)
new_dat_yo1 = data.frame(cbind(log_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))

#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gam1.6, newdata = new_dat_yo1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_yo1 = cbind(new_dat_yo1, fit, upper, lower)

# separate actual data by inundation flag
act.i0 = subset(yo.idys, yo.idys$inundation == 0)
act.i1 = subset(yo.idys, yo.idys$inundation == 1)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# plot the actual data points separately for inundated and non-inundated
colors = c("green", "blue")
plot(yo.idys$log_chla ~ yo.idys$log_qsdy, pch = 20, col = colors[yo.idys$inundation], las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(new_dat_yo1$log_qsdy, max(new_dat_yo1$log_qsdy), rev(new_dat_yo1$log_qsdy), new_dat_yo1$log_qsdy[1]),
        y = c(new_dat_yo1$lower, new_dat_yo1$upper[length(gamyo.i0$upper)], rev(new_dat_yo1$upper), new_dat_yo1$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = new_dat_yo1$log_qsdy, y = new_dat_yo1$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

```


#### Visualize gamyo1.3

```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.95
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(yo.idys$log_idys))
med.dowy = median(yo.idys$dowy)
qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
n = length(qsdy.ser)
new_dat_yo1 = data.frame(cbind(log_idys = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_yo1$inundation = sample(c(0,1), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gam1.3, newdata = new_dat_yo1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_yo1 = cbind(new_dat_yo1, fit, upper, lower)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# separate by inundation flag
gamyo.i0 = subset(new_dat_yo1, new_dat_yo1$inundation == 0)
gamyo.i1 = subset(new_dat_yo1, new_dat_yo1$inundation == 1)

# plot the actual data points
plot(yo.idys$log_chla ~ yo.idys$log_idys, pch = 20, col = gray(0.1,0.2), las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Inundation Days)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamyo.i0$log_idys, max(gamyo.i0$log_idys), rev(gamyo.i0$log_idys), gamyo.i0$log_idys[1]),
        y = c(gamyo.i0$lower, gamyo.i0$upper[length(gamyo.i0$upper)], rev(gamyo.i0$upper), gamyo.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i0$log_idys, y = gamyo.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamyo.i1$log_idys, max(gamyo.i1$log_idys), rev(gamyo.i1$log_idys), gamyo.i1$log_idys[1]),
        y = c(gamyo.i1$lower, gamyo.i1$upper[length(gamyo.i1$upper)], rev(gamyo.i1$upper), gamyo.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i1$log_idys, y = gamyo.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

```


#### Visualize gamyo1.4

```{r}

# gam1.4 <- gam(log_chla ~ te(log_qsdy, dowy, by = inund_fac), method = "REML", data = yo.idys, family = "gaussian")


# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(yo.idys$log_qsdy))
med.dowy = median(yo.idys$dowy)
qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
n = length(qsdy.ser)
new_dat_yo1 = data.frame(cbind(log_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_yo1$inund_fac = sample(c("none", "long", "short"), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gam1.4, newdata = new_dat_yo1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_yo1 = cbind(new_dat_yo1, fit, upper, lower)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# separate by inundation flag
gamyo.i0 = subset(new_dat_yo1, new_dat_yo1$inund_fac == "none")
gamyo.i1 = subset(new_dat_yo1, new_dat_yo1$inund_fac == "long")

# plot the actual data points
plot(yo.idys$log_chla ~ yo.idys$log_qsdy, pch = 20, col = gray(0.1,0.2), las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow same day)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gamyo.i0$log_qsdy, max(gamyo.i0$log_qsdy), rev(gamyo.i0$log_qsdy), gamyo.i0$log_qsdy[1]),
        y = c(gamyo.i0$lower, gamyo.i0$upper[length(gamyo.i0$upper)], rev(gamyo.i0$upper), gamyo.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i0$log_qsdy, y = gamyo.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gamyo.i1$log_qsdy, max(gamyo.i1$log_qsdy), rev(gamyo.i1$log_qsdy), gamyo.i1$log_qsdy[1]),
        y = c(gamyo.i1$lower, gamyo.i1$upper[length(gamyo.i1$upper)], rev(gamyo.i1$upper), gamyo.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gamyo.i1$log_qsdy, y = gamyo.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

```

## Upstream

```{r}
upstream <- alldata %>% filter(region == "above")
```

### Filter to upstream
```{r}
upstream_scaled <- upstream %>%
  mutate(across(.cols = c(diurnal_range:Sradmwk, t_qsdy:log_qsdy), ~ scale(.x)))
  
u.idys <- upstream_scaled

upstream_s_long <- pivot_longer(u.idys, cols = c(diurnal_range:Sradmwk, log_idys:log_qsdy), names_to = "Parameter", values_to = "Value")

ggplot(upstream_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```

### Model 
Need different models here, not inundation (maybe season or precip?)
```{r}
gamu0 <- gam(log_chla ~ 1, method = "REML", data = u.idys, family = "gaussian")
gamu1.0 <- gam(log_chla ~ s(log_qsdy), method = "REML", data = u.idys, family = "gaussian")
gamu1.1 <- gam(log_chla ~ s(log_qsdy) + s(dowy, bs = "cc"), method = "REML", data = u.idys, family = "gaussian")
gamu1.2 <- gam(log_chla ~ s(WTmwk), method = "REML", data = u.idys, family = "gaussian")
gamu1.3 <- gam(log_chla ~ s(WTmwk) + s(dowy, bs = "cc"), method = "REML", data = u.idys, family = "gaussian")
gamu1.4 <- gam(log_chla ~ s(Sradmwk), method = "REML", data = u.idys, family = "gaussian")
gamu1.5 <- gam(log_chla ~ s(Sradmwk) + s(dowy, bs = "cc"), method = "REML", data = u.idys, family = "gaussian")
gamu1.6 <- gam(log_chla ~ s(WTrangemwk), method = "REML", data = u.idys, family = "gaussian")
gamu1.7 <- gam(log_chla ~ s(WTrangemwk) + s(dowy, bs = "cc"), method = "REML", data = u.idys, family = "gaussian")
gamu1.8 <- gam(log_chla ~ te(WTmwk, dowy), method = "REML", data = u.idys, family = "gaussian")
gamu1.9 <- gam(log_chla ~ te(WTmwk, log_qsdy, dowy), method = "REML", data = u.idys, family = "gaussian")
gamu1.10 <- gam(log_chla ~ te(WTmwk, Sradmwk, dowy), method = "REML", data = u.idys, family = "gaussian")
gamu1.11 <- gam(log_chla ~ te(WTmwk, WTrangemwk, dowy), method = "REML", data = u.idys, family = "gaussian")
gamu1.12 <- gam(log_chla ~ te(WTmwk, diurnal_range, dowy), method = "REML", data = u.idys, family = "gaussian")
gamu1.13 <- gam(log_chla ~ s(diurnal_range), method = "REML", data = u.idys, family = "gaussian")
gamu1.14 <- gam(log_chla ~ s(diurnal_range) + s(dowy, bs = "cc"), method = "REML", data = u.idys, family = "gaussian")

# AIC
resu <- AIC(gamu0, gamu1.0, gamu1.1, gamu1.2, gamu1.3, gamu1.4, gamu1.5, gamu1.6, gamu1.7, gamu1.8, gamu1.9, gamu1.10, gamu1.11, gamu1.12, gamu1.13, gamu1.14)
resu %>% arrange(AIC)

write.csv(resu, "CompRegUpstream.csv")

summary(gamu1.2)
summary(gamu1.3)
summary(gamu1.11)
summary(gamu1.12)

par(mfrow = c(2,2))
gam.check(gamu1.2)
gam.check(gamu1.8)
gam.check(gamu1.7)
gam.check(gamu1.10)

summary(gamu1.7) # 38.1% R2 0.331
summary(gamu1.8) # 33.7% R2 0.301
summary(gamu1.2) # 28.97% R2 0.271
summary(gamu1.10) # 30.5% R2 0.29


resid = residuals(gamu1.10)
acf(resid)
pacf(resid)
```


## upstream best model by AIC
## Visualizing gamu1.11 (water temp. + its range) for Sac River across WTemp

```{r}

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# series of 5-95 percentile values of the predictor variable: WTmwk
wt.1 = quantile(u.idys$WTmwk, probs = 0.05)
wt.9 = quantile(u.idys$WTmwk, probs = 0.95)
wt.ser = seq(from = wt.1, to = wt.9, by = 0.02)
n = length(wt.ser)

# take 10th and 90th percentile values of all other predictors
wtrange.1 = quantile(u.idys$WTrangemwk, probs = 0.1)
wtrange.9 = quantile(u.idys$WTrangemwk, probs = 0.9)

# day of water year
dowy.fxd = quantile(u.idys$dowy, probs = 0.5)

new_dat_1 = data.frame(cbind(WTrangemwk = wtrange.1, WTmwk = wt.ser, dowy = dowy.fxd))
new_dat_2 = data.frame(cbind(WTrangemwk = wtrange.9, WTmwk = wt.ser, dowy = dowy.fxd))

new_dat = rbind(new_dat_1, new_dat_2)

#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gamu1.11, newdata = new_dat, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit
new_dat = cbind(new_dat, fit, upper, lower)

# separate by log_qsdy at high and low value
gam.i0 = subset(new_dat, new_dat$WTrangemwk == wtrange.1)
gam.i1 = subset(new_dat, new_dat$WTrangemwk == wtrange.9)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for high and low water temp. range
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# plot the actual data points separately for inundated and non-inundated
plot(u.idys$log_chla ~ u.idys$WTmwk, pch = 20, col = "green", las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Water temperature", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gam.i0$WTmwk, max(gam.i0$WTmwk), rev(gam.i0$WTmwk), gam.i0$WTmwk[1]),
        y = c(gam.i0$lower, gam.i0$upper[length(gam.i0$upper)], rev(gam.i0$upper), gam.i0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = gam.i0$WTmwk, y = gam.i0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gam.i1$WTmwk, max(gam.i1$WTmwk), rev(gam.i1$WTmwk), gam.i1$WTmwk[1]),
        y = c(gam.i1$lower, gam.i1$upper[length(gam.i1$upper)], rev(gam.i1$upper), gam.i1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gam.i1$WTmwk, y = gam.i1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))
```


#### Visualize gamu1.12 across day of water year

* Low and high water temperature
* Low and high tidal range
* Visualize dowy at 5-95 percentile


```{r}
# gamu1.12 <- gam(log_chla ~ te(WTmwk, diurnal_range, dowy), method = "REML", data = u.idys, family = "gaussian")  BEST MODEL

# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# series of 5-95 percentile values of the predictor variable: day of water year
dowy.1 =   floor(quantile(u.idys$dowy, probs = 0.05))
dowy.9 = ceiling(quantile(u.idys$dowy, probs = 0.95))
dowy.ser = seq(from = dowy.1, to = dowy.9, by = 1)
n = length(dowy.ser)

# take 10th and 90th percentile values of all other predictors
dr.1 = quantile(u.idys$diurnal_range, probs = 0.1)
dr.9 = quantile(u.idys$diurnal_range, probs = 0.9)

WTmwk.1 = quantile(u.idys$WTmwk, probs = 0.1)
WTmwk.9 = quantile(u.idys$WTmwk, probs = 0.9)

# dataset for log flow at 10 percentile; tily at 10 percentile
yo_dat_1 = data.frame(cbind(diurnal_range = rep(dr.1, n), WTmwk = rep(WTmwk.1, n), dowy = dowy.ser))

# dataset for log flow at 90 percentile; tily at 10 percentile
yo_dat_2 = data.frame(cbind(diurnal_range = rep(dr.9, n), WTmwk = rep(WTmwk.1, n), dowy = dowy.ser))

# dataset for log flow at 10 percentile; tily at 90 percentile
yo_dat_3 = data.frame(cbind(diurnal_range = rep(dr.1, n), WTmwk = rep(WTmwk.9, n), dowy = dowy.ser))

# dataset for log flow at 90 percentile; tily at 90 percentile
yo_dat_4 = data.frame(cbind(diurnal_range = rep(dr.9, n), WTmwk = rep(WTmwk.9, n), dowy = dowy.ser))

yo_dat = rbind(yo_dat_1, yo_dat_2, yo_dat_3, yo_dat_4)

#~ Use the *predict()* function to estimate model fit and standard error
preds.yo = predict(gamu1.12, newdata = yo_dat, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
fit = preds.yo$fit
upper = fit - CI_factor*preds.yo$se.fit
lower = fit + CI_factor*preds.yo$se.fit
yo_dat = cbind(yo_dat, fit, upper, lower)

# separate by log_qsdy at high and low value & tily at high and low value
yo.ldr.lWTmwk = subset(yo_dat, yo_dat$diurnal_range == dr.1 & yo_dat$WTmwk == WTmwk.1)
yo.hdr.lWTmwk = subset(yo_dat, yo_dat$diurnal_range == dr.9 & yo_dat$WTmwk == WTmwk.1)
yo.hdr.hWTmwk = subset(yo_dat, yo_dat$diurnal_range == dr.1 & yo_dat$WTmwk == WTmwk.9)
yo.ldr.hWTmwk = subset(yo_dat, yo_dat$diurnal_range == dr.9 & yo_dat$WTmwk == WTmwk.9)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for low and high flow for day of water year
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

tiff(filename = "figures/gam_upst_vis_WTmwk_dowy_dr.tiff", width = 10, height = 5, units = "in", res = 200)

# Plot the raw data and make the axes and labels aesthetically pleasing (margins = b,l,t,r)
par(mfrow = c(1,2), mar = c(2.5, 4.5, 2, 1) + 0.1, oma = c(1.5, 0, 0.5, 0))

# plot the actual data points separately for inundated and non-inundated
plot(u.idys$log_chla ~ u.idys$dowy, pch = 20, col = "green", las = 1, ylab = "", xlab = "", main = "Low water temperature")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Day of water year", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(yo.ldr.lWTmwk$dowy, max(yo.ldr.lWTmwk$dowy), rev(yo.ldr.lWTmwk$dowy), yo.ldr.lWTmwk$dowy[1]),
        y = c(yo.ldr.lWTmwk$lower, yo.ldr.lWTmwk$upper[length(yo.ldr.lWTmwk$upper)], rev(yo.ldr.lWTmwk$upper), yo.ldr.lWTmwk$lower[1]), border = NA,
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = yo.ldr.lWTmwk$dowy, y = yo.ldr.lWTmwk$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(yo.hdr.lWTmwk$dowy, max(yo.hdr.lWTmwk$dowy), rev(yo.hdr.lWTmwk$dowy), yo.hdr.lWTmwk$dowy[1]),
        y = c(yo.hdr.lWTmwk$lower, yo.hdr.lWTmwk$upper[length(yo.hdr.lWTmwk$upper)], rev(yo.hdr.lWTmwk$upper), yo.hdr.lWTmwk$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = yo.hdr.lWTmwk$dowy, y = yo.hdr.lWTmwk$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))


# plot the actual data points separately for inundated and non-inundated
plot(u.idys$log_chla ~ u.idys$dowy, pch = 20, col = "green", las = 1, ylab = "", xlab = "", main = "High water temperature")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Day of water year", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(yo.hdr.hWTmwk$dowy, max(yo.hdr.hWTmwk$dowy), rev(yo.hdr.hWTmwk$dowy), yo.hdr.hWTmwk$dowy[1]),
        y = c(yo.hdr.hWTmwk$lower, yo.hdr.hWTmwk$upper[length(yo.hdr.hWTmwk$upper)], rev(yo.hdr.hWTmwk$upper), yo.hdr.hWTmwk$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = yo.hdr.hWTmwk$dowy, y = yo.hdr.hWTmwk$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(yo.ldr.hWTmwk$dowy, max(yo.ldr.hWTmwk$dowy), rev(yo.ldr.hWTmwk$dowy), yo.ldr.hWTmwk$dowy[1]),
        y = c(yo.ldr.hWTmwk$lower, yo.ldr.hWTmwk$upper[length(yo.ldr.hWTmwk$upper)], rev(yo.ldr.hWTmwk$upper), yo.ldr.hWTmwk$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = yo.ldr.hWTmwk$dowy, y = yo.ldr.hWTmwk$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))

dev.off()

```




```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model for low and high water temperature for day of water year
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
library("plot3D")
library("ggplot2")

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# plot the actual data points separately for inundated and non-inundated
plot(u.idys$log_chla ~ u.idys$dowy, pch = 20, col = "gray", las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = "Day of water year", side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gam.lowwt$dowy, max(gam.lowwt$dowy), rev(gam.lowwt$dowy), gam.lowwt$dowy[1]),
        y = c(gam.lowwt$lower, gam.lowwt$upper[length(gam.lowwt$upper)], rev(gam.lowwt$upper), gam.lowwt$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = gam.lowwt$dowy, y = gam.lowwt$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gam.hghwt$dowy, max(gam.hghwt$dowy), rev(gam.hghwt$dowy), gam.hghwt$dowy[1]),
        y = c(gam.hghwt$lower, gam.hghwt$upper[length(gam.hghwt$upper)], rev(gam.hghwt$upper), gam.hghwt$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gam.hghwt$dowy, y = gam.hghwt$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))


# trying other ways to visualize

# surf3D(new.dat$log_qsdy, new.dat$WTmwk, new.dat$log_chla, colvar = new.dat$log_chla,
#        colkey = TRUE, box = TRUE, bty = "b", phi = 20, theta = 120)

#install.packages("visreg")
library("visreg")

visreg2d(gamu1.7, xvar='log_qsdy', yvar='WTmwk', scale='response', plot.type="image")
visreg2d(gamu1.7, xvar='log_qsdy', yvar='WTmwk', scale='response', plot.type="persp")

# ggplot(new.dat, aes(x = log_qsdy, y = WTmwk)) + geom_tile(aes(fill = fit))


```








## Cache
### Filter to cache
```{r}
cache <- alldata %>% filter(region == "cache")
```

```{r}
cache_scaled <- cache %>%
  mutate(across(.cols = c(log_idys,t_qsdy,diurnal_range:Sradmwk), ~ scale(.x)))
  
c.idys <- cache_scaled

cache_s_long <- pivot_longer(c.idys, cols = c(log_idys,t_qsdy,diurnal_range:Sradmwk), names_to = "Parameter", values_to = "Value")

ggplot(cache_s_long) + geom_violin(aes(x = Parameter, y = Value)) + facet_wrap(~Parameter, scales = "free")
```


### Model

```{r}
gamc0 <- gam(log_chla ~ 1, method = "REML", data = c.idys, family = "gaussian")
gamc1.0 <- gam(log_chla ~ s(t_qsdy, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.1 <- gam(log_chla ~ s(t_qsdy, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = c.idys, family = "gaussian")
gamc1.2 <- gam(log_chla ~ te(t_qsdy, dowy, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.3 <- gam(log_chla ~ te(log_idys, dowy, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.4 <- gam(log_chla ~ te(t_qsdy, dowy, by = inund_fac),  method = "REML", data = c.idys, family = "gaussian")
gamc1.5 <- gam(log_chla ~ s(WTmwk, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.6 <- gam(log_chla ~ s(WTmwk, by = inundation) + s(dowy, bs = "cc"), method = "REML", data = c.idys, family = "gaussian")
gamc1.7 <- gam(log_chla ~ s(t_qsdy, by = inundation) + s(WTmwk, bs = "cc"), method = "REML", data = c.idys, family = "gaussian")
gamc1.8 <- gam(log_chla ~ s(t_qsdy, by = inundation) + s(Sradmwk, bs = "cc"), method = "REML", data = c.idys, family = "gaussian")
gamc1.9 <- gam(log_chla ~ te(t_qsdy, WTmwk, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.10 <- gam(log_chla ~ te(t_qsdy, Sradmwk, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.11 <- gam(log_chla ~ te(log_idys, Sradmwk, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.12 <- gam(log_chla ~ te(log_idys, WTmwk, by = inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.13 <- gam(log_chla ~ te(t_qsdy, dowy), method = "REML", data = c.idys, family = "gaussian")
gamc1.14 <- gam(log_chla ~ te(log_idys, dowy), method = "REML", data = c.idys, family = "gaussian")

gamc1.15 <- gam(log_chla ~ s(total_inund_last_year, k = 6), method = "REML", data = c.idys, family = "gaussian")
gamc1.16 <- gam(log_chla ~ s(days_since_last_inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.17 <- gam(log_chla ~ s(diurnal_range), method = "REML", data = c.idys, family = "gaussian")
gamc1.18 <- gam(log_chla ~ s(WTrangemwk), method = "REML", data = c.idys, family = "gaussian")
gamc1.19 <- gam(log_chla ~ s(t_qsdy), method = "REML", data = c.idys, family = "gaussian")

gamc1.20 <- gam(log_chla ~ te(t_qsdy, total_inund_last_year), method = "REML", data = c.idys, family = "gaussian")
gamc1.21 <- gam(log_chla ~ te(t_qsdy, days_since_last_inundation), method = "REML", data = c.idys, family = "gaussian")
gamc1.22 <- gam(log_chla ~ te(t_qsdy, diurnal_range), method = "REML", data = c.idys, family = "gaussian")
gamc1.23 <- gam(log_chla ~ te(t_qsdy, WTrangemwk), method = "REML", data = c.idys, family = "gaussian")
gamc1.24 <- gam(log_chla ~ te(t_qsdy, WTmwk), method = "REML", data = c.idys, family = "gaussian")


resc <- (AIC(gamc0, gamc1.0, gamc1.1, gamc1.2, gamc1.3, gamc1.4, gamc1.5, gamc1.6, gamc1.7, gamc1.8, gamc1.9, gamc1.10, gamc1.11, gamc1.12, gamc1.13, gamc1.14,  gamc1.16, gamc1.17, gamc1.18, gamc1.19, gamc1.20, gamc1.21, gamc1.22, gamc1.23, gamc1.24))
resc %>% arrange(AIC)

# Best is 1.9

par(mfrow = c(2,2))
gam.check(gamc1.9)
gam.check(gamc1.10)
gam.check(gamc1.22)
gam.check(gamc1.2)

summary(gamc1.9) # 40% R2 0.329
summary(gamc1.10) # 36% R2 0.289
summary(gamc1.22) # 28.8% R2 0.248
summary(gamc1.2) # 37.6% 0.297

resid = residuals(gamc1.2)
acf(resid)
pacf(resid)

```

#### Model validation


```{r}

  # validate model gamc1.2 - log_chla ~ te(log_qsdy, dowy, by = inundation)
  # -----------------------------------------------------------------------
  # validate the final model by plotting residuals and save to a tiff
  tiff(filename = "figures/gam_model_validation_qsdy_yolo.tiff", width = 6, height = 9, units = "in", res = 100)
  resid = residuals(gamc1.2)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamc1.2), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(c.idys$t_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(c.idys$dowy, resid, xlab = "day of water year",   ylab = "Residuals")
  plot(c.idys$inundation, resid, xlab = "inundation flag",   ylab = "Residuals")
  par(op)
  dev.off()
```

#### Visualize gamc1.2
green = no inundation, blue = inundation
```{r}
# Determine the confidence interval to present in your figure and determine the
# t-distribution value of which to multiply the standard error
interval_value = 0.88
t_distribution_probability = 1.0 - (1-interval_value)/2
CI_factor = qt(t_distribution_probability, Inf)

#~ Generate a data frame with the sequence from the minimum to the maximum
# observed value of the predictor variable: day of water year
mod_q_range = range(na.omit(c.idys$t_qsdy))
lqsdy.05 = quantile(c.idys$t_qsdy, probs = 0.05)
lqsdy.95 = quantile(c.idys$t_qsdy, probs = 0.95)
#med.dowy = median(c.idys$dowy)
med.dowy = quantile(c.idys$dowy, probs = 0.5)
#qsdy.ser = seq(from = mod_q_range[1], to = mod_q_range[2], by = 0.05)
qsdy.ser = seq(from = lqsdy.05, to = lqsdy.95, by = 0.02)
n = length(qsdy.ser)
new_dat_c1 = data.frame(cbind(t_qsdy = qsdy.ser, dowy = rep(med.dowy, n)))
new_dat_c1$inundation = sample(c(0,1), n, replace = TRUE)
                       
#~ Use the *predict()* function to estimate model fit and standard error
preds = predict(gamc1.2, newdata = new_dat_c1, se.fit = TRUE)

# Multiply the standard error fit by the t-distribution value to estimate
# the upper and lower confidence interval limits
fit = preds$fit
upper = fit - CI_factor*preds$se.fit
lower = fit + CI_factor*preds$se.fit

new_dat_c1 = cbind(new_dat_c1, fit, upper, lower)

# separate by inundation flag
gam.ci0 = subset(new_dat_c1, new_dat_c1$inundation == 0)
gam.ci1 = subset(new_dat_c1, new_dat_c1$inundation == 1)

# separate actual data by inundation flag
act.ci0 = subset(c.idys, c.idys$inundation == 0)
act.ci1 = subset(c.idys, c.idys$inundation == 1)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~ Visualize the model
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Plot the raw data and make the axes and labels aesthetically pleasing
par(mfrow = c(1,1), mar = c(2.5, 4.5, 2,4.5) + 0.1, oma = c(1.5,0,0.5,0))

# plot the actual data points separately for inundated and non-inundated
colors = c("green", "blue")
plot(c.idys$log_chla ~ c.idys$t_qsdy, pch = 20, col = colors[c.idys$inundation], las = 1, ylab = "", xlab = "")
#mtext(text = "Log(Chlorophyll-a)", side = 4, line = 3)
mtext(text = expression(log[e]~"(Chlorophyll-a)"), side = 2, line = 2.5)
#y_labs = c(0.0001,0.0003,0.001, 0.003,0.01,0.033,0.1,0.33, 1, 3, 8, 20, 55, 150)[c(T,F)]
#axis(side = 4, at = log(y_labs), labels = y_labs, las=1)
mtext(text = expression(log[e]~"(Flow)"), side = 1, line = 2.25)

# Use the polygon function to add the CI
polygon(x = c(gam.ci0$t_qsdy, max(gam.ci0$t_qsdy), rev(gam.ci0$t_qsdy), gam.ci0$t_qsdy[1]),
        y = c(gam.ci0$lower, gam.ci0$upper[length(gam.ci0$upper)], rev(gam.ci0$upper), gam.ci0$lower[1]), border = NA, 
        col = rgb(20, 200, 20, alpha = 150, maxColorValue = 255))

# Use the lines() function to add the mean model prediction
lines(x = gam.ci0$t_qsdy, y = gam.ci0$fit, lwd = 2, col = rgb(20, 150, 20, alpha = 255, maxColorValue = 255))

# Use the polygon function to add the CI
polygon(x = c(gam.ci1$t_qsdy, max(gam.ci1$t_qsdy), rev(gam.ci1$t_qsdy), gam.ci1$t_qsdy[1]),
        y = c(gam.ci1$lower, gam.ci1$upper[length(gam.ci1$upper)], rev(gam.ci1$upper), gam.ci1$lower[1]), border = NA, 
        col = rgb(20, 20, 200, alpha = 150, maxColorValue=255))

# Use the lines() function to add the mean model prediction
lines(x = gam.ci1$t_qsdy, y = gam.ci1$fit, lwd = 2, col = rgb(20, 20, 150, alpha = 255, maxColorValue = 255))
```

