---
title: "make_gam_model"
author: "Catarina Pien"
date: "1/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
Variety of Modeling Techniques for Connectivity Synthesis
* LM
* GLS
* GLMM
* GAM
* GAMM
What will we choose?

Last updated 1/19/2022 after working with Pascale/Shruti with original dataset

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connectivity Modeling - Chlorophyll a

## Preparation

### Call packages
```{r}
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
```

### Read in data

Old - leaving in for now
```{r}
# covars <- read_csv("data_model/gam_data_covars.csv")

# temporary while waiting to clean up other data ------
# inundation_doy1998 <- read_csv("data_model/inundation_with_doy1998.csv")
# 
# covars <- readRDS("bayes_models/mod_covariates_complete.rds") %>% 
#   select(-yolo, -height_sac, -inund_days, -inun, -past_topped) %>%
#   left_join(inundation_doy1998, by = c("doy1998", "date"))
# 
# chla_all <- readRDS("bayes_models/Chla_all.rds")%>%
#     # only want chl
#     select(station, date, chl) %>%
#     filter(chl > 0 & station != '11455420') %>%
#     filter(complete.cases(.)) %>% # keep only complete records
#     arrange(station, date) %>%
#     # add a diff time as above for join purposes
#     mutate(doy1998 = as.numeric(difftime(date, as.Date("1998-01-01"), "day")) + 1,
#            station_id = as.numeric(as.factor(station))) %>%
#     mutate(log_chla = log(chl))
# 
# final_covars <- covars %>%
#     mutate(Q_1day = lag(flow_usgs_verona, 1), #this one is not working
#            Q_mwk = rollapply(flow_usgs_verona, 7, mean, align='right', partial=TRUE),
#            T_mwk = rollapply(daymet_tmax, 7, mean, align='right', partial=TRUE),
#            Srad_mwk = rollapply(daymet_srad, 7, mean, align='right', partial=TRUE)) %>%
#     rename(Q_sday = flow_usgs_verona)
# 
# chla_uniq_date = chla_all %>% group_by(doy1998) %>% sample_n(1)
```

```{r}
# chla_covars <- left_join(chla_uniq_date, final_covars, by=c("doy1998", "date"))
# 
# x <- na.omit(chla_covars) %>% select(log_chla, Q_sday, Q_1day, Q_mwk, T_mwk, Srad_mwk, inund_days, doy1998)
```

### Dataset with all chla (multiple stations)
```{r}
# gam_data_allstations <- left_join(chla_all, final_covars, by = c("doy1998", "date")) %>%
#   filter(chl > 0 & station != '11455420') %>%
#   filter(complete.cases(.)) %>%
#   mutate(fWY = factor(water_year),
#          doy = yday(date),
#          station_id = factor(station_id))
# 
# str(gam_data_allstations)
```

### Clean model dataset - current
```{r}
#4/8/2022
chla_covars <- read_csv("data_model/model_chla_covars_gam.csv") %>%
  mutate(log_chla = log(chlorophyll + 1))

rv_data <- chla_covars %>% filter(station_wq_chl == "657")
```

For linear models, if we want one observation per date, use for now until we modify the order. Some of the variables have changed though, lags not in here. 
```{r}
chla_single <- chla_covars %>%
  group_by(doy1998) %>% sample_n(1)

x <- na.omit(chla_single) %>% select(log_chla, Q_sday, WTmwk, Sradmwk, inund_days, doy1998)
```

## Modeling

### LM
All variables

```{r}
lm_first = lm(log_chla ~ Q_sday + Q_1day + Q_mwk + T_mwk + Srad_mwk + inund_days, data = x)
```

#### Check vif (variance inflation factor). 
vif > 4 should be eliminated step by step, removing variable with largest vif first.
```{r}
vif(lm_first)

# remove the covar with the biggest number; rerun vif
lm_firstb = lm(log_chla ~ Q_sday + Q_mwk + T_mwk + Srad_mwk + inund_days, data = x)
vif(lm_firstb)

# remove the next most correlated covar; rerun vif until all values below 3
lm_firstc = lm(log_chla ~ Q_sday + Srad_mwk + inund_days, data = x)
vif(lm_firstc)
```

#### dredge - a function to run all instances of a global model and evaluated
```{r}
# them based off AIC - output commented below
options(na.action = "na.fail")
dredge_results_fullmodel <- dredge(lm_first)
dredge_results <- dredge(lm_firstc)
```

#### Check for autocorrelation in model, predictor, response
```{r}
acf(lm_firstc$residuals)
acf(x$log_chla)
acf(x$Q_sday)
```

### GLS

#### Run Models 
Temperature
```{r}
gls.0 <- gls(log_chla ~ Q_sday + T_mwk + inund_days, na.action = na.omit, 
              data = x)
acf(gls.0$residuals)


gls.1 <- gls(log_chla ~ Q_sday + T_mwk + inund_days, na.action = na.omit, 
            data = x,
            correlation = corAR1(form =~ doy1998))
  
nresid = residuals(gls.1, type = "normalized")
acf(nresid)
  
gls.2 <- gls(log_chla ~ Q_sday + T_mwk + inund_days, na.action = na.omit, 
            data = x,
            correlation = corARMA(form =~ doy1998, p=1, q=1))
nresid2 = residuals(gls.2, type = "normalized")
acf(nresid2)

gls.3 <- gls(log_chla ~ Q_sday + T_mwk + inund_days, na.action = na.omit, 
             data = x,
            correlation = corARMA(form =~ doy1998, p=2, q=1))
nresid3 = residuals(gls.3, type = "normalized")
acf(nresid3)

# Compare models
anova(gls.0, gls.1, gls.2, gls.3)
```

Srad
```{r}
gls.4 <- gls(log_chla ~ Q_sday + Srad_mwk + inund_days, na.action = na.omit, data = x,
               correlation = corARMA(form =~ doy1998, p=1, q=1))
nresid4 = residuals(gls.4, type = "normalized")
acf(nresid4)

AIC(gls.4, gls.2) # Temperature better than Srad
```

#### Final Model 
Remove temperature: final model is just flow and inundation
```{r}
# dropping Temperature gives better AIC hence gls 5 is final model
gls.5 <- gls(log_chla ~ Q_sday + inund_days, na.action = na.omit, data = x,
               correlation = corARMA(form =~ doy1998, p=1, q=1))
gls.final = update(gls.5, method="ML")

# check residuals are not autocorrelated
nresid5 = residuals(gls.final, type = "normalized")
acf(nresid5)

# Get p-values
anova(gls.final)
```

```{r}
# interaction gives a worse AIC hence no interaction
gls.6 = gls(log_chla ~ inund_days + Q_sday + inund_days:Q_sday, na.action = na.omit, data = x, correlation = corARMA(form =~ doy1998, p=1, q=1))

AIC(gls.final, gls.6) # How do I check this??
```

#### Model validation
```{r}
# validate the final model by plotting residuals and save to a tiff
  #tiff(filename = "figures/gls_model_validation.tiff", width = 10, height = 6, units = "in", res = 300)
  op = par(mfrow = c(2, 3), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gls.final), nresid, xlab = "Fitted values", ylab = "Norm. Residuals")
  #plot(gls.final, add.smooth = FALSE, which = 1)
  # save residuals
  nresid = residuals(gls.final, type = "normalized")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(nresid, xlab = 'Norm. Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(nresid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(x$Q_sday, nresid, xlab = "Flow same day", ylab = "Norm. Residuals")
  plot(x$inund_days, nresid, xlab = "Consecutive inundation days", ylab = "Norm. Residuals")
  par(op)
```

### GLMM
```{r}
gls.7 = gls(log_chla ~ inund_days + Q_sday + , na.action = na.omit, data = x, correlation = corARMA(form =~ doy1998, p=1, q=1))
```



### GAMM
* Are we interested in month, season, day of water year as interaction terms? e.g. inund_days by month?

* tp is default
* cc = cyclic smooth
* by = doesn't need a smooth
* by = will make a smoother for each category

#### Models

```{r}
# Add in day of year
# Make water year a factor
# Remove outlier
gam_data <- rv_data %>%
  mutate(doy = yday(date),
         fWY = factor(water_year),
         inundation= factor(inundation),
         wateRshedTools::add_WYD(rv_data, datecolumn = "date")) %>%
  filter(chlorophyll < 20) %>%
  select(-DOY, -WY) %>%
  rename(dowy = DOWY)
str(gam_data)
 #------------------------


# Look at data
ggplot(gam_data) + geom_point(aes(date, chlorophyll))
ggplot(gam_data) + geom_point(aes(Sradmwk, WTmwk))
```

```{r}
# Flow, inundation days, water temperature, day of year
gam1 <- gam(log_chla~ s(Q_sday, k = 10) + s(inund_days, k = 15) + s(WTmwk, k = 10) + s(dowy, k = 10, bs = "cc") + s(water_year, k = 6), method = "REML", data = gam_data)

par(mfrow = c(2,2))
gam.check(gam1)
summary(gam1)
plot(gam1, pages = 1, all.terms = TRUE, se = TRUE)
gamplot <- plot(gam1, pages = 2, all.terms = TRUE, se = TRUE, ylab = "Chlorophyll (ug/L)")

# Interaction example
# Flow, inundation days, water temperature, day of year 
# Add flow by inundation days
gam2 <- gam(log_chla~s(inund_days, k = 12) + s(Q_sday, by = inundation, k = 6) + inundation +  s(WTmwk, k = 10) + s(dowy, k = 10, bs = "cc"), method = "REML", data = gam_data)

gam.check(gam2)
summary(gam2)
plot(gam2, pages = 1, all.terms = TRUE, se = TRUE, ylab = "log Chlorophyll (ug/L)")

# Single model
# Flow
gam3 <- gam(log_chla~ s(Q_sday, bs = "tp", k = 11), method = "REML", data = gam_data)

gam.check(gam3)
summary(gam3)
plot(gam3, pages = 1, all.terms = TRUE)

# Inundation days
gam4 <- gam(log_chla~ s(inund_days, bs = "tp", k = 12), method = "REML", data = gam_data)

gam.check(gam4)
summary(gam4)
plot(gam4, pages = 1, all.terms = TRUE)



```

```{r}
BIC(gam3, gam4)
```

##### Export plots for IEP/ISB presentation (Model gam1)
```{r}
png("figures/gamplotresults.png", width = 3.5, height = 3, units = "in", pointsize = 12, res = 300)
plot(gam1, pages = 2, all.terms = TRUE, se = TRUE, ylab = "log Chlorophyll")
dev.off()
png("figures/gamplotresults_flow_inund.png", width = 5.5, height = 3, units = "in", pointsize = 12, res = 300)
plot(gam1, pages = 1, all.terms = TRUE, se = TRUE, ylab = "log Chlorophyll")
dev.off()
```


#### Random effect
```{r}
# With station as random effect
gamm1 <- gam(log_chla~s(inund_days, k = 12) + s(Q_sday, k = 6) + s(station_id, bs = "re") , method = "REML", data = gam_data_allstations)

par(mfrow = c(2,2))
gam.check(gamm1)
summary(gamm1)
plot(gamm1, pages = 1, all.terms = TRUE, se = TRUE)
```

#### Zuur GAM
```{r}
gamm2 <- gamm(log_chla~s(inund_days, bs = "tp") + s(Q_sday, bs = "tp") + s(doy, bs = "cc"), random = list(station_id = ~1), method = "REML", data = gam_data_allstations)
plot(gamm2)
```


#### Interactions
```{r}
gam4 <- gam(log_chla~s(inund_days, k = 12) + s(Q_sday, k = 6) + s(Srad_mwk, k = 15) + te(doy, water_year, k = c(15,10), bs = "cc"), method = "REML", data = gam_data)
gam.check(gam4)
summary(gam4)
plot(gam4, pages = 1, all.terms = TRUE)
par(mfrow = c(1,1))
vis.gam(x = gam4,
        view = c("doy", "water_year"),
        color = "heat", plot.type = "contour")

gam5 <- gam(log_chla~ s(inund_days, k = 12) +s(Q_sday, k = 6)+s(Srad_mwk, by = fWY) + fWY, method = "REML", data = filter(gam_data, water_year>2012))

par(mfrow = c(2,2))
gam.check(gam5)
summary(gam5)
plot(gam5)

vis.gam(x = gam5,
        view = c("fWY", "inund_days"))
```

* I think this is similar to vif but for GAM
```{r}
gam_full <- gam(log_chla ~ s(Q_sday) + s(WTmwk) + s(Sradmwk) + s(inund_days), data = gam_data, method = "REML")

concurvity(gam_full, full = TRUE)
concurvity(gam_full, full = FALSE)
round(concurvity(gam_full, full = FALSE)$worst, 2)
```




* Shruti's notes
```{r}
# Global model call: lm(formula = log_chla ~ Q_sday + Srad_mwk + past_topped, data = x,
  #                       na.omit = TRUE)
  # ---
  #   Model selection table
  # (Int) pst_tpp     Q_sdy   Srd_mwk df   logLik   AICc delta weight
  # 4 0.5803 0.02551 5.329e-06            4 -523.349 1054.8  0.00  0.334
  # 8 0.4168 0.02381 6.584e-06 0.0003927  5 -522.356 1054.8  0.05  0.326
  # 2 0.6607 0.03325                      3 -524.781 1055.6  0.83  0.221
  # 6 0.5725 0.03332           0.0002395  4 -524.385 1056.8  2.07  0.119
  # 7 0.2002         1.706e-05 0.0006215  4 -530.340 1068.8 13.98  0.000
  # 3 0.4447         1.622e-05            3 -532.850 1071.7 16.97  0.000
  # 1 0.7595                              2 -564.603 1133.2 78.45  0.000
  # 5 0.6926                   0.0001823  3 -564.406 1134.9 80.08  0.000
  # Models ranked by AICc(x)

  # lm_0 = lm(log_chla ~ 1, data = x)

  #  num_rows = nrow(chla_all)

  # unique(covars$past_topped)

  # write out
  # if needed write to rds files:
  # write_rds(chla_all, "bayes_models/mod_chla_data.rds")
  # write_rds(covars, file = "bayes_models/mod_covariates_complete.rds")
  # cat("Model datasets saved in list:\nmodel_df$covars and model_df$chla_all")
```

