---
title: "visualize_gam_upstream"
author: "Catarina Pien"
date: '2022-10-05'
output: html_document
editor_options: 
  chunk_output_type: console
---
Cat
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

library(tidymv)
```

* validation plots
* actual data points
* grid of flow vs water temperature, colored by chlorophyll
* marker of the extent of data for each inundation phase
* same scale and axis range for each factor
* random effect plot (partial effect of each)
```{r}
load("data_model/gams_origdata.Rdata")
```

# Filter datasets to region
```{r}
upstream <- alldata %>% filter(region == "above")
```

# Model validation
```{r}
  # validate the final model by plotting residuals and save to a png to save space
  png(filename = "../figures/gamu6d.png", width = 7, height = 9, units = "in", res = 300)
  resid = residuals(gamu6d)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamu6d), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(upstream$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(upstream$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(upstream$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  #pacf(resid)
```

### PREDICT MODEL OUTPUTS FOR SIMULATED DATASET

## Visualize the Upstream Final Model as a 2x2 grid ignoring random effect of station

```{r}
# subset the upstream data frame each inundation factor
u.none = upstream %>% filter(inund_fac2 == "none")
u.shrt = upstream %>% filter(inund_fac2 == "short")
u.long = upstream %>% filter(inund_fac2 == "long")

# Get values 
u_table <- upstream %>% 
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() %>% unlist()

# divide all covars into this many intervals in simulated dataset
grid.size = 49

# calculate interval size for all 3 inundation factor simulation datasets
int.lqsdy = (u_table[2] - u_table[1])/(grid.size)
int.wtmwk = (u_table[4] - u_table[3])/(grid.size)

# create sequences of values for model inputs
seq.lqsdy = seq(u_table[1], u_table[2], by = int.lqsdy)
seq.wtmwk = seq(u_table[3], u_table[4], by = int.wtmwk)

# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)

# use i as subscript for log_qsdy
for (i in 1:(grid.size+1)) {
  # use j as subscript for WTmwk
  pred.none = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL), 
                          exclude_terms = "s(station)")
  pred.shrt = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL), 
                          exclude_terms = "s(station)")
  pred.long = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL), 
                          exclude_terms = "s(station)")
  chl.mat.none[i,] = pred.none$fit
  chl.mat.shrt[i,] = pred.shrt$fit
  chl.mat.long[i,] = pred.long$fit
}

# Color palette
# col_range = colorRampPalette(viridis(25, option = "turbo"))
# col_pal = col_range(1000)
col_range1 = colorRampPalette(c("darkorchid4","blueviolet","blue4","blue","deepskyblue2","deepskyblue","yellow2"))
col_range2 = colorRampPalette(c("yellow2","darkorange1","chocolate1","chocolate4","cornsilk2"))
col_range3 = colorRampPalette(c("cornsilk2","gray","lightgray","white"))
col_pal1 = col_range1(40)
col_pal2 = col_range2(240)
col_pal3 = col_range3(720)
col_pal = c(col_pal1, col_pal2, col_pal3)

# decide breaks for the color palette
# adjust to actual chlorophyll values existing in the dataset
# more than max is just one color; less than max is another
temp = cbind(chl.mat.none, chl.mat.shrt, chl.mat.long)
chl.ymin = min(upstream$chlorophyll)
chl.ymax = max(upstream$chlorophyll)
chla.min = min(min(temp), chl.ymin)
chla.max = max(max(temp), chl.ymax)
chla.int = (chl.ymax - chl.ymin)/998
breaks = c(chla.min, seq(chl.ymin, chl.ymax, chla.int), chla.max)

# calculating axes labels
lbl.lqsdy = round(seq(u_table[1], u_table[2], by = (u_table[2] - u_table[1])/5), digits = 2)
lbl.wtmwk = round(seq(u_table[3], u_table[4], by = (u_table[4] - u_table[3])/5), digits = 2)

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(u.none$log_qsdy, u.none$WTmwk), chull(u.none$log_qsdy, u.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(u.shrt$log_qsdy, u.shrt$WTmwk), chull(u.shrt$log_qsdy, u.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(u.long$log_qsdy, u.long$WTmwk), chull(u.long$log_qsdy, u.long$WTmwk)[1])
# get x,y coordinates of the hull
u.none.hull = cbind(u.none[cnvx.hull.none, 31], u.none[cnvx.hull.none, 17])
u.shrt.hull = cbind(u.shrt[cnvx.hull.shrt, 31], u.shrt[cnvx.hull.shrt, 17])
u.long.hull = cbind(u.long[cnvx.hull.long, 31], u.long[cnvx.hull.long, 17])

# constrain to 5-95 percentile values
u.none.hull = u.none.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < u_table[1], u_table[1], ifelse(log_qsdy > u_table[2], u_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < u_table[3], u_table[3], ifelse(WTmwk    > u_table[4], u_table[4], WTmwk))) %>%
  mutate(inund_fac2 = "none")

u.shrt.hull = u.shrt.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < u_table[1], u_table[1], ifelse(log_qsdy > u_table[2], u_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < u_table[3], u_table[3], ifelse(WTmwk    > u_table[4], u_table[4], WTmwk)))%>%
  mutate(inund_fac2 = "short")

u.long.hull = u.long.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < u_table[1], u_table[1], ifelse(log_qsdy > u_table[2], u_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < u_table[3], u_table[3], ifelse(WTmwk    > u_table[4], u_table[4], WTmwk)))%>%
  mutate(inund_fac2 = "long")

```


## MAKE THE ACTUAL FIGURE

```{r}
png(filename = "../figures/upstream_2dfigure.png", width = 12, height = 4, units = "in", res = 300)

par(mfrow = c(1, 3), mar = c(4, 2.75, 1, 2.5) + 0.1, oma = c(0, 2.25, 2, 2), las = 1, cex = 1)

###-------------- upstream: No inundation --------------###

#png(filename = "../figures/yo_none_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.none, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(u.none.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.none.hull[1, 1]), 
        y = c(u.none.hull[[2]], u.none.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.none.hull[1, 2], u.none.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = u.none$log_qsdy, y = u.none$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj = 0.5, las = 0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj = 0.5, las = 0)
mtext(text = "upstream: No inundation", side = 3,
      line = 0.75, adj = 0.5, las = 0, font = 1.5)

#dev.off()

###--------- upstream: short inundation (< 3 weeks) ---------###
#png(filename = "../figures/yo_shrt_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.shrt, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(u.shrt.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.shrt.hull[1, 1]), 
        y = c(u.shrt.hull[[2]], u.shrt.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.shrt.hull[1, 2], u.shrt.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = u.shrt$log_qsdy, y = u.shrt$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "upstream: short inundation (< 3 weeks)", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

#dev.off()


###--------- Upstream: long inundation (> 3 weeks) ---------###
#png(filename = "../figures/yo_long_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.long, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(u.long.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.long.hull[1, 1]), 
        y = c(u.long.hull[[2]], u.long.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.long.hull[1, 2], u.long.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = u.long$log_qsdy, y = u.long$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "upstream: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()
```

Make hulls
```{r}
u.shrt.hull2 <- data.frame(log_qsdy = c(u.shrt.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.shrt.hull[1, 1]), 
        WTmwk = c(u.shrt.hull[[2]], u.shrt.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.shrt.hull[1, 2], u.shrt.hull[1, 2])) %>%
  mutate(inund_fac2 = "short")


u.none.hull2 <- data.frame(log_qsdy = c(u.none.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.none.hull[1, 1]), 
        WTmwk = c(u.none.hull[[2]], u.none.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.none.hull[1, 2], u.none.hull[1, 2])) %>%
  mutate(inund_fac2 = "none")

u.long.hull2 <- data.frame(log_qsdy = c(u.long.hull[[1]], u_table[2],         u_table[2], u_table[1], u_table[1], u_table[2], u_table[2],         u.long.hull[1, 1]), 
        WTmwk = c(u.long.hull[[2]], u.long.hull[1, 2], u_table[4], u_table[4], u_table[3], u_table[3], u.long.hull[1, 2], u.long.hull[1, 2]))%>%
  mutate(inund_fac2 = "long")
```

Predictions
```{r}
pred.none = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "none")

pred.shrt = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "short")

pred.long = predict_gam(gamu6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "long")
 
pred_upstream <- rbind(pred.none, pred.shrt, pred.long)
hull_upstream <- rbind(u.shrt.hull2, u.none.hull2, u.long.hull2)

upstream_f <- upstream %>% 
  filter(log_qsdy < u_table[2] & log_qsdy > u_table[1],
         WTmwk < u_table[4] & WTmwk > u_table[3])
```

Plot code
```{r}
pred_upstream$inund_fac2 <- factor(pred_upstream$inund_fac2, levels = c("none", "short", "long"))
hull_upstream$inund_fac2 <- factor(hull_upstream$inund_fac2, levels = c("none", "short", "long"))

library(viridis)

ggplot() + 
  geom_tile(data = pred_upstream, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_upstream, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = upstream_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  #scale_fill_steps2(midpoint=2, n.breaks=12, low = "green", mid = "blue", high = "yellow")+
  #scale_fill_steps2(n.breaks=13, breaks = c(0, 0.25, 0.5, 0.75, 1, 1.5, 2, 2.5, 3, 4, 5, 10, 12),low = viridis::viridis(13)[1], mid = viridis::viridis(13)[8], high = viridis::viridis(13)[13])+
  #viridis::scale_fill_viridis(option = "turbo") +
 scale_fill_gradientn(colours = c("mediumpurple", "ghostwhite", "cyan","indianred1", "springgreen4", "bisque1", "mediumorchid4","palegreen", "magenta", "aquamarine", "yellow", "orange1","white", "thistle2", "brown"),
     breaks = c(0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8,0.9, 1, 1.5, 2,5,8, 12))+
  facet_wrap(~inund_fac2) + 
  theme_classic()
```




```{r}
ggplot() + 
  geom_tile(data = pred.shrt, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = u.shrt.hull2, aes(x = log_qsdy, y = WTmwk), fill = "grey", alpha = 0.3)+
  geom_point(data = u.shrt, aes(log_qsdy, WTmwk), color = "gray4", shape = 1) +
  #scale_fill_steps2(n.breaks=13, breaks = c(0, 0.25, 0.5, 0.75, 1, 1.5, 2, 2.5, 3, 4, 5, 10, 12),low = viridis::viridis(13)[1], mid = viridis::viridis(13)[8], high = viridis::viridis(13)[13])+
  #viridis::scale_fill_viridis(option = "turbo") +
 scale_fill_gradientn(colours = c("mediumpurple", "ghostwhite", "cyan","indianred1", "springgreen4", "bisque1", "mediumorchid4","palegreen", "magenta", "aquamarine", "yellow", "orange1","white", "brown"),
     breaks = c(0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8,0.9, 1, 1.5, 2,5, 12))+
  theme_classic()


ggplot() + 
  geom_tile(data = pred.long, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = u.long.hull2, aes(x = log_qsdy, y = WTmwk), fill = "grey", alpha = 0.3)+
  geom_point(data = u.long, aes(log_qsdy, WTmwk), color = "gray4", shape = 1) +
  #scale_fill_steps2(n.breaks=13, breaks = c(0, 0.25, 0.5, 0.75, 1, 1.5, 2, 2.5, 3, 4, 5, 10, 12),low = viridis::viridis(13)[1], mid = viridis::viridis(13)[8], high = viridis::viridis(13)[13])+
   scale_fill_gradientn(colours = c("mediumpurple", "ghostwhite", "cyan","indianred1", "springgreen4", "bisque1", "mediumorchid4","palegreen", "magenta", "aquamarine", "yellow", "orange1","white", "brown"),
     breaks = c(0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8,0.9, 1, 1.5, 2,5, 12))+
  theme_classic()
```

