---
title: "visualize_gam_yolo"
author: "Catarina Pien"
date: '2022-10-05'
output: html_document
---

Shruti

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* validation plots
* actual data points
* grid of flow vs water temperature, colored by chlorophyll
* marker of the extent of data for each inundation phase
* same scale and axis range for each factor
* random effect plot (partial effect of each)

## Load required packages

```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)

```


## Read in datasets and final models

```{r}
load("../data_model/gams_origdata.Rdata")

```

## Filter to Yolo Region

```{r}
yolo <- alldata %>% filter(region == "yolo")

```


### gamyo6d - Yolo final model validation

Model: gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + inund_fac2 + s(station, bs = "re"), method = "REML", data = yolo, family = "gaussian")

```{r}
  # validate the final model by plotting residuals and save to a png to save space
  png(filename = "../figures/gamyo6d.png", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamyo6d)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamyo6d), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(yolo$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(yolo$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(yolo$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  #pacf(resid)
```

### PREDICT MODEL OUTPUTS FOR SIMULATED DATASET

## Visualize the Yolo Final Model as a 2x2 grid ignoring random effect of station

```{r}

library(tidymv)

# subset the yolo data frame each inundation factor
yo.none = yolo %>% filter(inund_fac2 == "none")
yo.shrt = yolo %>% filter(inund_fac2 == "short")
yo.long = yolo %>% filter(inund_fac2 == "long")

# Get values 
y_table <- yolo %>% 
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() %>% unlist()

# divide all covars into this many intervals in simulated dataset
grid.size = 49

# calculate interval size for all 3 inundation factor simulation datasets
int.lqsdy = (y_table[2] - y_table[1])/(grid.size)
int.wtmwk = (y_table[4] - y_table[3])/(grid.size)

# create sequences of values for model inputs
seq.lqsdy = seq(y_table[1], y_table[2], by = int.lqsdy)
seq.wtmwk = seq(y_table[3], y_table[4], by = int.wtmwk)

# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)

# use i as subscript for log_qsdy
for (i in 1:(grid.size+1)) {
  # use j as subscript for WTmwk
  pred.none = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL), 
                          exclude_terms = "s(station)")
  pred.shrt = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL), 
                          exclude_terms = "s(station)")
  pred.long = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL), 
                          exclude_terms = "s(station)")
  chl.mat.none[i,] = pred.none$fit
  chl.mat.shrt[i,] = pred.shrt$fit
  chl.mat.long[i,] = pred.long$fit
}

# Color palette
# col_range = colorRampPalette(viridis(25, option = "turbo"))
# col_pal = col_range(1000)
col_range1 = colorRampPalette(c("darkorchid4","blueviolet","blue4","blue","deepskyblue2","deepskyblue","yellow2"))
col_range2 = colorRampPalette(c("yellow2","darkorange1","chocolate1","chocolate4","cornsilk2"))
col_range3 = colorRampPalette(c("cornsilk2","gray","lightgray","white"))
col_pal1 = col_range1(40)
col_pal2 = col_range2(240)
col_pal3 = col_range3(720)
col_pal = c(col_pal1, col_pal2, col_pal3)

# decide breaks for the color palette
# adjust to actual chlorophyll values existing in the dataset
# more than max is just one color; less than max is another
temp = cbind(chl.mat.none, chl.mat.shrt, chl.mat.long)
chl.ymin = min(yolo$chlorophyll)
chl.ymax = max(yolo$chlorophyll)
chla.min = min(min(temp), chl.ymin)
chla.max = max(max(temp), chl.ymax)
chla.int = (chl.ymax - chl.ymin)/998
breaks = c(chla.min, seq(chl.ymin, chl.ymax, chla.int), chla.max)

# calculating axes labels
lbl.lqsdy = round(seq(y_table[1], y_table[2], by = (y_table[2] - y_table[1])/5), digits = 2)
lbl.wtmwk = round(seq(y_table[3], y_table[4], by = (y_table[4] - y_table[3])/5), digits = 2)

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(yo.none$log_qsdy, yo.none$WTmwk), chull(yo.none$log_qsdy, yo.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(yo.shrt$log_qsdy, yo.shrt$WTmwk), chull(yo.shrt$log_qsdy, yo.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(yo.long$log_qsdy, yo.long$WTmwk), chull(yo.long$log_qsdy, yo.long$WTmwk)[1])
# get x,y coordinates of the hull
yo.none.hull = cbind(yo.none[cnvx.hull.none, 31], yo.none[cnvx.hull.none, 17])
yo.shrt.hull = cbind(yo.shrt[cnvx.hull.shrt, 31], yo.shrt[cnvx.hull.shrt, 17])
yo.long.hull = cbind(yo.long[cnvx.hull.long, 31], yo.long[cnvx.hull.long, 17])

# constrain to 5-95 percentile values
yo.none.hull = yo.none.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))
yo.shrt.hull = yo.shrt.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))
yo.long.hull = yo.long.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))

```


## MAKE THE ACTUAL FIGURE

```{r}
png(filename = "../figures/yo_2dfigure.png", width = 12, height = 4, units = "in", res = 300)

par(mfrow = c(1, 3), mar = c(4, 2.75, 1, 2.5) + 0.1, oma = c(0, 2.25, 2, 2), las = 1, cex = 1)

###-------------- Yolo: No inundation --------------###

#png(filename = "../figures/yo_none_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.none, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.none.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.none.hull[1, 1]), 
        y = c(yo.none.hull[[2]], yo.none.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.none.hull[1, 2], yo.none.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.none$log_qsdy, y = yo.none$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj = 0.5, las = 0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj = 0.5, las = 0)
mtext(text = "Yolo: No inundation", side = 3,
      line = 0.75, adj = 0.5, las = 0, font = 1.5)

#dev.off()

###--------- Yolo: short inundation (< 3 weeks) ---------###
#png(filename = "../figures/yo_shrt_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.shrt, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.shrt.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.shrt.hull[1, 1]), 
        y = c(yo.shrt.hull[[2]], yo.shrt.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.shrt.hull[1, 2], yo.shrt.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.shrt$log_qsdy, y = yo.shrt$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: short inundation (< 3 weeks)", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

#dev.off()


###--------- Downstream: long inundation (> 3 weeks) ---------###
#png(filename = "../figures/yo_long_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.long, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.long.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.long.hull[1, 1]), 
        y = c(yo.long.hull[[2]], yo.long.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.long.hull[1, 2], yo.long.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.long$log_qsdy, y = yo.long$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

#dev.off()
```





