---
title: "visualize_gam_yolo"
author: "Catarina Pien"
date: '2022-10-05'
output: html_document
editor_options: 
  chunk_output_type: console
---

Shruti

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* validation plots
* actual data points
* grid of flow vs water temperature, colored by chlorophyll
* marker of the extent of data for each inundation phase
* same scale and axis range for each factor
* random effect plot (partial effect of each)

## Load required packages

```{r}
#detach(gam)
library(dplyr)
library(readr)
library(lubridate)
library(glue)
library(tidyr)
library(zoo)
library(car)
library(MuMIn)
library(psych)
library(mgcv)
library(ggplot2)
library(viridis)
library(nlme)
library(kader)#cuberoot
library(mgcv)

```


## Read in datasets and final models

```{r}
load("data_model/gams_origdata.Rdata")

```

## Filter to Yolo Region

```{r}
yolo <- alldata %>% filter(region == "yolo")

```


### gamyo6d - Yolo final model validation

Model: gam(log_chla ~ te(log_qsdy, WTmwk, by = inund_fac2) + inund_fac2 + s(station, bs = "re"), method = "REML", data = yolo, family = "gaussian")

```{r}
  # validate the final model by plotting residuals and save to a png to save space
  png(filename = "../figures/gamyo6d.png", width = 7, height = 9, units = "in", res = 200)
  resid = residuals(gamyo6d)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamyo6d), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  plot(yolo$log_qsdy, resid, xlab = "Flow",   ylab = "Residuals")
  plot(yolo$WTmwk, resid, xlab = "WTmwk",   ylab = "Residuals")
  plot(yolo$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  #pacf(resid)
```

### PREDICT MODEL OUTPUTS FOR SIMULATED DATASET

## Visualize the Yolo Final Model as a 2x2 grid ignoring random effect of station

```{r}

library(tidymv)

# subset the yolo data frame each inundation factor
yo.none = yolo %>% filter(inund_fac2 == "none")
yo.shrt = yolo %>% filter(inund_fac2 == "short")
yo.long = yolo %>% filter(inund_fac2 == "long")

# Get values 
y_table <- yolo %>% 
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() %>% unlist()

# divide all covars into this many intervals in simulated dataset
grid.size = 49

# calculate interval size for all 3 inundation factor simulation datasets
int.lqsdy = (y_table[2] - y_table[1])/(grid.size)
int.wtmwk = (y_table[4] - y_table[3])/(grid.size)

# create sequences of values for model inputs
seq.lqsdy = seq(y_table[1], y_table[2], by = int.lqsdy)
seq.wtmwk = seq(y_table[3], y_table[4], by = int.wtmwk)

# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)

# use i as subscript for log_qsdy
for (i in 1:(grid.size+1)) {
  # use j as subscript for WTmwk
  pred.none = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL), 
                          exclude_terms = "s(station)")
  pred.shrt = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL), 
                          exclude_terms = "s(station)")
  pred.long = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL), 
                          exclude_terms = "s(station)")
  chl.mat.none[i,] = pred.none$fit
  chl.mat.shrt[i,] = pred.shrt$fit
  chl.mat.long[i,] = pred.long$fit
}

# Color palette
# col_range = colorRampPalette(viridis(25, option = "turbo"))
# col_pal = col_range(1000)
col_range1 = colorRampPalette(c("darkorchid4","blueviolet","blue4","blue","deepskyblue2","deepskyblue","yellow2"))
col_range2 = colorRampPalette(c("yellow2","darkorange1","chocolate1","chocolate4","cornsilk2"))
col_range3 = colorRampPalette(c("cornsilk2","gray","lightgray","white"))
col_pal1 = col_range1(40)
col_pal2 = col_range2(240)
col_pal3 = col_range3(720)
col_pal = c(col_pal1, col_pal2, col_pal3)

# decide breaks for the color palette
# adjust to actual chlorophyll values existing in the dataset
# more than max is just one color; less than max is another
temp = cbind(chl.mat.none, chl.mat.shrt, chl.mat.long)
chl.ymin = min(yolo$chlorophyll)
chl.ymax = max(yolo$chlorophyll)
chla.min = min(min(temp), chl.ymin)
chla.max = max(max(temp), chl.ymax)
chla.int = (chl.ymax - chl.ymin)/998
breaks = c(chla.min, seq(chl.ymin, chl.ymax, chla.int), chla.max)

# calculating axes labels
lbl.lqsdy = round(seq(y_table[1], y_table[2], by = (y_table[2] - y_table[1])/5), digits = 2)
lbl.wtmwk = round(seq(y_table[3], y_table[4], by = (y_table[4] - y_table[3])/5), digits = 2)

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(yo.none$log_qsdy, yo.none$WTmwk), chull(yo.none$log_qsdy, yo.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(yo.shrt$log_qsdy, yo.shrt$WTmwk), chull(yo.shrt$log_qsdy, yo.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(yo.long$log_qsdy, yo.long$WTmwk), chull(yo.long$log_qsdy, yo.long$WTmwk)[1])
# get x,y coordinates of the hull
yo.none.hull = cbind(yo.none[cnvx.hull.none, 31], yo.none[cnvx.hull.none, 17])
yo.shrt.hull = cbind(yo.shrt[cnvx.hull.shrt, 31], yo.shrt[cnvx.hull.shrt, 17])
yo.long.hull = cbind(yo.long[cnvx.hull.long, 31], yo.long[cnvx.hull.long, 17])

# constrain to 5-95 percentile values
yo.none.hull = yo.none.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))
yo.shrt.hull = yo.shrt.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))
yo.long.hull = yo.long.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < y_table[1], y_table[1], ifelse(log_qsdy > y_table[2], y_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < y_table[3], y_table[3], ifelse(WTmwk    > y_table[4], y_table[4], WTmwk)))

```


## MAKE THE ACTUAL FIGURE

```{r}
#png(filename = "../figures/yo_2dfigure.png", width = 12, height = 4, units = "in", res = 300)

par(mfrow = c(1, 3), mar = c(4, 2.75, 1, 2.5) + 0.1, oma = c(0, 2.25, 2, 2), las = 1, cex = 1)

###-------------- Yolo: No inundation --------------###

#png(filename = "../figures/yo_none_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.none, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.none.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.none.hull[1, 1]), 
        y = c(yo.none.hull[[2]], yo.none.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.none.hull[1, 2], yo.none.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.none$log_qsdy, y = yo.none$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj = 0.5, las = 0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj = 0.5, las = 0)
mtext(text = "Yolo: No inundation", side = 3,
      line = 0.75, adj = 0.5, las = 0, font = 1.5)

#dev.off()

###--------- Yolo: short inundation (< 3 weeks) ---------###
#png(filename = "../figures/yo_shrt_2dfigure.png", width = 4, height = 4, units = "in", res = 300)
# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.shrt.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.shrt.hull[1, 1]), 
        y = c(yo.shrt.hull[[2]], yo.shrt.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.shrt.hull[1, 2], yo.shrt.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.shrt$log_qsdy, y = yo.shrt$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: short inundation (< 3 weeks)", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

#dev.off()


###--------- Downstream: long inundation (> 3 weeks) ---------###
#png(filename = "../figures/yo_long_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.long, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(yo.long.hull[[1]], y_table[2],         y_table[2], y_table[1], y_table[1], y_table[2], y_table[2],         yo.long.hull[1, 1]), 
        y = c(yo.long.hull[[2]], yo.long.hull[1, 2], y_table[4], y_table[4], y_table[3], y_table[3], yo.long.hull[1, 2], yo.long.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = yo.long$log_qsdy, y = yo.long$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "Yolo: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)
image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.shrt, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))



#dev.off()
```

## Cat's code: same axes
### Make inverse hulls from Shruti's hulls
```{r}
# short
yo.shrt.hull2 <- data.frame(
  log_qsdy = c(yo.shrt.hull$log_qsdy, y_table["maxQ"], y_table["maxQ"], y_table["minQ"], y_table["minQ"],
               y_table["maxQ"], y_table["maxQ"], yo.shrt.hull$log_qsdy[1]), 
       
  WTmwk = c(yo.shrt.hull$WTmwk, yo.shrt.hull$WTmwk[1], y_table["maxWT"], y_table["maxWT"], y_table["minWT"],
            y_table["minWT"], yo.shrt.hull$WTmwk[1], yo.shrt.hull$WTmwk[1])) %>%
  
  mutate(inund_fac2 = "short")

# none
yo.none.hull2 <- data.frame(log_qsdy = c(yo.none.hull[[1]], y_table["maxQ"],         y_table["maxQ"], y_table["minQ"], y_table["minQ"], y_table["maxQ"], y_table["maxQ"],         yo.none.hull$log_qsdy[1]), 
        WTmwk = c(yo.none.hull$WTmwk, yo.none.hull$WTmwk[1], y_table["maxWT"], y_table["maxWT"], y_table["minWT"], y_table["minWT"], yo.none.hull$WTmwk[1], yo.none.hull$WTmwk[1])) %>%
  mutate(inund_fac2 = "none")

#long
yo.long.hull2 <- data.frame(log_qsdy = c(yo.long.hull[[1]], y_table["maxQ"],         y_table["maxQ"], y_table["minQ"], y_table["minQ"], y_table["maxQ"], y_table["maxQ"],         yo.long.hull$log_qsdy[1]), 
        WTmwk = c(yo.long.hull$WTmwk, yo.long.hull$WTmwk[1], y_table["maxWT"], y_table["maxWT"], y_table["minWT"], y_table["minWT"], yo.long.hull$WTmwk[1], yo.long.hull$WTmwk[1]))%>%
  mutate(inund_fac2 = "long")
```

### Predictions
```{r}
pred.none = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "none")

pred.shrt = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "short")

pred.long = predict_gam(gamyo6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "long")
 
pred_yolo <- rbind(pred.none, pred.shrt, pred.long)
hull_yolo <- rbind(yo.shrt.hull2, yo.none.hull2, yo.long.hull2)

yolo_f <- yolo %>% 
  filter(log_qsdy < y_table[2] & log_qsdy > y_table[1],
         WTmwk < y_table[4] & WTmwk > y_table[3])
```

### Plot 
```{r}
pred_yolo$inund_fac2 <- factor(pred_yolo$inund_fac2, levels = c("none", "short", "long"))
hull_yolo$inund_fac2 <- factor(hull_yolo$inund_fac2, levels = c("none", "short", "long"))

library(viridis)

(yolo_same <- ggplot() + 
  geom_tile(data = pred_yolo, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_yolo, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = yolo_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  #scale_fill_steps2(midpoint=2, n.breaks=12, low = "green", mid = "blue", high = "yellow")+
  #scale_fill_steps2(n.breaks=13, breaks = c(0, 0.25, 0.5, 0.75, 1, 1.5, 2, 2.5, 3, 4, 5, 10, 12),low = viridis::viridis(13)[1], mid = viridis::viridis(13)[8], high = viridis::viridis(13)[13])+
  #viridis::scale_fill_viridis(option = "turbo") +
 scale_fill_gradientn(colours = c("mediumpurple", "ghostwhite", "cyan","indianred1", "springgreen4", "bisque1", "mediumorchid4","palegreen", "magenta", "aquamarine", "yellow", "orange1","white", "thistle2", "brown"),
     breaks = c(0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8,0.9, 1, 1.5, 2,5,8, 12))+
  facet_wrap(~inund_fac2) +
  labs(x = "log Daily Flow (cfs)", y = "Mean weekly water temperature (°C)", fill = "log chlorophyll a (ug/L)", title = "Yolo") +
  theme_classic() +
  theme(legend.position = "top"))
```


Write plot
```{r}
png(here::here("figures","yolo_gam_samescales.png"), width = 7, height = 4, units = "in", res = 300)
yolo_same
dev.off()
```


## Cat's code: different axes

### Make new hulls based on new mins and maxes customized to inundation factor

* Short: log_qsdy = 10-11/WTmwk: 7.5-13
* Long: log_qsdy = 10.25-11/WTmwk: 7.5-15
```{r}
y_table2 <- yolo %>% 
  group_by(inund_fac2) %>%
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() 

maxQshort = 11.1
minQshort = 7.6
minWTshort = 7.5
maxWTshort = 18.2

maxQlong = 11.5
minQlong = 8
minWTlong = 8.9
maxWTlong = 18

maxQnone = 8.5
minQnone = 4.4
minWTnone = 7.4
maxWTnone = 19.5

f_createpolygon <- function(df_hull, maxQ, minQ, maxWT, minWT, inundationfactor) {
  hull <- data.frame(
    log_qsdy = c(df_hull$log_qsdy, maxQ,          maxQ, minQ, minQ, maxQ, maxQ),
    
    WTmwk = c(df_hull$WTmwk,    df_hull$WTmwk[1], maxWT, maxWT, minWT, minWT, df_hull$WTmwk[1]),
    inund_fac2 = inundationfactor) 
  
  return(hull)
  }

yo.shrt.hullnew <- f_createpolygon(yo.shrt.hull, maxQshort, minQshort, maxWTshort, minWTshort, "short")
yo.long.hullnew <- f_createpolygon(yo.long.hull, maxQlong, minQlong, maxWTlong, minWTlong, "long")
yo.none.hullnew <- f_createpolygon(yo.none.hull, maxQnone, minQnone, maxWTnone, minWTnone, "none")

hull_yolonew <- rbind(yo.shrt.hullnew, yo.none.hullnew, yo.long.hullnew) 
```

```{r}
ggplot(yo.long.hullnew) + geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))
ggplot(yo.shrt.hullnew) + geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))
ggplot(yo.none.hullnew) + 
  geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))

```

### Filter predictions
```{r}
pred_yolonew <- pred_yolo %>%
  filter(case_when(inund_fac2 == "none" ~ log_qsdy >= minQnone & log_qsdy <= maxQnone,
                   inund_fac2 == "short" ~ log_qsdy >= minQshort & log_qsdy <= maxQshort,
                   inund_fac2 == "long" ~ log_qsdy >= minQlong & log_qsdy <= maxQlong)) %>%
  filter(case_when(inund_fac2 == "none" ~ WTmwk >= minWTnone & WTmwk <= maxWTnone,
                   inund_fac2 == "short" ~ WTmwk >= minWTshort & WTmwk <= maxWTshort,
                   inund_fac2 == "long" ~ WTmwk >= minWTlong & WTmwk <= maxWTlong)) %>%
  filter(fit<5) %>%
  mutate(chl = 10^fit)

max(pred_yolonew$log_qsdy)
```

### Plot
```{r}
pred_yolonew$inund_fac2 <- factor(pred_yolonew$inund_fac2, levels = c("none", "short", "long"))
hull_yolonew$inund_fac2 <- factor(hull_yolonew$inund_fac2, levels = c("none", "short", "long"))

(yolo_diff <- ggplot() + 
  geom_tile(data = pred_yolonew, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_yolonew, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = yolo_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  viridis::scale_fill_viridis(option = "turbo") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  facet_wrap(~inund_fac2, scales = "free") + 
  labs(x = "log Daily Flow (cfs)", y = "Mean weekly water temperature (°C)", fill = "log chlorophyll a (ug/L)", title = "Yolo") +
  theme_classic() +
  theme(legend.position = "top"))
```

Write plot
```{r}
png(here::here("figures","yolo_gam_diffscales.png"), width = 7, height = 4, units = "in", res = 300)
yolo_diff
dev.off()

save(yolo_diff, yolo_same, file = here::here("data_model", "yolo_gam_plots.RData"))
```
