---
title: "visualize_gam_downstream"
author: "Catarina Pien"
date: '2022-10-05'
output: html_document
editor_options: 
  chunk_output_type: console
---
Pascale
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
* validation plots
* actual data points
* grid of flow vs water temperature, colored by chlorophyll
* marker of the extent of data for each inundation phase
* same scale and axis range for each factor
* random effect plot (partial effect of each)
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidymv)
library(gratia)
library(mgcv)
library(patchwork)
```

```{r}
load("data_model/gams_origdata.Rdata")
```

## Filter datasets to region
```{r}
downstream <- alldata %>% filter(region == "below")
```

## Model validation
```{r}
  # validate the final model by plotting residuals and save to a png to save space
  png(filename = "../figures/gamd6d.png", width = 7, height = 9, units = "in", res = 300)
  resid = residuals(gamd6d)
  op = par(mfrow = c(3, 2), mar = c(5, 4, 1, 2), cex = 1.2)
  # Plot 1: Residuals vs. Fitted values; should be centered around 0
  plot(fitted(gamd6d), resid, xlab = "Fitted values", ylab = "Residuals")
  # Plot 2: histogram of the residuals; should be centered around 0
  hist(resid, xlab = 'Residuals', main = "")
  # Plot 3: is there autocorrelation in the residuals?
  acf(resid)
  # Plots 4,5,6: the Residuals vs. all the predictors; should be centered around 0
  # none of these seem to work, and I dont see where the connection between the data values and residuals are being made...
  plot(downstream$log_qsdy, resid, xlab = "Flow", ylab = "Residuals")
  plot(downstream$WTmwk, resid, xlab = "WTmwk", ylab = "Residuals")
  plot(downstream$inund_fac2, resid, xlab = "inundation factor",   ylab = "Residuals")
  par(op)
  dev.off()
  
  #the plots produced are (from left-to-right, top-to-bottom)
  appraise(gamd6d)

#a quantile-quantile (QQ) plot of deviance residuals,
#a scatterplot of deviance residuals against the linear predictor,
#a histogram of deviance residuals, and
#a scatterplot of observed vs fitted values
  
  draw(gamd6d, residuals = TRUE)
  
```


### PREDICT MODEL OUTPUTS FOR SIMULATED DATASET

#### Visualize the Upstream Final Model as a 2x2 grid ignoring random effect of station

```{r}
# subset the downstream data frame each inundation factor
d.none = downstream %>% filter(inund_fac2 == "none")
d.shrt = downstream %>% filter(inund_fac2 == "short")
d.long = downstream %>% filter(inund_fac2 == "long")

# Get values 
d_table <- downstream %>% 
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() %>% unlist()

# divide all covars into this many intervals in simulated dataset
grid.size = 49

# calculate interval size for all 3 inundation factor simulation datasets
int.lqsdy = (d_table[2] - d_table[1])/(grid.size)
int.wtmwk = (d_table[4] - d_table[3])/(grid.size)

# create sequences of values for model inputs
seq.lqsdy = seq(d_table[1], d_table[2], by = int.lqsdy)
seq.wtmwk = seq(d_table[3], d_table[4], by = int.wtmwk)

# arrange the predictions in a grid.size x grid.size matrix ready to display as an image
# create a matrix to hold the value
chl.mat.none = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.shrt = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)
chl.mat.long = matrix(data = NA, nrow = grid.size+1, ncol = grid.size+1)

# use i as subscript for log_qsdy
for (i in 1:(grid.size+1)) {
  # use j as subscript for WTmwk
  pred.none = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL), 
                          exclude_terms = "s(station)")
  pred.shrt = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL), 
                          exclude_terms = "s(station)")
  pred.long = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy[i], WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL), 
                          exclude_terms = "s(station)")
  chl.mat.none[i,] = pred.none$fit
  chl.mat.shrt[i,] = pred.shrt$fit
  chl.mat.long[i,] = pred.long$fit
}

# Color palette
# col_range = colorRampPalette(viridis(25, option = "turbo"))
# col_pal = col_range(1000)
col_range1 = colorRampPalette(c("darkorchid4","blueviolet","blue4","blue","deepskyblue2","deepskyblue","yellow2"))
col_range2 = colorRampPalette(c("yellow2","darkorange1","chocolate1","chocolate4","cornsilk2"))
col_range3 = colorRampPalette(c("cornsilk2","gray","lightgray","white"))
col_pal1 = col_range1(40)
col_pal2 = col_range2(240)
col_pal3 = col_range3(720)
col_pal = c(col_pal1, col_pal2, col_pal3)

# decide breaks for the color palette
# adjust to actual chlorophyll values existing in the dataset
# more than max is just one color; less than max is another
temp = cbind(chl.mat.none, chl.mat.shrt, chl.mat.long)
chl.ymin = min(downstream$chlorophyll)
chl.ymax = max(downstream$chlorophyll)
chla.min = min(min(temp), chl.ymin)
chla.max = max(max(temp), chl.ymax)
chla.int = (chl.ymax - chl.ymin)/998
breaks = c(chla.min, seq(chl.ymin, chl.ymax, chla.int), chla.max)

# calculating axes labels
lbl.lqsdy = round(seq(d_table[1], d_table[2], by = (d_table[2] - d_table[1])/5), digits = 2)
lbl.wtmwk = round(seq(d_table[3], d_table[4], by = (d_table[4] - d_table[3])/5), digits = 2)

# calculate the outer hull coordinates for log_qsdy and WTmwk - gives the relevant row numbers
cnvx.hull.none = c(chull(d.none$log_qsdy, d.none$WTmwk), chull(d.none$log_qsdy, d.none$WTmwk)[1])
cnvx.hull.shrt = c(chull(d.shrt$log_qsdy, d.shrt$WTmwk), chull(d.shrt$log_qsdy, d.shrt$WTmwk)[1])
cnvx.hull.long = c(chull(d.long$log_qsdy, d.long$WTmwk), chull(d.long$log_qsdy, d.long$WTmwk)[1])
# get x,y coordinates of the hull
d.none.hull = cbind(d.none[cnvx.hull.none, 31], d.none[cnvx.hull.none, 17])
d.shrt.hull = cbind(d.shrt[cnvx.hull.shrt, 31], d.shrt[cnvx.hull.shrt, 17])
d.long.hull = cbind(d.long[cnvx.hull.long, 31], d.long[cnvx.hull.long, 17])

# constrain to 5-95 percentile values
d.none.hull = d.none.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[1], d_table[1], ifelse(log_qsdy > d_table[2], d_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < d_table[3], d_table[3], ifelse(WTmwk    > d_table[4], d_table[4], WTmwk))) %>%
  mutate(inund_fac2 = "none")

d.shrt.hull = d.shrt.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[1], d_table[1], ifelse(log_qsdy > d_table[2], d_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < d_table[3], d_table[3], ifelse(WTmwk    > d_table[4], d_table[4], WTmwk)))%>%
  mutate(inund_fac2 = "short")

d.long.hull = d.long.hull %>% 
  mutate(log_qsdy = ifelse(log_qsdy < d_table[1], d_table[1], ifelse(log_qsdy > d_table[2], d_table[2], log_qsdy))) %>% 
  mutate(WTmwk    = ifelse(WTmwk    < d_table[3], d_table[3], ifelse(WTmwk    > d_table[4], d_table[4], WTmwk)))%>%
  mutate(inund_fac2 = "long")

```


#### MAKE THE ACTUAL FIGURE

```{r}
png(filename = "../figures/downstream_2dfigure.png", width = 12, height = 4, units = "in", res = 300)

par(mfrow = c(1, 3), mar = c(4, 2.75, 1, 2.5) + 0.1, oma = c(0, 2.25, 2, 2), las = 1, cex = 1)

###-------------- downstream: No inundation --------------###

#png(filename = "../figures/yo_none_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.none, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(d.none.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.none.hull[1, 1]), 
        y = c(d.none.hull[[2]], d.none.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.none.hull[1, 2], d.none.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = d.none$log_qsdy, y = d.none$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj = 0.5, las = 0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj = 0.5, las = 0)
mtext(text = "downstream: No inundation", side = 3,
      line = 0.75, adj = 0.5, las = 0, font = 1.5)

#dev.off()

###--------- downstream: short inundation (< 3 weeks) ---------###
#png(filename = "../figures/yo_shrt_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.shrt, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(d.shrt.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.shrt.hull[1, 1]), 
        y = c(d.shrt.hull[[2]], d.shrt.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.shrt.hull[1, 2], d.shrt.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = d.shrt$log_qsdy, y = d.shrt$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "downstream: short inundation (< 3 weeks)", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

#dev.off()


###--------- downstream: long inundation (> 3 weeks) ---------###
#png(filename = "../figures/yo_long_2dfigure.png", width = 4, height = 4, units = "in", res = 300)

image(x = seq.lqsdy, y = seq.wtmwk, z = chl.mat.long, col = col_pal, breaks = breaks,
      ylab = "", xlab = "", axes = F, ylim = range(seq.wtmwk), xlim = range(seq.lqsdy))

# create breaks and labels for x and y axes
axis(side = 1, at = lbl.lqsdy, labels = lbl.lqsdy)
axis(side = 2, at = lbl.wtmwk, labels = lbl.wtmwk)

# dull the outside polygon where data does not exist
polygon(x = c(d.long.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.long.hull[1, 1]), 
        y = c(d.long.hull[[2]], d.long.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.long.hull[1, 2], d.long.hull[1, 2]), 
        col = gray(1, 0.7), border = NA)
points(x = d.long$log_qsdy, y = d.long$WTmwk, col = "white")

mtext(text = "Log of same day flow", side = 1,
      line = 2.5, adj=0.5, las=0)
mtext(text = "Weekly mean water temperature", side = 2,
      line = 3, adj=0.5, las=0)
mtext(text = "downstream: long inundation", side = 3,
      line = 0.75, adj=0.5, las=0, font=2)

dev.off()
```

Make inverse hulls
```{r}
d.shrt.hull2 <- data.frame(log_qsdy = c(d.shrt.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.shrt.hull[1, 1]), 
        WTmwk = c(d.shrt.hull[[2]], d.shrt.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.shrt.hull[1, 2], d.shrt.hull[1, 2])) %>%
  mutate(inund_fac2 = "short")


d.none.hull2 <- data.frame(log_qsdy = c(d.none.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.none.hull[1, 1]), 
        WTmwk = c(d.none.hull[[2]], d.none.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.none.hull[1, 2], d.none.hull[1, 2])) %>%
  mutate(inund_fac2 = "none")

d.long.hull2 <- data.frame(log_qsdy = c(d.long.hull[[1]], d_table[2],         d_table[2], d_table[1], d_table[1], d_table[2], d_table[2],         d.long.hull[1, 1]), 
        WTmwk = c(d.long.hull[[2]], d.long.hull[1, 2], d_table[4], d_table[4], d_table[3], d_table[3], d.long.hull[1, 2], d.long.hull[1, 2]))%>%
  mutate(inund_fac2 = "long")
```

## Cat's code - same axes
### Predictions
```{r}
pred.none = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "none", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "none")

pred.shrt = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "short", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "short")

pred.long = predict_gam(gamd6d, values = list(log_qsdy = seq.lqsdy, WTmwk = seq.wtmwk, inund_fac2 = "long", station = NULL, by = int.lqsdy), 
                          exclude_terms = "s(station)") %>%
  mutate(inund_fac2 = "long")
 
pred_downstream <- rbind(pred.none, pred.shrt, pred.long)
hull_downstream <- rbind(d.shrt.hull2, d.none.hull2, d.long.hull2)

downstream_f <- downstream %>% 
  filter(log_qsdy < d_table[2] & log_qsdy > d_table[1],
         WTmwk < d_table[4] & WTmwk > d_table[3])
```

### Plot code: same 
```{r}
pred_downstream$inund_fac2 <- factor(pred_downstream$inund_fac2, levels = c("none", "short", "long"))
hull_downstream$inund_fac2 <- factor(hull_downstream$inund_fac2, levels = c("none", "short", "long"))

library(viridis)

(downstream_same <- ggplot() + 
  geom_tile(data = pred_downstream, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_downstream, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = downstream_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  #scale_fill_steps2(midpoint=2, n.breaks=12, low = "green", mid = "blue", high = "yellow")+
  #scale_fill_steps2(n.breaks=13, breaks = c(0, 0.25, 0.5, 0.75, 1, 1.5, 2, 2.5, 3, 4, 5, 10, 12),low = viridis::viridis(13)[1], mid = viridis::viridis(13)[8], high = viridis::viridis(13)[13])+
  # viridis::scale_fill_viridis(option = "turbo") +
  scale_fill_gradientn(colours = c("mediumpurple", "ghostwhite", "cyan","indianred1", "springgreen4", "bisque1", "mediumorchid4","palegreen", "magenta", "aquamarine", "yellow", "orange1","white", "thistle2", "brown"),
                       breaks = c(-5, -1, 0, 0.5, 0.6, 0.7, 0.8,0.9, 1, 1.5, 2,5,8, 12))+
     # guide = guide_legend(keywidth = 2))+
  facet_wrap(~inund_fac2) + 
  labs(x = "log Daily Flow (cfs)", y = "Mean weekly water temperature (°C)", fill = "log chlorophyll a (ug/L)", title = "Downstream") +
  theme_classic() +
  theme(legend.position = "top"))
```

Write plot
```{r}
png(here::here("figures","downstream_gam_samescales.png"), width = 7, height = 4, units = "in", res = 300)
downstream_same
dev.off()
```



## Cat's code - Different axes
```{r}
d_table2 <- downstream %>% 
  group_by(inund_fac2) %>%
  summarize(minQ = quantile(log_qsdy, 0.05),
            maxQ = quantile(log_qsdy, 0.95),
            minWT = quantile(WTmwk, 0.05),
            maxWT = quantile(WTmwk, 0.95)) %>%
  as.data.frame() 

maxQshort = 11.5
minQshort = 10
minWTshort = 8
maxWTshort = 15

maxQlong = 11.5
minQlong = 10.9
minWTlong = 9
maxWTlong = 16

maxQnone = 11.5
minQnone = 8.5
minWTnone = 7.9
maxWTnone = 16.6

f_createpolygon <- function(df_hull, maxQ, minQ, maxWT, minWT, inundationfactor) {
  hull <- data.frame(
    log_qsdy = c(df_hull$log_qsdy, maxQ,          maxQ, minQ, minQ, maxQ, maxQ),
    
    WTmwk = c(df_hull$WTmwk,    df_hull$WTmwk[1], maxWT, maxWT, minWT, minWT, df_hull$WTmwk[1]),
    inund_fac2 = inundationfactor) 
  
  return(hull)
  }

d.shrt.hullnew <- f_createpolygon(d.shrt.hull, maxQshort, minQshort, maxWTshort, minWTshort, "short")
d.long.hullnew <- f_createpolygon(d.long.hull, maxQlong, minQlong, maxWTlong, minWTlong, "long")
d.none.hullnew <- f_createpolygon(d.none.hull, maxQnone, minQnone, maxWTnone, minWTnone, "none")

hull_downstreamnew <- rbind(d.shrt.hullnew, d.none.hullnew, d.long.hullnew) 
```

```{r}
ggplot(d.long.hullnew) + geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))
ggplot(d.shrt.hullnew) + geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))
ggplot(d.none.hullnew) + 
  geom_polygon(aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(aes(x = log_qsdy, y = WTmwk))

```

### Filter predictions
```{r}
pred_downstreamnew <- pred_downstream %>%
  filter(case_when(inund_fac2 == "none" ~ log_qsdy >= minQnone & log_qsdy <= maxQnone,
                   inund_fac2 == "short" ~ log_qsdy >= minQshort & log_qsdy <= maxQshort,
                   inund_fac2 == "long" ~ log_qsdy >= minQlong & log_qsdy <= maxQlong)) %>%
  filter(case_when(inund_fac2 == "none" ~ WTmwk >= minWTnone & WTmwk <= maxWTnone,
                   inund_fac2 == "short" ~ WTmwk >= minWTshort & WTmwk <= maxWTshort,
                   inund_fac2 == "long" ~ WTmwk >= minWTlong & WTmwk <= maxWTlong)) %>%
  mutate(chl = exp^fit)

max(pred_downstreamnew$log_qsdy)
```

### Plot
```{r}
pred_downstreamnew$inund_fac2 <- factor(pred_downstreamnew$inund_fac2, levels = c("none", "short", "long"))
hull_downstreamnew$inund_fac2 <- factor(hull_downstreamnew$inund_fac2, levels = c("none", "short", "long"))

(downstream_diff <- ggplot() + 
  geom_tile(data = pred_downstreamnew, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_downstreamnew, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = downstream_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  viridis::scale_fill_viridis(option = "turbo") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  facet_wrap(~inund_fac2, scales = "free") + 
  labs(x = "log Daily Flow (cfs)", y = "Mean weekly water temperature (°C)", fill = "log chlorophyll a (ug/L)", title = "Downstream") +
  theme_classic() +
  theme(legend.position = "top"))
```

Write plot
```{r}
png(here::here("figures","downstream_gam_diffscales.png"), width = 7, height = 4, units = "in", res = 300)
downstream_diff
dev.off()

save(downstream_diff, downstream_same, file = here::here("data_model", "downstream_gam_plots.RData"))
```

## Cat's code: same y axis 
```{r}
maxWTall = 16.6
minWTall = 7.9

d.shrt.hullsamey <- f_createpolygon(d.shrt.hull, maxQshort, minQshort, maxWTall, minWTall, "short")
d.long.hullsamey <- f_createpolygon(d.long.hull, maxQlong, minQlong, maxWTall, minWTall, "long")
d.none.hullsamey <- f_createpolygon(d.none.hull, maxQnone, minQnone, maxWTall, minWTall, "none")

hull_downstreamsamey<- rbind(d.shrt.hullsamey, d.none.hullsamey, d.long.hullsamey) 
```

```{r}
pred_downstreamsamey <- pred_downstreamnew <- pred_downstream %>%
  filter(case_when(inund_fac2 == "none" ~ log_qsdy >= minQnone & log_qsdy <= maxQnone,
                   inund_fac2 == "short" ~ log_qsdy >= minQshort & log_qsdy <= maxQshort,
                   inund_fac2 == "long" ~ log_qsdy >= minQlong & log_qsdy <= maxQlong)) %>%
  filter(case_when(inund_fac2 == "none" ~ WTmwk >= minWTall & WTmwk <= maxWTall,
                   inund_fac2 == "short" ~ WTmwk >= minWTall & WTmwk <= maxWTall,
                   inund_fac2 == "long" ~ WTmwk >= minWTall & WTmwk <= maxWTall)) %>%
  filter(fit>0) %>% 
  mutate(chl = 10^fit)
```


```{r}
pred_downstreamsamey$inund_fac2 <- factor(pred_downstreamsamey$inund_fac2, levels = c("none", "short", "long"))
hull_downstreamsamey$inund_fac2 <- factor(hull_downstreamsamey$inund_fac2, levels = c("none", "short", "long"))

(downstream_diffsamey <- ggplot() + 
  geom_tile(data = pred_downstreamsamey, aes(log_qsdy, WTmwk, fill = fit)) +
  geom_polygon(data = hull_downstreamsamey, aes(x = log_qsdy, y = WTmwk), fill = "white", alpha = 0.7)+
  geom_point(data = downstream_f, aes(log_qsdy, WTmwk), color = "gray6", shape = 1) +
  viridis::scale_fill_viridis(option = "turbo") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  facet_wrap(~inund_fac2, scales = "free_x") + 
  labs(x = "log Daily Flow (cfs)", y = "Mean weekly water temperature (°C)", fill = "log chl a\n(ug/L)", title = "downstream") +
  theme_classic() +
  theme(legend.position = "right"))
```

```{r}
png(here::here("figures","downstream_gam_sameyscales.png"), width = 7.5, height = 4, units = "in", res = 300)
downstream_diffsamey
dev.off()

save(downstream_diff, downstream_same, downstream_diffsamey, file = here::here("data_model", "downstream_gam_plots.RData"))
```





## Pascale's code

### Customizing plots in gratia
```{r}
draw(gamd6d) & theme_bw()

p1 <- draw(gamd6d, select = "te(log_qsdy,WTmwk):inund_fac2long")
p2 <- draw(gamd6d, select = "te(log_qsdy,WTmwk):inund_fac2short")
p3 <- draw(gamd6d, select = "te(log_qsdy,WTmwk):inund_fac2none")

p1 + p2 + p3 + plot_layout(ncol = 3)

# evaluate the smooths
sm <- smooth_estimates(gamd6d) %>%
  add_confint()

# add partial residuals to data
downstream <- downstream %>%
  add_partial_residuals(gamd6d)

# one smoother at a time, plotted by covar
sm %>%
  filter(smooth == "te(log_qsdy,WTmwk):inund_fac2long") %>%
  ggplot() +
  geom_rug(aes(x = log_qsdy),
           data = downstream,
           sides = "b", length = grid::unit(0.02, "npc")) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, x = log_qsdy),
              alpha = 0.2) +
  geom_point(aes(x = log_qsdy, y = `te(log_qsdy,WTmwk):inund_fac2long`),
             data = downstream, cex = 1.5, colour = "steelblue3") +
  geom_line(aes(x = log_qsdy, y = est), lwd = 1.2) +
  labs(y = "Partial effect", title = "Long Inundation")

# add color and size for station and temp

plt <- sm %>%
  filter(smooth == "te(log_qsdy,WTmwk):inund_fac2long") %>%
  ggplot() +
  geom_rug(aes(x = log_qsdy),
           data = downstream,
           sides = "b", length = grid::unit(0.02, "npc")) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, x = log_qsdy),
              alpha = 0.2) +
  geom_line(aes(x = log_qsdy, y = est), lwd = 1.2) +
  labs(y = "Partial effect", title = "s(x2)")

plt +
  geom_point(aes(x = log_qsdy, y = `te(log_qsdy,WTmwk):inund_fac2long`,
                 colour = station), # <-- map fac to colour aesthetic
             data = downstream, cex = 1.5)

plt +
  geom_point(aes(x = log_qsdy, y = `te(log_qsdy,WTmwk):inund_fac2long`,
                 colour = WTmwk, size = WTmwk), # <-- map fac to colour aesthetic
             data = downstream, alpha = 0.3) +   # <-- deleted cex
  scale_colour_viridis_c(option = "plasma")

# seems to work better if subset ahead instead of filter

none_sm <- subset(sm, inund_fac2 == "none")
none_dat <- subset(downstream, inund_fac2 == "none")
colnames(none_dat)[5] <- "est"

short_sm <- subset(sm, inund_fac2 == "short")
short_dat <- subset(downstream, inund_fac2 == "short")
colnames(short_dat)[5] <- "est"

long_sm <- subset(sm, inund_fac2 == "long")
long_dat <- subset(downstream, inund_fac2 == "long")
colnames(long_dat)[5] <- "est"

plot_long <- ggplot() +
  geom_rug(aes(x = log_qsdy),
           data = long_sm,
           sides = "b", length = grid::unit(0.02, "npc")) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, x = log_qsdy),
              alpha = 0.2, data = long_sm) +
  geom_line(aes(x = log_qsdy, y = est), lwd = 1.2, data = long_sm) +
  labs(y = "Partial effect", title = "Long Inundation") +
  xlab("Sacramento River outflow") +
  theme_bw()

plot_long + geom_point(aes(x = log_qsdy, y = est,
                 colour = WTmwk, size = WTmwk), data = long_dat, alpha = 0.3) +   
  scale_colour_viridis_c(option = "plasma")

plot_long + geom_point(aes(x = log_qsdy, y = est,
                 colour = station), data = long_dat, alpha = 0.3) 
# OR
plot_long <- ggplot() +
  geom_rug(aes(x = WTmwk),
           data = long_sm,
           sides = "b", length = grid::unit(0.02, "npc")) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, x = WTmwk),
              alpha = 0.2, data = long_sm) +
  geom_line(aes(x = WTmwk, y = est), lwd = 1.2, data = long_sm) +
  labs(y = "Partial effect", title = "Long Inundation") +
  xlab("Sacramento River temperature") +
  theme_bw()



png("hgam_results_temp.png", bg = "transparent", width = 11, height = 20, units = "in", pointsize = 12, res = 350)

plot_1 + plot_2 + plot_3 + plot_4 + plot_layout(nrow = 4)

dev.off()
```

# station
```{r}
ggplot() +
  geom_point(aes(x = chlorophyll, y = `s(station)`, color = station),data = downstream)

sm_657 <- subset(sm, station == "657")
dat_657 <- subset(downstream, station == "657")
colnames(dat_657)[5] <- "est"


sm %>%
  filter (smooth == "s(station)") %>%
  ggplot()+
  geom_point(aes(x = station, y = est)) +
  geom_errorbar(aes(x = station, ymin = est-se, ymax = est+se))+
  theme_bw()

ggplot()+
  geom_boxplot(aes(x = station, y = chlorophyll), downstream)
```
